\section{The Athanasian Creed}
\fancyhead[RE,LO]{The Athanasian Creed}
\rub{The Athanasian Creed shall be sung or said, standing, replacing the Apostles' Creed for the Major Hours on Christmas Day, Epiphany Day, Saint Matthias, Easter Day, Ascension Day, Whitsunday, Trinity Sunday, and on the Feasts of St. John Baptist, St. James, St. Matthew, Sts. Simon \& Jude, and St.Andrew.}
\elcol{
\lett{W}{hosoever} will be saved: before all things it is necessary that he hold the Catholick Faith.\par
Which Faith except every one do keep whole and undefiled: without doubt he shall perish everlastingly.
\par
And the Catholick Faith is this: That we worship one God in Trinity, and Trinity in Unity;\par
Neither confounding the Persons: nor dividing the Substance.\par
For there is one Person of the Father, another of the Son: and another of the Holy Ghost.\par
But the Godhead of the Father, of the Son, and of the Holy Ghost, is all one: the Glory equal, the Majesty co-eternal.\par
Such as the Father is, such is the Son: and such is the Holy Ghost.\par
The Father uncreate, the Son uncreate: and the Holy Ghost uncreate.\par
The Father incomprehensible, the Son incomprehensible: and the Holy Ghost incomprehensible.\par
The Father eternal, the Son eternal: and the Holy Ghost eternal.\par
And yet they are not three eternals: but one eternal.\par
As also there are not three incomprehensibles, nor three uncreated: but one uncreated, and one incomprehensible.\par
So likewise the Father is Almighty, the Son Almighty: and the Holy Ghost Almighty.\par
And yet they are not three Almighties: but one Almighty.\par
So the Father is God, the Son is God: and the Holy Ghost is God.\par
And yet they are not three Gods: but one God.\par
So likewise the Father is Lord, the Son Lord: and the Holy Ghost Lord.\par
And yet not three Lords: but one Lord.\par
For like as we are compelled by the Christian verity: to acknowledge every Person by himself to be God and Lord;\par
So are we forbidden by the Catholick Religion: to say there be three Gods, or three Lords.\par
The Father is made of none: neither created, nor begotten.\par
The Son is of the Father alone: not made, nor created, but begotten.\par
The Holy Ghost is of the Father (alone): neither made, nor created, nor begotten, but proceeding.\par
So there is one Father, not three Fathers; one Son, not three Sons: one Holy Ghost, not three Holy Ghosts.\par
And in this Trinity none is afore, or after other: none is greater, or less than another;\par
But the whole three Persons are co-eternal together: and co-equal.\par
So that in all things, as is aforesaid: the Unity in Trinity, and the Trinity in Unity is to be worshipped.\par
He therefore that will be saved: must thus think of the Trinity.
\par
Furthermore it is necessary to everlasting salvation: that he also believe rightly the Incarnation of our Lord Jesus Christ.\par
For the right Faith is that we believe and confess: that our Lord Jesus Christ, the Son of God, is God and Man;\par
God, of the Substance of the Father, begotten before the worlds: and Man, of the Substance of his Mother, born in the world;\par
Perfect God, and Perfect Man: of a reasonable soul and human flesh subsisting;\par
Equal to the Father, as touching his Godhead: and inferior to the Father, as touching his Manhood.\par
Who although he be God and Man: yet he is not two, but one Christ;\
One, not by conversion of the Godhead into flesh: but by taking of the Manhood into God;\par
One altogether, not by confusion of Substance: but by unity of Person.\par
For as the reasonable soul and flesh is one man: so God and Man is one Christ.\par
Who suffered for our salvation: descended into hell, rose again the third day from the dead.\par
He ascended into heaven, he sitteth on the right hand of the Father, God Almighty: from whence he shall come to judge the quick and the dead.\par
At whose coming all men shall rise again with their bodies: and shall give account for their own works.\par
And they that have done good shall go into life everlasting: and they that have done evil into everlasting fire.
\par
This is the Catholick Faith: which except a man believe faithfully, he cannot be saved.\par
\par
Glory be to the Father, and to the Son: and to the Holy Ghost;
As it was in the beginning, is now, and ever shall be: world without end. Amen.
}
{\lett{Q}{uicúmque} vult salvus esse, * ante ómnia opus est, ut téneat cathólicam fidem:\par
Quam nisi quisque íntegram inviolatámque serváverit, * absque dúbio in {\ae}térnum períbit.
\par
Fides autem cathólica h{\ae}c est: * ut unum Deum in Trinitáte, et Trinitátem in unitáte venerémur.\par
Neque confundéntes persónas, * neque substántiam separántes.\par
Alia est enim persóna Patris, ália Fílii, * ália Spíritus Sancti:\par
Sed Patris, et Fílii, et Spíritus Sancti una est divínitas, * {\ae}quális glória, co{\ae}térna majéstas.\par
Qualis Pater, talis Fílius, * talis Spíritus Sanctus.\par
Increátus Pater, increátus Fílius, * increátus Spíritus Sanctus.\par
Imménsus Pater, imménsus Fílius, * imménsus Spíritus Sanctus.\par
{\ae}térnus Pater, {\ae}térnus Fílius, * {\ae}térnus Spíritus Sanctus.\par
Et tamen non tres {\ae}térni, * sed unus {\ae}térnus.\par
Sicut non tres increáti, nec tres imménsi, * sed unus increátus, et unus imménsus.\par
Simíliter omnípotens Pater, omnípotens Fílius, * omnípotens Spíritus Sanctus.\par
Et tamen non tres omnipoténtes, * sed unus omnípotens.\par
Ita Deus Pater, Deus Fílius, * Deus Spíritus Sanctus.\par
Ut tamen non tres Dii, * sed unus est Deus.\par
Ita Dóminus Pater, Dóminus Fílius, * Dóminus Spíritus Sanctus.\par
Et tamen non tres Dómini, * sed unus est Dóminus.\par
Quia, sicut singillátim unamquámque persónam Deum ac Dóminum confitéri christiána veritáte compéllimur: * ita tres Deos aut Dóminos dícere cathólica religióne prohibémur.\par
Pater a nullo est factus: * nec creátus, nec génitus.\par
Fílius a Patre solo est: * non factus, nec creátus, sed génitus.\par
Spíritus Sanctus a Patre (solo): * non factus, nec creátus, nec génitus, sed procédens.\par
Unus ergo Pater, non tres Patres: unus Fílius, non tres Fílii: * unus Spíritus Sanctus, non tres Spíritus Sancti.\par
Et in hac Trinitáte nihil prius aut postérius, nihil majus aut minus: * sed tot{\ae} tres persón{\ae} co{\ae}térn{\ae} sibi sunt et co{\ae}quáles.\par
Ita ut per ómnia, sicut jam supra dictum est, * et únitas in Trinitáte, et Trínitas in unitáte veneránda sit.\par
Qui vult ergo salvus esse, * ita de Trinitáte séntiat.
\par
Sed necessárium est ad {\ae}térnam salútem, * ut Incarnatiónem quoque Dómini nostri Jesu Christi fidéliter credat.\par
Est ergo fides recta ut credámus et confiteámur, * quia Dóminus noster Jesus Christus, Dei Fílius, Deus et homo est.\par
Deus est ex substántia Patris ante sǽcula génitus: * et homo est ex substántia matris in sǽculo natus.\par
Perféctus Deus, perféctus homo: * ex ánima rationáli et humána carne subsístens.\par
{\ae}quális Patri secúndum divinitátem: * minor Patre secúndum humanitátem.\par
Qui licet Deus sit et homo, * non duo tamen, sed unus est Christus.\par
Unus autem non conversióne divinitátis in carnem, * sed assumptióne humanitátis in Deum.\par
Unus omníno, non confusióne substánti{\ae}, * sed unitáte persón{\ae}.\par
Nam sicut ánima rationális et caro unus est homo: * ita Deus et homo unus est Christus.\par
Qui passus est pro salúte nostra: descéndit ad ínferos: * tértia die resurréxit a mórtuis.\par
Ascéndit ad c{\ae}los, sedet ad déxteram Dei Patris omnipoténtis: * inde ventúrus est judicáre vivos et mórtuos.\par
Ad cujus advéntum omnes hómines resúrgere habent cum corpóribus suis; * et redditúri sunt de factis própriis ratiónem.\par
Et qui bona egérunt, ibunt in vitam {\ae}térnam: * qui vero mala, in ignem {\ae}térnum.
\par
H{\ae}c est fides cathólica, * quam nisi quisque fidéliter firmitérque credíderit, salvus esse non póterit.\par
\par
Glória Patri, et Fílio, * et Spirítui Sancto: Sicut erat in princípio, et nunc, et semper, * et in sǽcula s{\ae}culórum. Amen.
}