\section{Benediction of the Blessed Sacrament}
\fancyhead[RE,LO]{Benediction of the Blessed Sacrament}
\rub{When the priest brings the Sacrament out of the tabernacle, all kneeling, singing the following hymn:}
\subsection{O Salutaris Hostia}
\elcol{\lett{O}{} Saving Victim op'ning wide * The Gate of heav'n to man below,}{\lett{O}{} salutáris Hóstia, * Qu{\ae} caeli pandis óstium:}
\elcol{Our foes press on from ev'ry side, * Thine aid supply, thy strength bestow.}{Bella premunt hostília, * Da robur, fer auxílium.}
\elcol{All praise and thanks to thee ascend * For evermore, blest One in Three;}{Uni trinóque Dómino * Sit sempitérna glória,}
\elcol{O grant us life that shall not end, * In our true native land with thee. Amen.}
{Qui vitam sine término * Nobis donet in pátria. Amen.}
\subsection{Tantum Ergo Sacramentum}
\rub{There may be a time for silence. Then, the following hymn, or another suitable devotion, is sung:}
\elcol{\lett{T}{herefore} we, before Him bending, * This great Sacrament revere;}{\lett{T}{ántum} ergo Sacraméntum * Venerémur cérnui}
\elcol{Types and shadows have their ending, * For the newer rite is here;}{Et antíquum documéntum, * Nóvo cédat rítui,}
\elcol{Faith, our outward sense befriending, * Makes our inward vision clear.}{Prǽstet fídes suppleméntum * Sénsuum deféctui.}
\elcol{Glory let us give and blessing * To the Father and the Son,}{Genitóri, Genitóque, * Laus et jubilátio;}
\elcol{Honour, thanks, and praise addressing, * While eternal ages run;}{Sálus, hónor, vírtus quoque * Sit et benedíctio.}
\elcol{Ever too His love confessing * Who from One with Both is One. Amen.\\}{Procedénti ab unóque * Cómpar sit laudátio. Amen.\\}
\elcol{\textit{\scshape ℣.} Thou gavest them bread from heaven. (Alleluia)}{℣. Panem de c{\ae}lis pr{\ae}stitísti eis. (Allelúia)}
\elcol{\textit{\scshape ℟.} Containing in itself all sweetness. (Alleluia)\\}{℟. Omne delectaméntum in se habéntem. (Allelúia)\\}
\elcol{Let us pray:}{Orémus:}
\elcol{O God, who in this wonderful Sacrament hast left us a perpetual Memorial of Thy Passion: Grant us, we beseech thee, so to venerate the Sacred Mysteries of thy Body and Blood, that we may ever perceive within ourselves the fruit of thy redemption; who livest and reignest, world without end. \textit{Amen.}}{Deus, qui nobis sub sacraménto mirábili, passiónis tu{\ae} memóriam reliquísti: tríbue, quǽsumus, ita nos córporis et sánguinis tui sacra mystéria venerári, ut redemptiónis tu{\ae} fructum in nobis iúgiter sentiámus. Qui vivis et regnas in s{\ae}cula s{\ae}culorum. \textit{Amen.}}
\rub{While the Priest blesses the congregation with the monstrance, it is pious to pray during the benediction:}
\begin{center}
	Blessed, praised, hallowed, and adored be Jesus in the most Holy Sacrament.
\end{center}
\subsection{Divine Praises}
\rub{After placing the monstrance on the altar, the priest says the Divine Praises, all kneeling.}
\elcoldent{Blessed be God.}{Benedíctus Deus.}
\elcoldent{Blessed be His Holy Name.}{Benedíctum Nomen Sanctum eius.}
\elcoldent{Blessed be Jesus Christ, true God and true Man.}{Benedíctus Iesus Christus, verus Deus et verus homo.}
\elcoldent{Blessed be the Name of Jesus.}{Benedíctum Nomen Iesu.}
\elcoldent{Blessed be His Most Sacred Heart.}{Benedíctum Cor eius sacratíssimum.}
\elcoldent{Blessed be Jesus in the Most Holy Sacrament of the Altar.}{Benedíctus Iesus in sanctíssimo altáris Sacraménto.}
\elcoldent{Blessed be God the Holy Ghost, the Comforter.}{Benedíctus Sanctus Spíritus, Paraclítus.}
\elcoldent{Blessed be the Great Mother of God, Mary most Holy.}{Benedícta magna Mater Dei, María sanctíssima.}
\elcoldent{Blessed be the name of Mary, Virgin and Mother.}{Benedíctum nomen Marí{\ae}, Vírginis et Matris.}
\elcoldent{Blessed be Saint Joseph, her most chaste spouse.}{Benedíctus sanctus Ioseph, eius castíssimus Sponsus.}
\elcoldent{Blessed be God in His Angels and in His Saints.}{Benedíctus Deus in Angelis suis, et in Sanctis suis.}
\rub{The Blessed Sacrament is replaced in the Tabernacle. All stand and pray:}
\elcol{\textit{Ant.} Let us forever adore the Most Holy Sacrament. (Alleluia.)}{\textit{Ant.} Adorémus in {\ae}térnum Sanctissimum Sacramentum. (Allelúia.)}
\elcol{\lett{O}{} praise the \textsc{Lord}, all ye nations; * praise him, all ye peoples.}{\lett{L}{audáte} Dóminum, omnes gentes: * laudáte eum, omnes pópuli:}
\elcol{For his merciful kindness is ever more and more toward us; * 
and the truth of the \textsc{Lord} endureth for ever. Praise the \textsc{Lord}.}{Quóniam confirmáta est super nos misericórdia ejus: * et véritas Dómini manet in {\ae}térnum.}
\elcol{Glory be to the Father, and to the Son, *
    and to the Holy Ghost:}{Glória Patri, et Fílio, * et Spirítui Sancto.}
\elcol{As it was in the beginning, is now, and ever shall be, *
    world without end. Amen.}{Sicut erat in princípio, et nunc, et semper, * et in sǽcula sæculórum. Amen.}
\elcol{\textit{Ant.} Let us forever adore the Most Holy Sacrament. (Alleluia.)}{\textit{Ant.} Adorémus in {\ae}térnum Sanctíssimum Sacraméntum. (Allelúia.)}