\section{The Order of Compline}
\fancyhead[RE,LO]{Compline}
\englat{
℣. Sir, ask a blessing.\\
\textit{Blessing.} The Lord Almighty grant us a quiet night and a perfect end.\\
℟. Amen.\\
\par
\noindent
Brethren: Be sober, be vigilant; because your adversary the devil, as a roaring lion, walketh about, seeking whom he may devour: whom resist, steadfast in the faith.\\
But thou, O Lord, have mercy upon us.\\
℟. Thanks be to God.\\
\textit{\scshape ℣.} Our help {\ding{64}} is in the Name of the Lord.\\
\textit{\scshape ℟.} Who hath made heaven and earth.
}{
℣. Iube, domne, benedícere.\\
\textit{Benedictio.} Noctem quiétam et finem perféctum concédat nobis Dóminus omnípotens.\\
℟. Amen.\\
\par
\noindent
Fratres: Sóbrii estóte, et vigiláte: quia adversárius vester diábolus tamquam leo rúgiens círcuit, qu{\ae}rens quem dévoret: cui resístite fortes in fide.\\
Tu autem, Dómine, miserére nobis.\\
℟. Deo grátias.\\
\textit{\scshape ℣.} Adjutórium nóstrum {\ding{64}} in nómine Dómini.\\
\textit{\scshape ℟.} Qui fecit c{\ae}lum et terram.
}
\rub{The Our Father is said in secret.}
\subsection{Confession}
\rub{The Minister and the People then say the following Confession.}
\englat{%
I confess to God Almighty, to blessed Mary ever-Virgin, to blessed Michael the Archangel, to blessed John Baptist, to the holy Apostles Peter and Paul, and to all the Saints that I have sinned exceedingly in thought, word, and deed, by my fault, by my own fault, by my own most grievous fault. Wherefore I beg blessed Mary ever-Virgin, blessed Michael the Archangel, blessed John Baptist, the holy Apostles Peter and Paul, and all the Saints, to pray for me to the Lord our God.\\
\par\noindent
℣. God Almighty have mercy on us, forgive us our sins, and bring us to everlasting life. Amen.\\
℣. The Almighty {\ding{64}} and merciful Lord grant us pardon, absolution, and remission of our sins. Amen.\\
}{%
Confíteor Deo omnipoténti, beátæ Maríæ semper Vírgini, beáto Michaéli Archángelo, beáto Joánni Baptístæ, sanctis Apóstolis Petro et Paulo, et ómnibus Sanctis, quia peccávi nimis, cogitatióne, verbo et ópere: mea culpa, mea culpa, mea máxima culpa. Ídeo precor beátam Maríam semper Vírginem, beátum Michaélem Archángelum, beátum Joánnem Baptístam, sanctos Apóstolos Petrum et Paulum, et omnes Sanctos, oráre pro me ad Dóminum Deum nostrum.\\
\par\noindent
℣. Misereátur nostri omnípotens Deus, et dimíssis peccátis nostris, perdúcat nos ad vitam {\ae}térnam. Amen.\\
℣. Indulgéntiam, {\ding{64}} absolutiónem et remissiónem peccatórum nostrórum tríbuat nobis omnípotens et miséricors Dóminus. Amen.
}
\englat{
\textit{\scshape ℣.} Turn {\ding{61}} us then, O God our Saviour.\\
\textit{\scshape ℟.} And let thine anger cease from us.\\
\textit{\scshape ℣.} O God, {\ding{64}} make speed to save me.\\
\textit{\scshape ℟.} O Lord, make haste to help me.\\
\textit{\scshape ℣.} Glory be to the Father, and to the Son, and to the Holy Ghost;\\
\textit{\scshape ℟.} As it was in the beginning, is now, and ever shall be, world without end. Amen. (Alleluia)
}{
\textit{\scshape ℣.} Convérte nos {\ding{64}} Deus, salutáris noster.\\
\textit{\scshape ℟.} Et avérte iram tuam a nobis.\\
\textit{\scshape ℣.} Deus {\ding{64}} in adjutórium meum inténde.\\
\textit{\scshape ℟.} Dómine, ad adjuvándum me festína.\\
\textit{\scshape ℣.} Glória Patri, et Fílio, et Spirítui Sancto.\\
\textit{\scshape ℟.} Sicut erat in princípio, et nunc, et semper, et in sǽcula s{\ae}culórum. Amen. (Allelúia)
}
\subsection{The Psalms}
\psa{Psalm 4: \textit{Cum invocarem.}}
\lett{H}{ear} me when I call, O God of my righteousness: thou hast set me at liberty when I was in trouble; have mercy upon me, and hearken unto my prayer.
\par
O ye sons of men, how long will ye blaspheme mine honour: and have such pleasure in vanity, and seek after leasing?
\par
Know this also, that the Lord hath chosen to himself the man that is godly: when I call upon the Lord, he will hear me.
\par
Stand in awe, and sin not: commune with your own heart, and in your chamber, and be still.
\par
Offer the sacrifice of righteousness: and put your trust in the Lord.
\par
There be many that say: Who will shew us any good?
\par
Lord, lift thou up: the light of thy countenance upon us.
\par
Thou hast put gladness in my heart: since the time that their corn and wine and oil increased.
\par
I will lay me down in peace, and take my rest: for it is thou, Lord, only, that makest me dwell in safety.
\psa{Psalm 90: \textit{Qui habitat.}}
\lett{W}{hoso} dwelleth under the defence of the most High: shall abide under the shadow of the Almighty.
\par
I will say unto the Lord, Thou art my hope, and my strong hold: my God, in him will I trust.
\par
For he shall deliver thee from the snare of the hunter: and from the noisome pestilence.
\par
He shall defend thee under his wings, and thou shalt be safe under his feathers: his faithfulness and truth shall be thy shield and buckler.
\par
Thou shalt not be afraid for any terror by night: nor for the arrow that flieth by day;
\par
For the pestilence that walketh in darkness: nor for the sickness that destroyeth in the noon-day.
\par
A thousand shall fall beside thee, and ten thousand at thy right hand: but it shall not come nigh thee.
\par
Yea, with thine eyes shalt thou behold: and see the reward of the ungodly.
\par
For thou, Lord, art my hope: thou hast set thine house of defence very high.
\par
There shall no evil happen unto thee: neither shall any plague come nigh thy dwelling.
\par
For he shall give his angels charge over thee: to keep thee in all thy ways.
\par
They shall bear thee in their hands: that thou hurt not thy foot against a stone.
\par
Thou shalt go upon the lion and adder: the young lion and the dragon shalt thou tread under thy feet.
\par
Because he hath set his love upon me, therefore will I deliver him: I will set him up, because he hath known my Name.
\par
He shall call upon me, and I will hear him: yea, I am with him in trouble; I will deliver him, and bring him to honour.
\par
With long life will I satisfy him: and shew him my salvation.
\psa{Psalm 134: \textit{Ecce nunc.}}
\lett{B}{ehold} now, praise the Lord: all ye servants of the Lord;
\par
Ye that by night stand in the house of the Lord: even in the courts of the house of our God.
\par
Lift up your hands in the sanctuary: and praise the Lord.
\par
The Lord that made heaven and earth: give thee blessing out of Sion.
\subsection{Hymn. \textit{Te lucis ante terminum.}}
\englat{%
To thee, before the close of day\\
Creator of the world, we pray\\
That with thy wonted favour thou\\
Wouldst be our Guard and Keeper now.\\
\par
\noindent
From all ill dreams defend our sight\\
From nightly fears and fantasies\\
Tread under foot our ghostly foe\\
That no pollution we may know.\\
\par
\noindent
O Father, that we ask be done\\
Through Jesus Christ, thine only Son\\
Who with the Holy Ghost and thee\\
Shall live and reign eternally. Amen.\\
}{%
Te lucis ante términum,\\
Rerum Creátor, póscimus,\\
Ut pro tua cleméntia\\
Sis pr{\ae}sul et custódia.\\
\par
\noindent
Procul recédant sómnia,\\
Et nóctium phantásmata;\\
Hostémque nóstrum cómprime,\\
Ne polluántur córpora.\\
\par
\noindent
Jesu, tibi sit glória,\\
Qui apparuísti Géntibus,\\
Cum Patre et almo Spíritu,\\
In sempitérna sǽcula. Amen.
}
\subsection{Chapter Verse}
\englat{
Thou, O Lord, art in the midst of us, and we are called by thy Name; leave us not, O Lord our God. \vr{Jer. 14:9}\\
\textit{\scshape ℟.} Thanks be to God.\\
\textit{\scshape ℣.} Keep us, O Lord, as the apple of an eye (alleluia).\\
\textit{\scshape ℟.} Hide us under the shadow of thy wings (alleluia).\\
Lord, have mercy upon us.\\
Christ, have mercy upon us.\\
Lord, have mercy upon us.
}{
Tu autem in nobis es, Dómine, et nomen sanctum tuum invocátum est super nos: ne derelínquas nos, Dómine, Deus noster. \vr{Jer. 14:9}\\
\textit{\scshape ℟.} Deo grátias.\\
\textit{\scshape ℣.} Custódi nos, Dómine, ut pupíllam óculi (allelúia).\\
\textit{\scshape ℟.} Sub umbra alárum tuárum prótege nos (allelúia).\\
Kýrie, eléison.\\
Christe, eléison.\\
Kýrie, eléison.
}
\rub{Our Father is said secretly until:}
\englat{
℣. And lead us not into temptation:
℟. But deliver us from evil. 
}{
℣. Et ne nos indúcas in tentatiónem:\\
℟. Sed líbera nos a malo.
}
\subsection{Preces}
\englat{
\lett{I}{Believe} in God the Father Almighty, Maker of heaven and earth:
    And in Jesus Christ his only Son our Lord: Who was conceived by the Holy Ghost, Born of the Virgin Mary: Suffered under Pontius Pilate, Was crucified, dead, and buried: He descended into hell; The third day he rose again from the dead: He ascended into heaven, And sitteth on the right hand of God the Father Almighty: From thence he shall come to judge the quick and the dead.\par
    I believe in the Holy Ghost: The holy Catholic Church; The Communion of Saints: The Forgiveness of sins:\\
    \par\noindent
℣. The Resurrection {\ding{64}} of the body:\\
℟. And the Life everlasting. Amen.\\
℣. Vouchsafe, O Lord, this night.\\
℟. To keep us without sin.\\
℣. O Lord, hear my prayer.\\
℟. And let my cry come unto thee.\\
℣. The Lord be with you.\\
℟. And with thy spirit.\\
℣. Let us pray.
}
{
\lett{C}{redo} in Deum, Patrem omnipoténtem, Creatórem c{\ae}li et terr{\ae}.
     Et in Jesum Christum, Fílium ejus únicum, Dóminum nostrum: qui concéptus est de Spíritu Sancto, natus ex María Vírgine, passus sub Póntio Piláto, crucifíxus, mórtuus, et sepúltus: descéndit ad ínferos; tértia die resurréxit a mórtuis; ascéndit ad c{\ae}los; sedet ad déxteram Dei Patris omnipoténtis: inde ventúrus est judicáre vivos et mórtuos.\par
     Credo in Spíritum Sanctum, sanctam Ecclésiam cathólicam, Sanctórum communiónem, remissiónem peccatórum,\\
     \par\noindent
℣. carnis {\ding{64}} resurrectiónem:\\
℟. vitam {\ae}térnam. Amen.\\
℣. Dignáre, Dómine, nocta ista.\\
℟. Sine peccáto nos custodíre.\\
℣. Dómine, exáudi oratiónem meam.\\
℟. Et clamor meus ad te véniat.\\
℣. Dóminus vobíscum.\\
℟. Et cum spíritu tuo.\\
℣. Orémus.
 }
\subsection{The Collect}
Visit, we beseech thee, O Lord, this habitation, and drive far from it the snares of the enemy. Let thy holy angels dwell herein to preserve us in peace, and let thy blessing be upon us evermore, through Jesus Christ our Lord.\\
\textit{\scshape ℟.} Amen.
\englat{
\textit{\scshape ℣.} The Lord be with you. (O Lord, hear my prayer.)\\
\textit{\scshape ℟.} And with thy spirit. (And let my cry come unto thee.)\\
\textit{\scshape ℣.} Let us bless the Lord.\\
\textit{\scshape ℟.} Thanks be to God.\\
\textit{\scshape ℣.} The Almighty and merciful Lord, {\ding{64}} the Father, the Son and the Holy Ghost, bless and preserve us.\\
\textit{\scshape ℟.} Amen.
}	
{
	\textit{\scshape ℣.} Dóminus vobíscum. (Dómine, exáudi oratiónem meam.)\\
	\textit{\scshape ℟.} Et cum spiritu tuo. (Et clamor meus ad te véniat.)\\
	\textit{\scshape ℣.} Benedicámus Dómino.\\
	\textit{\scshape ℟.} Deo grátias.\\
	\textit{\scshape ℣.} Benedícat et custódiat nos omnípotens et miséricors Dóminus, {\ding{64}} Pater, et Fílius, et Spíritus Sanctus.\\
	\textit{\scshape ℟.} Amen.
}
\rub{The appropriate Marian Anthem (p. \pageref{mary}) is said or sung.}
\englat{
℣. May the divine help {\ding{64}} remain with us always.\\
℟. And with our absent brethren. Amen.
}{
℣. Divínum auxílium {\ding{64}} máneat semper nobíscum.\\
℟. Et cum frátribus nostris abséntibus. Amen.
}