\section{The Sacrament of Confirmation}
\fancyhead[RE,LO]{The Sacrament of Confirmation}

    Bishop. Our help is in the Name of the Lord;
    Answer. Who hath made heaven and earth.
    Bishop. Blessed be the Name of the Lord;
    Answer. Henceforth, world without end.
    Bishop. Lord, hear our prayer.
    Answer. And let our cry come unto thee.

Bishop. Let us pray.

ALMIGHTY and everliving God, who hast vouchsafed to regenerate these thy servants by Water and the Holy Ghost, and hast given unto them forgiveness of all their sins; Strengthen them, we beseech thee, O Lord, with the Holy Ghost, the Comforter, and daily increase in them thy manifold gifts of grace: the spirit of wisdom and under-standing, the spirit of counsel and ghostly strength, the spirit of knowledge and true godliness; and fill them, O Lord, with the spirit of thy holy fear, now and for ever. Amen. 
 

 
¶ Then all of them in order kneeling before the Bishop, he shall lay his hand upon the head of every one severally, saying,

DEFEND, O Lord, this thy Child with thy heavenly grace; that he may continue thine for ever; and daily increase in thy Holy Spirit more and more, until he come unto thy everlasting kingdom. Amen.

¶ Then shall the Bishop say,

Answer. 
The Lord be with you. 
And with thy spirit.
Bishop. Let us pray.

¶ Then shall the Bishop say the Lord's Prayer, the People kneeling and repeating it with him.

OUR Father, who art in heaven, Hallowed be thy Name. Thy kingdom come. Thy will be done, On earth as it is in heaven. Give us this day our daily bread. And forgive us our trespasses, As we forgive those who trespass against us. And lead us not into temptation, But deliver us from evil. For thine is the kingdom, and the power, and the glory, for ever and ever. Amen.

¶ Then shall the Bishop say,

ALMIGHTY and everliving God, who makest us both to will and to do those things which are good, and accept-able unto thy Divine Majesty; We make our humble supplications unto thee for these thy servants, upon whom, after the example of thy holy Apostles, we have now laid our hands, to certify them, by this sign, of thy favour and gracious goodness towards them. Let thy fatherly hand, we beseech thee, ever be over them; let thy Holy Spirit ever be with them; and so lead them in the knowledge and obedience of thy Word, that in the end they may obtain everlasting life; through our Lord Jesus Christ, who with thee and the same Holy Spirit liveth and reigneth ever, one God, world without end. Amen.

O ALMIGHTY Lord, and everlasting God, vouchsafe, we beseech thee, to direct, sanctify, and govern, both our hearts and bodies, in the ways of thy laws, and in the works of thy commandments; that, through thy most mighty protection, both here and ever, we may be pre served in body and soul; through our Lord and Saviour Jesus Christ. Amen.

¶ Then the Bishop shall bless them, saying thus,

THE Blessing of God Almighty, the Father, the Son, and the Holy Ghost, be upon you, and remain with you for ever. Amen.

¶ The Minister shall not omit earnestly to move the Persons confirmed to come, without delay, to the Lord’s Supper.

¶ And there shall none be admitted to the Holy Communion, until such time as he be confirmed, or be ready and desirous to be confirmed.