\section{Evensong for the Dead}
\fancyhead[RE,LO]{Evensong for the Dead}
\textit{Ant.} I will walk *
\psa{Psalm 114. \textit{Dilexi, quoniam.}}\par
\lett{I}{am} well pleased * that the Lord hath heard the voice of my prayer.
\par
That he hath inclined his ear unto me * therefore will I call upon him as long as I live.
\par
The snares of death compassed me round about * and the pains of hell gat hold upon me.
\par
I shall find trouble and heaviness, and I will call upon the Name of the Lord * O Lord, I beseech thee, deliver my soul.
\par
Gracious is the Lord, and righteous * yea, our God is merciful
\par
The Lord preserveth the simple * I was in misery, and he helped me.
\par
Turn again then unto thy rest, O my soul * for the Lord hath rewarded thee.
\par
And why? thou hast delivered my soul from death * mine eyes from tears, and my feet from falling.
\par
I will walk before the Lord * in the land of the living.\\
\par\noindent
℣. Rest eternal * grant unto them, O Lord.\\
℟. And let light perpetual * shine upon them.\\
\par\noindent
\textit{Ant.} I will walk before the Lord in the land of the living.\\
\textit{Ant.} Woe is me *
\psa{Psalm 119. \textit{Ad Dominum.}}
\lett{W}{hen} I was in trouble I called upon the Lord * and he heard me.
\par
Deliver my soul, O Lord, from lying lips * and from a deceitful tongue.
\par
What reward shall be given or done unto thee, thou false tongue * even mighty and sharp arrows, with hot burning coals.
\par
Woe is me, that I am constrained to dwell with Mesech * and to have my habitation among the tents of Kedar.
\par
My soul hath long dwelt among them * that are enemies unto peace.
\par
I labour for peace, but when I speak unto them thereof * they make them ready to battle.\\
\par\noindent
℣. Rest eternal * grant unto them, O Lord.\\
℟. And let light perpetual * shine upon them.\\
\textit{Ant.} Woe is me that I am constrained to dwell with Mesech.\\
\textit{Ant.} The Lord *
\psa{Psalm 120. \textit{Levavi oculos.}}
\lett{I}{will} lift up mine eyes unto the hills * from whence cometh my help.
\par
My help cometh even from the Lord * who hath made heaven and earth.
\par
He will not suffer thy foot to be moved * and he that keepeth thee will not sleep.
\par
Behold, he that keepeth Israel * shall neither slumber nor sleep.
\par
The Lord himself is thy keeper * the Lord is thy defence upon thy right hand.
\par
So that the sun shall not burn thee by day * neither the moon by night.
\par
The Lord shall preserve thee from all evil * yea, it is even he that shall keep thy soul.
\par
The Lord shall preserve thy going out, and thy coming in * from this time forth for evermore.\\
\par\noindent
℣. Rest eternal * grant unto them, O Lord.\\
℟. And let light perpetual * shine upon them.\\
\textit{Ant.} The Lord shall preserve thee from all evil * yea, it is even he that shall keep thy soul.\\
\textit{Ant.} If thou, Lord, wilt be extreme *
\psa{Psalm 129. \textit{De profundis.}}
\lett{O}{ut} of the deep have I called unto thee, O Lord * Lord, hear my voice.
\par
O let thine ears consider well * the voice of my complaint.
\par
If thou, Lord, wilt be extreme to mark what is done amiss * O Lord, who may abide it?
\par
For there is mercy with thee * therefore shalt thou be feared.
\par
I look for the Lord; my soul doth wait for him * in his word is my trust.
\par
My soul fleeth unto the Lord * before the morning watch, I say, before the morning watch.
\par
O Israel, trust in the Lord, for with the Lord there is mercy * and with him is plenteous redemption.
\par
And he shall redeem Israel * from all his sins.\\
\par\noindent
℣. Rest eternal * grant unto them, O Lord.\\
℟. And let light perpetual * shine upon them.\\
\textit{Ant.} If thou, Lord, wilt be extreme to mark what is done amiss, O Lord, who may abide it?\\
\textit{Ant.} Despise not then, *
\psa{Psalm 137. \textit{Confitebor tibi.}}
\lett{I}{will} give thanks unto thee, O Lord, with my whole heart * even before the gods will I sing praise unto thee.
\par
I will worship toward thy holy temple, and praise thy Name, because of thy loving-kindness and truth * for thou hast magnified thy Name and thy word above all things.
\par
When I called upon thee, thou heardest me * and enduedst my soul with much strength.
\par
All the kings of the earth shall praise thee, O Lord * for they have heard the words of thy mouth.
\par
Yea, they shall sing in the ways of the Lord * that great is the glory of the Lord.
\par
For though the Lord be high, yet hath he respect unto the lowly * as for the proud, he beholdeth them afar off.
\par
Though I walk in the midst of trouble, yet shalt thou refresh me * thou shalt stretch forth thy hand upon the furiousness of mine enemies, and thy right hand shall save me.
\par
The lord shall make good his loving-kindness toward me * yea, thy mercy, O Lord, endureth for ever; despise not then the works of thine own hands.\\
\par\noindent
℣. Rest eternal * grant unto them, O Lord.\\
℟. And let light perpetual * shine upon them.\\
\textit{Ant.} Despise not then, O Lord, the works of thine own hands.\\
\par\noindent
℣. I heard a voice from heaven, saying unto me.\\
℟. Blessed are the dead which die in the Lord.\\
\textit{Ant.} All that the Father *
\subsection{The Magnificat}
\englat{%
\lett{M}{y} soul {\ding{64}} doth magnify the Lord, * and my spirit hath rejoiced in God my Saviour.\\
    For he hath regarded * the lowliness of his handmaiden.\\
    For behold, from henceforth * all generations shall call me blessed.\\
    For he that is mighty hath magnified me; * and holy is his Name.\\
    And his mercy is on them that fear him * throughout all generations.\\
    He hath showed strength with his arm; * he hath scattered the proud in the imagination of their hearts.\\
    He hath put down the mighty from their seat, * and hath exalted the humble and meek.\\
    He hath filled the hungry with good things; * and the rich he hath sent empty away.\\
    He remembering his mercy hath holpen his servant Israel; * as he promised to our forefathers, Abraham and his seed, for ever.\\
℣. Rest eternal * grant unto them, O Lord.\\
℟. And let light perpetual * shine upon them.
}
{%
\lett{M}{agníficat} {\ding{64}} ánima mea Dóminum.\\
Et exsultávit spíritus meus: * in Deo, salutári meo.\\
Quia respéxit humilitátem ancíll{\ae} su{\ae}: * ecce enim ex hoc beátam me dicent omnes generatiónes.\\
Quia fecit mihi magna qui potens est: * et sanctum nomen ejus.\\
Et misericórdia ejus, a progénie in progénies: * timéntibus eum.\\
Fecit poténtiam in brácchio suo: * dispérsit supérbos mente cordis sui.\\
Depósuit poténtes de sede: * et exaltávit húmiles.\\
Esuriéntes implévit bonis: * et dívites dimísit inánes.\\
Suscépit Israël púerum suum: * recordátus misericórdi{\ae} su{\ae}.\\
Sicut locútus est ad patres nostros: * Ábraham, et sémini ejus in s{\ae}cula.\\
℣. Réquiem ætérnam * dona eis, Dómine.\\
℟. Et lux perpétua * lúceat eis.
}

\textit{Ant.} All that the Father giveth me shall come to me; and him that cometh to me, I will in no wise cast out.
\rub{The Our Father is here said, in secret, kneeling.}
℣. And lead us not into temptation.\\
℟. But deliver us from evil.\\
\rub{Psalm 145 is not said on All Souls Day, on All the Departed of the Order, on the day of death or burial, nor at any time when the Office is recited with Double rite.}
\psa{Psalm 145}
\lett{P}{raise} the Lord, O my soul; while I live will I praise the Lord * yea, as long as I have any being, I will sing praises unto my God.
\par
O put not your trust in princes, nor in any child of man * for there is no help in them.
\par
For when the breath of man goeth forth he shall turn again to his earth * and then all his thoughts perish.
\par
Blessed is he that hath the God of Jacob for his help * and whose hope is in the Lord his God;
\par
Who made heaven and earth, the sea, and all that therein is * who keepeth his promise for ever;
\par
Who helpeth them to right that suffer wrong * who feedeth the hungry.
\par
The Lord looseth men out of prison * the Lord giveth sight to the blind.
\par
The Lord helpeth them that are fallen * the Lord careth for the righteous.
\par
The Lord careth for the strangers, he defendeth the fatherless and widow * as for the way of the ungodly, he turneth it upside down.
\par
The Lord thy God, O Sion, shall be King for evermore * and throughout all generations.\\
\par\noindent
℣. Rest eternal * grant unto them, O Lord.\\
℟. And let light perpetual * shine upon them.\\
℣. From the gate of hell.\\
℟. Deliver \textit{his soul}, O Lord.\\
℣. May \textit{he} rest in peace.\\
℟. Amen.\\
℣. O Lord, hear my prayer.\\
℟. And let my cry come unto thee.\\
℣. The Lord be with you.\\
℟. And with thy spirit.\\
℣. Let us pray.
\rub{The appropriate collect is here said.}
℣. Rest eternal * grant unto them, O Lord.\\
℟. And let light perpetual * shine upon them.\\
℣. May they rest in peace.\\
℟. Amen.