\section{The Order of Evensong}
\fancyhead[RE,LO]{Evening Prayer}
\elcol{℣. O Lord, {\ding{61}} open thou our lips.}{℣. Dómine, {\ding{61}} lábia nostra apéries.}
\elcol{℟. And our mouth shall show forth thy praise.}{℟. Et os nostrum annuntiábit laudem tuam.}
\elcol{℣. O God, {\ding{64}} make speed to save us.}{℣. Deus, {\ding{64}} in adiutórium nostrum inténde.}
\elcol{℟. O Lord, make haste to help us.}{℟. Dómine, ad adiuvándum nos festína.}
\elcol{℣. Glory be to the Father, and to the Son, and to the Holy Ghost.}{℣. Glória Patri, et Fílio, * et Spirítui Sancto:}
\elcol{℟. As it was in the beginning, is now, and ever shall be, world without end. Amen.}{℟. Sicut erat in princípio, et nunc, et semper, * et in sǽcula s{\ae}culórum. Amen.}
\elcol{℣. Praise ye the Lord.}{℣. Laudáte Dominum.}
\elcol{℟. The Lord's Name be praised.}{℟. Sit Nomen Domini Benedictum.}
\rub{Then shall be sung or said a portion of the Psalms according to the use of this Book. After each Psalm (or Psalm portion) and the \textit{Magnificat} and the \textit{Nunc Dimittis} shall be prayed the \textit{Gloria Patri.}\\
After which shall be read the First Lesson, according to the Daily Office Lectionary.}
\subsection{The Magnificat}
\elcol{\lett{M}{y} soul {\ding{64}} doth magnify the Lord, * and my spirit hath rejoiced in God my Saviour.\par
    For he hath regarded * the lowliness of his handmaiden.\par
    For behold, from henceforth * all generations shall call me blessed.\par
    For he that is mighty hath magnified me; * and holy is his Name.\par
    And his mercy is on them that fear him * throughout all generations.\par
    He hath showed strength with his arm; * he hath scattered the proud in the imagination of their hearts.\par
    He hath put down the mighty from their seat, * and hath exalted the humble and meek.\par
    He hath filled the hungry with good things; * and the rich he hath sent empty away.\par
    He remembering his mercy hath holpen his servant Israel; * as he promised to our forefathers, Abraham and his seed, for ever.}{\lett{M}{agníficat} {\ding{64}} ánima mea Dóminum.\par
Et exsultávit spíritus meus: * in Deo, salutári meo.\par
Quia respéxit humilitátem ancíll{\ae} su{\ae}: * ecce enim ex hoc beátam me dicent omnes generatiónes.\par
Quia fecit mihi magna qui potens est: * et sanctum nomen ejus.\par
Et misericórdia ejus, a progénie in progénies: * timéntibus eum.\par
Fecit poténtiam in brácchio suo: * dispérsit supérbos mente cordis sui.\par
Depósuit poténtes de sede: * et exaltávit húmiles.\par
Esuriéntes implévit bonis: * et dívites dimísit inánes.\par
Suscépit Israël púerum suum: * recordátus misericórdi{\ae} su{\ae}.\par
Sicut locútus est ad patres nostros: * Ábraham, et sémini ejus in s{\ae}cula.
}
\rub{Then Second Lesson shall then be read followed by the \textit{Nunc Dimittis}.}
\subsection{The Nunc Dimittis}
\elcol{\lett{L}{ord,} {\ding{64}} now lettest thou thy servant depart in peace, * according to thy word.\par
    For mine eyes have seen * thy salvation,\par
    Which thou hast prepared * before the face of all people;\par
    To be a light to lighten the Gentiles, * and to be the glory of thy people Israel.
}{\lett{N}{unc} dimíttis {\ding{64}} servum tuum, Dómine, * secúndum verbum tuum in pace:\par
Quia vidérunt óculi mei * salutáre tuum,\par
Quod parásti * ante fáciem ómnium populórum,\par
Lumen ad revelatiónem géntium, * et glóriam plebis tu{\ae} Isra{\"e}l.}
\subsection{The Apostles' Creed}
\elcol{\lett{I}{Believe} in God the Father Almighty, Maker of heaven and earth:
    And in Jesus Christ his only Son our Lord: Who was conceived by the Holy Ghost, Born of the Virgin Mary: Suffered under Pontius Pilate, Was crucified, dead, and buried: He descended into hell; The third day he rose again from the dead: He ascended into heaven, And sitteth on the right hand of God the Father Almighty: From thence he shall come to judge the quick and the dead.\par
    I believe in the Holy Ghost: The holy Catholic Church; The Communion of Saints: The Forgiveness of sins: The Resurrection of the body: {\ding{64}} And the Life everlasting. Amen.
}{\lett{C}{redo} in Deum, Patrem omnipoténtem, Creatórem c{\ae}li et terr{\ae}.
     Et in Jesum Christum, Fílium ejus únicum, Dóminum nostrum: qui concéptus est de Spíritu Sancto, natus ex María Vírgine, passus sub Póntio Piláto, crucifíxus, mórtuus, et sepúltus: descéndit ad ínferos; tértia die resurréxit a mórtuis; ascéndit ad c{\ae}los; sedet ad déxteram Dei Patris omnipoténtis: inde ventúrus est judicáre vivos et mórtuos.\par
     Credo in Spíritum Sanctum, sanctam Ecclésiam cathólicam, Sanctórum communiónem, remissiónem peccatórum, carnis {\ding{64}} resurrectiónem, vitam {\ae}térnam. Amen.}
\rub{And after that, these Prayers following, the people devoutly kneeling; the Minister first pronouncing (the words in parentheses if he is not in major orders),}
\elcol{℣. The Lord be with you. (O Lord, hear my prayer.)}{℣. Dóminus vobíscum. (Dómine, exáudi oratiónem meam.)}
\elcol{℟. And with thy spirit. (And let my cry come unto thee.)}{℟. Et cum spiritu tuo. (Et clamor meus ad te véniat.)}
\elcol{℣. Let us pray.}{℣. Orémus.}
\elcol{℣. Lord, have mercy upon us.}{℣. Kýrie, eléison.}
\elcol{℟. Christ, have mercy upon us.}{℟. Christe, eléison.}
\elcol{℣. Lord, have mercy upon us.}{℣. Kýrie, eléison.}
\subsection{Our Father}
\rub{All kneeling shall say the Lord's Prayer.}
\elcol{
\lett{O}{ur} Father, who art in heaven, Hallowed be thy Name. Thy kingdom come. Thy will be done on earth, As it is in heaven. Give us this day our daily bread. And forgive us our trespasses, As we forgive those who trespass against us. And lead us not into temptation; But deliver us from evil. Amen.}
{
\lett{P}{ater} noster, qui es in c{\ae}lis, sanctificétur nomen tuum: advéniat regnum tuum: fiat volúntas tua, sicut in c{\ae}lo et in terra. Panem nostrum quotidiánum da nobis hódie: et dimítte nobis débita nostra, sicut et nos dimíttimus debitóribus nostris: et ne nos indúcas in tentatiónem: sed líbera nos a malo. Amen.}
\subsection{The Preces}
\rub{The Minister, standing, shall say:}
\elcol{℣. O Lord, show thy mercy upon us.}{℣. Ostende nobis Domine misericordiam tuam.}
\elcol{℟. And grant us thy salvation.}{℟. Et salutare tuum da nobis.}
\elcol{℣. O Lord, save the State.}{℣. Domine salvam fac Civitatem.}
\elcol{℟. And mercifully hear us when we call upon thee.}{℟. Et exaudi nos cum invocamus te.}
\elcol{℣. Endue thy Ministers with righteousness.}{℣. Sacerdotes tui induantur Justitia.}
\elcol{℟. And make thy chosen people joyful.}{℟. Et sancti tui exultent.}
\elcol{℣. O Lord, save thy people.}{℣. Salvum fac Populum tuum Domine.}
\elcol{℟. And bless thine inheritance.}{℟. Et benedic H{\ae}reditati tu{\ae}.}
\elcol{℣. Give peace in our time, O Lord.}{℣. Da pacem Domine in diebus nostris.}
\elcol{℟. For it is thou, Lord, only, that makest us dwell in safety.}{℟. Quia non est alius qui pugnet pro nobis, nisi tu Deus noster.}
\elcol{℣. O God, make clean our hearts within us.}{℣. Cor mundum crea in nobis, O Deus.}
\elcol{℟. And take not thy Holy Spirit from us.}{℟. Et Spiritum Sanctum tuum ne auferas a nobis.}
\subsection{The Collects}
\rub{Then shall follow the Collect for the Day and then the following collects.}
\subsubsection{A Collect for Peace}
\lett{O}{God,} from whom all holy desires, all good counsels, and all just works do proceed; Give unto thy servants that peace which the world cannot give; that our hearts may be set to obey thy commandments, and also that by thee, we, being defended from the fear of our enemies, may pass our time in rest and quietness; through the merits of Jesus Christ our Saviour. Amen.
\subsubsection{A Collect for Aid against Perils}
\lett{L}{ighten} our darkness, we beseech thee, O Lord; and by thy great mercy defend us from all perils and dangers of this night; for the love of thy only Son, our Saviour, Jesus Christ. Amen.
\subsection{Conclusion}
\elcol{℣. The Lord be with you. (O Lord, hear my prayer.)}{℣. Dóminus vobíscum. (Dómine, exáudi oratiónem meam.)}
\elcol{℟. And with thy spirit. (And let my cry come unto thee.)}{℟. Et cum spiritu tuo. (Et clamor meus ad te véniat.)}
\elcol{℣. Let us bless the Lord. (alleluia, alleluia.)}{℣. Benedicámus Dómino (allelúia, allelúia).}
\elcol{℟. Thanks be to God. (alleluia, alleluia.)}{℟. Deo grátias (allelúia, allelúia).}
\elcol{℣. May the souls {\ding{64}} of the faithful departed, through the mercy of God, rest in peace.}{℣. Fidélium ánim{\ae} {\ding{64}} per misericórdiam Dei requiéscant in pace.}
\elcol{℟. Amen.}{℟. Amen.}
\rub{The appropriate Marian Anthem (p. \pageref{mary}) may be said or sung here.}
\subsection{Third Collects}
\subsubsection{A Prayer for The President of the United States, and all in Civil Authority}
\lett{O}{Lord}, our heavenly Father, the high and mighty Ruler of the universe, who dost from thy throne behold all the dwellers upon earth; Most heartily we beseech thee, with thy favour to behold and bless thy servant THE PRESIDENT OF THE UNITED STATES, and all others in authority; and so replenish them with the grace of thy Holy Spirit, that they may always incline to thy will, and walk in thy way. Endue them plenteously with heavenly gifts; grant them in health and prosperity long to live; and finally, after this life, to attain everlasting joy and felicity; through Jesus Christ our Lord. Amen.
\subsubsection{A Prayer for the Clergy and People}
\lett{A}{lmighty} and everlasting God, from whom cometh every good and perfect gift; Send down upon our Bishops, and other Clergy, and upon the Congregations committed to their charge, the healthful Spirit of thy grace; and, that they may truly please thee, pour upon them the continual dew of thy blessing. Grant this, O Lord, for the honour of our Advocate arid Mediator, Jesus Christ. Amen.
\subsubsection{A Prayer for all Sorts \& Conditions of Men}
\lett{O}{God}, the Creator and Preserver of all mankind, we humbly beseech thee for all sorts and conditions of men; that thou wouldest be pleased to make thy ways known unto them, thy saving health unto all nations. More especially we pray for thy holy Church universal; that it may be so guided and governed by thy good Spirit, that all who profess and call themselves Christians may be led into the way of truth, and hold the faith in unity of spirit, in the bond of peace, and in righteousness of life. Finally, we commend to thy fatherly goodness all those who are any ways afflicted, or distressed, in mind, body, or estate; (especially those for whom our prayers are desired;) that it may please thee to comfort and relieve them, according to. their several necessities; giving them patience under their sufferings, and a happy issue out of all their afflictions. And this we beg for Jesus Christ's sake. Amen.
\subsubsection{A General Thanksgiving}
\rub{The General Thanksgiving may be said by the Congregation with the Minister.}
\lett{A}{lmighty} God, Father of all mercies, we, thine unworthy servants, do give thee most humble and hearty thanks for all thy goodness and loving-kindness to us and to all men; (to those who desire now to offer up their praises and thanksgivings for thy late mercies vouchsafed unto them.) We bless thee for our creation, preservation, and all the blessings of this life; but above all, for thine inestimable love in the redemption of the world by our Lord Jesus Christ; for the means of grace, and for the hope of glory. And, we beseech thee, give us that due sense of all thy mercies, that our hearts may he unfeignedly thankful; and that we show forth thy praise, not only with our lips, but in our lives, by giving up our selves to thy service, and by walking before thee in holiness and righteousness all our days; through Jesus Christ our Lord, to whom, with thee and the Holy Ghost, be all honour and glory, world without end. Amen.
\subsubsection{A Prayer of St. Chrysostom}
\lett{A}{lmighty} God, who hast given us grace at this time with one accord to make our common supplications unto thee; and dost promise that when two or three are gathered together in thy Name thou wilt grant their requests; Fulfil now, O Lord, the desires and petitions of thy servants, as may be most expedient for them; granting us in this world knowledge of thy truth, and in the world to come life everlasting. Amen.
\subsubsection{The Grace}
\lett{T}{he} grace of our Lord Jesus Christ, {\ding{64}} and the love of God, and the fellowship of the Holy Ghost, be with us all evermore. Amen.\par
\begin{center}
	\textsc{Here endeth the Order of Evening Prayer.}
\end{center}