\section{The Fore-Office}
\fancyhead[RE,LO]{The Fore-Office}
\rub{The Minister shall begin the Major Hours by reading of the following Scriptures.}
\begin{multicols}{2}
\seas{Morning}
The \textsc{Lord} is in his holy temple: let all the earth keep silence before him.\vr{Hab. 2:20}
\par
  \vr{Ps. 122:1}I was glad when they said unto me, We will go into the house of the \textsc{Lord}.
    \par
    Let the words of my mouth, and the meditation of my heart, be alway acceptable in thy sight, O \textsc{Lord}, my strength and my redeemer.\vr{Ps. 19:14}
    \par
    O send out thy light and thy truth, that they may lead me, and bring me unto thy holy hill, and to thy dwelling.\vr{Ps. 43:3}
    \par
    Thus saith the high and lofty One that inhabiteth eternity, whose name is Holy; I dwell in the high and holy place, with him also that is of a contrite and humble spirit, to revive the spirit of the humble, and to revive the heart of the contrite ones.\vr{Is. 57:15}
    \seas{Evening}
    Let my prayer be set forth in thy sight as the incense; and let the lifting up of my hands be an evening sacrifice.\vr{Ps. 141:2}
    \par
    The \textsc{Lord} is in his holy temple: let all the earth keep silence before him.\vr{Hab. 2:20}
    \par
    \textsc{Lord}, I have loved the habitation of thy house, and the place where thine honour dwelleth.\vr{Ps.26:8}
    \par
    O worship the \textsc{Lord} in the beauty of holiness; let the whole earth stand in awe of him.\vr{Ps. 96:9}
    \par
    Let the words of my mouth, and the meditation of my heart, be alway acceptable in thy sight, O \textsc{Lord}, my strength and my redeemer.\vr{Ps. 19:14-15}
%    \par
	\seas{Advent}
	Repent ye, for the Kingdom of heaven is at hand.\vr{Matt. 3:2}
	\par
	For, behold, the darkness shall cover the earth, and gross darkness the people: but the Lord shall arise upon thee, and his glory shall be seen upon thee. \vr{Is. 40:2}
	\par
    O Lord, Almighty God of our fathers, whom all men fear, and tremble before thy power; for the majesty of thy glory cannot be borne, and thine angry threatening toward sinners is importable: \vr{Mann. 1,5}
%    \par
	\seas{Christmas}
	Behold, I bring you good tidings of great joy, which shall be to all people. For unto you is born this day in the city of David a Saviour, which is Christ the Lord.\vr{Lk. 2:10-11}
	\par
	Herein was the love of God manifested in us, that God hath sent his only begotten Son into the world, that we might live through him. \vr{1 Jn. 4:9}
	\par
	Thy merciful promise is unmeasurable and unsearchable; for thou art the most high Lord, of great compassion, longsuffering, very merciful, and repentest of the evils of men. \vr{Mann. 6-7}
%	\par
	\seas{Epiphany}
	And the Gentiles shall come to thy light, and kings to the brightness of thy rising.\vr{Is. 60:3}\par
	From the rising of the sun even unto the going down of the same my Name shall be great among the Gentiles; and in every place incense shall be offered unto my Name, and a pure offering: for my Name shall be great among the heathen, saith the \textsc{Lord} of hosts.\vr{Mal. 1:11}
	\par
%    Awake, awake; put on thy strength, O Zion; put on thy beautiful garments, O Jerusalem.\vr{Is. 52:1}
%    \par
    Thou, O Lord, according to thy great goodness hast promised repentance and forgiveness to them that have sinned against thee: and of thine infinite mercies hast appointed repentance unto sinners, that they may be saved. \vr{Mann. 8}
%    \par
	\seas{Lent}
	Rend your heart, and not your garments, and turn unto the \textsc{Lord} your God: for he is gracious and merciful, slow to anger, and of great kindness, and repenteth him of the evil.\vr{Jl. 2:13}\par
    The sacrifices of God are a broken spirit: a broken and a contrite heart, O God, thou wilt not despise.\vr{Psalm 51:17}\par
    I have sinned, O Lord, I have sinned, and I acknowledge mine iniquities: wherefore, I humbly beseech thee, forgive me, O Lord, forgive me, and destroy me not with mine iniquities. \vr{Mann. 12}
%    \par
    \seas{Passiontide}
    Is it nothing to you, all ye that pass by? behold, and see if there be any sorrow like unto my sorrow which is done unto me, wherewith the LORD hath afflicted me.\vr{Lam. 1:12}\par
    Be not angry with me for ever, by reserving evil for me; neither condemn me to the lower parts of the earth. For thou art the God, even the God of them that repent; and in me thou wilt shew all thy goodness: for thou wilt save me, that am unworthy, according to thy great mercy. \vr{Mann. 13}
	\seas{Eastertide}
	He is risen. The Lord is risen indeed.\vr{Mk. 16:6}
	\par
    This is the day which the \textsc{Lord} hath made; we will rejoice and be glad in it.\vr{Psalm 118:24}
    \par
    I will praise thee for ever all the days of my life: for all the powers of the heavens do praise thee, and thine is the glory for ever and ever. Amen. \vr{Mann. 15}
%    \par
	\seas{Ascensiontide}
	Christ is not entered into the holy places made with hands, which are the figures of the true; but into heaven itself, now to appear in the presence of God for us.\vr{Heb. 9:24}
	\par
	Seeing that we have a great High Priest, that is passed into the heavens, Jesus the Son of God, let us come boldly unto the throne of grace, that we may obtain mercy, and find grace to help in time of need.\vr{Heb. 4:14,16}
%	\par
	\seas{Whitsuntide}
	Ye shall receive power, after that the Holy Ghost is come upon you: and ye shall be witnesses unto me both in Jerusalem, and in all Jud{\ae}a, and in Samaria, and unto the uttermost part of the earth.\vr{Acts 1:8}
	\par
    Because ye are sons, God hath sent forth the Spirit of his Son into your hearts, crying, Abba, Father.\vr{Gal. 4:6}
%    \par
	\seas{Trinitytide}
	Holy, holy, holy, Lord God Almighty, which was, and is, and is to come.\vr{Rev. 4:8}
	\par
	Holy, holy, holy, is the LORD of hosts: the whole earth is full of his glory. \vr{Is. 6:3}
	\par
	God is love; and he that abideth in love abideth in God and God in him.\vr{1 Jn. 4:16}
%	\par
%	\seas{The Blessed Sacrament}
%	Whoso eateth my flesh, and drinketh my blood, hath eternal life; and I will raise him up at the last day. \vr{Jn. 6:54}
%	\par
%	The cup of blessing which we bless, is it not the communion of the blood of Christ? The bread which we break, is it not the communion of the body of Christ? \vr{1 Cor. 10:16}
	\seas{Blessed Sacrament}
	I am the living bread which came down from heaven: if any man eat of this bread, he shall live for ever: and the bread that I will give is my flesh, which I will give for the life of the world. \vr{Jn. 6:51}\par
	But when they came to Jesus, and saw that he was dead already, they brake not his legs: but one of the soldiers with a spear pierced his side, and forthwith came there out blood and water.\vr{Jn.19:33-4}\par
%	Take my yoke upon you, and learn of me; for I am meek and lowly in heart: and ye shall find rest unto your souls. \vr{Matt. 11:29}
	For God so loved the world, that he gave his only begotten Son, that whosoever believeth in him should not perish, but have everlasting life. \vr{Jn. 3:16}
	\seas{Transfiguration}
	Arise, shine; for thy light is come, and the glory of the Lord is risen upon thee. \vr{Is. 60:1}
%	\seas{The Holy Name of Jesus}
%	And when eight days were accomplished for the circumcising of the child, his name was called \textsc{Jesus}, which was so named of the angel before he was conceived in the womb. \vr{Lk. 2:21}
%	\seas{Christ the King}
%	I will extol my God, and my soul shall praise the King of Heaven, and shall rejoice in his greatness. \vr{Tob. 13:7}
%	\par
%	For it pleased the Father that in him should all fulness dwell; and, having made peace through the blood of his cross, by him to reconcile all things unto himself; by him, I say, whether they be things in earth, or things in heaven. \vr{Col. 1:19}
%	\seas{The Holy Rood}
%	But God forbid that I should glory, save in the cross of our Lord Jesus Christ, by whom the world is crucified unto me, and I unto the world. \vr{Gal. 6:14}
	\seas{The Blessed Virgin Mary}
	Therefore the Lord himself shall give you a sign; Behold, a virgin shall conceive, and bear a son, and shall call his name Immanuel. \vr{Is. 6:14}
	\par
	Strength and honour are her clothing; and she shall rejoice in time to come. She openeth her mouth with wisdom; and in her tongue is the law of kindness. \vr{Prov. 31:25-6}
%	\seas{The Holy Angels}
%	Bless the Lord, ye his angels, that excel in strength, that do his commandments, hearkening unto the voice of his word. \vr{Ps. 103:20}
%	\par
%	Again, the kingdom of heaven is like unto a net, that was cast into the sea, and gathered of every kind: which, when it was full, they drew to shore, and sat down, and gathered the good into vessels, but cast the bad away. So shall it be at the end of the world: the angels shall come forth, and sever the wicked from among the just. \vr{Matt. 13:47-9}
%	\seas{The Holy Apostles}
%	The righteous shall shine, and run to and fro like sparks among the stubble. They shall judge the nations, and have dominion over the people, and their Lord shall reign for ever. \vr{Wis. 3:7}
%	\seas{The Holy Martyrs}
%	These be they that have put off mortal clothing, and put on the immortal, and have confessed the Name of God: now are they crowned, and receive palms. \vr{2 Esdras 2:45}
%	\par
%	My soul drew near even unto death. They compassed me on every side, and there was no man to help me: I looked for the succour of men, but there was none. Then I thought upon thy mercy, O Lord. My prayer was heard: for thou savedst me from destruction. \vr{Sir. 51:6,8,11-2}
%	\seas{Holy Bishops \& Priests}
%	He blessed him with glory, that he should execute the office of the priesthood, to bless the people in his Name, and to offer sacrifices to the Lord, incense, and a sweet savour. \vr{Sir. 44:15}
%	\seas{Holy Confessors \& Monks}
%	Blessed is the man that is found without blemish: and hath not gone after gold. Who is he? And we will call him blessed: for wonderful things hath he doen among his people. \vr{Sir. 31:8}
%	\seas{Holy Women}
%	Who can find a virtuous woman? For her price is far above rubies. The heart of her husband doth safely trust in her, so that he shall have no need of spoil. \vr{Prov. 31:10-1}
	\seas{Saints}
	But the souls of the righteous are in the hand of God, and there shall no torment touch them.\vr{Wis. 3:1}\par
	Look at the generations of old, and see; did ever any trust in the Lord, and was confounded? \vr{Sir. 2:10}
	\par
	The righteous shall shine, and run to and fro like sparks among the stubble. They shall judge the nations, and have dominion over the people, and their Lord shall reign for ever. \vr{Wis. 3:7}
	\seas{Dedication of a Church}
	And I heard a great voice out of heaven, saying, Behold the tabernacle of God is with men, and he will dwell with them, and they shall be his people, and God himself shall be with them, and be their God.\vr{Rev. 21:3}
	\par
	And the Lord appeared to Solomon by night, and said unto him, I have heard thy prayer, and have chosen this place to myself for an house of sacrifice. For now have I chosen and sanctified this house, that my name may be there for ever: and mine eyes and mine heart shall be there perpetually. \vr{2 Chron. 7:12,16}
	\par
	And the temple of God was opened in heaven, and there was seen in his temple the ark of his testament: and there were lightnings, and voices, and thunderings, and an earthquake, and great hail. And there appeared a great wonder in heaven; a woman clothed with the sun, and the moon under her feet, and upon her head a crown of twelve stars: and she being with child cried, travailing in birth, and pained to be delivered. \vr{Rev. 11:19-12:2}
\end{multicols}
\subsection{The General Confession}
\lett{D}{early} beloved brethren, the Scripture moveth us, in sundry places, to acknowledge and confess our manifold sins and wickedness; and that we should not dissemble nor cloak them before the face of Almighty God our heavenly Father; but confess them with an humble, lowly, penitent, and obedient heart; to the end that we may obtain forgiveness of the same, by his infinite goodness and mercy. And although we ought, at all times, humbly to acknowledge our sins before God; yet ought we chiefly so to do, when we assemble and meet together to render thanks for the great benefits that we have received at his hands, to set forth his most worthy praise, to hear his most holy Word, and to ask those things which are requisite and necessary, as well for the body as the soul. Wherefore I pray and beseech you, as many as are here present, to accompany me with a pure heart, and humble voice, unto the throne of the heavenly grace, saying--
\rub{or}
\lett{L}{et} us humbly confess our sins unto Almighty God.
\rub{To be prayed by all kneeling,}
\lett{A}{lmighty} and most merciful Father; We have erred, and strayed from thy ways like lost sheep. We have followed too much the devices and desires of our own hearts. We have offended against thy holy laws. We have left undone those things which we ought to have done; And we have done those things which we ought not to have done; And there is no health in us. But thou, O Lord, have mercy upon us, miserable offenders. Spare thou those, O God, who confess their faults. Restore thou those who are penitent; According to thy promises declared unto mankind in Christ Jesus our Lord. And grant, O most merciful Father, for his sake; That we may hereafter live a godly, righteous, and sober life, To the glory of thy holy Name. Amen.
\subsection{The Absolution}
\rub{The Priest stands and declares,}
\lett{A}{lmighty} God, the Father of our Lord Jesus Christ, who desireth not the death of a sinner, but rather that he may turn from his wickedness and live, hath given power, and commandment, to his Ministers, to declare and pronounce to his people, being penitent, the Absolution and Remission of their sins. He pardoneth and {\ding{64}} absolveth all those who truly repent, and unfeignedly believe his holy Gospel.
\par
Wherefore let us beseech him to grant us true repentance, and his Holy Spirit, that those things may please him which we do at this present; and that the rest of our life hereafter may be pure and holy; so that at the last we may come to his eternal joy; through Jesus Christ our Lord. Amen.
\rub{When the Minister is not a Priest, he stands and says instead,}
\lett{G}{rant}, we beseech thee, merciful Lord, to thy faithful people pardon and peace, that they may be cleansed from all their sins, and serve thee with a quiet mind; through Jesus Christ our Lord. Amen.
\rub{or}
\lett{G}{od} Almighty have mercy upon us, forgive us our sins, and bring us to everlasting life. Amen.
\lett{T}{he} Almighty and merciful Lord grant unto us pardon, absolution, and remission of our sins. Amen.
\rub{The Minister kneeling, all say,}
\lett{O}{ur} Father, who art in heaven, Hallowed be thy Name. Thy kingdom come. Thy will be done on earth, As it is in heaven. Give us this day our daily bread. And forgive us our trespasses, As we forgive those who trespass against us. And lead us not into temptation; But deliver us from evil: For thine is the kingdom, and the power, and the glory, for ever and ever. Amen.
\par\noindent
\lett{H}{ail} Mary, full of grace; The Lord is with thee; Blessed art thou amongst women, And blessed is the fruit of thy womb, Jesus. Holy Mary, Mother of God, Pray for us sinners, now and at the hour of our death. Amen.