\section{The Liturgy of St. Gregory}
\fancyhead[RE,LO]{The Liturgy of St. Gregory}
\rub{The Introit is then said or sung followed by the Kyrie.}
\subsection{The Kyrie}
\elcoldent{℣. Lord, have mercy upon us.}{℣. Kyrie, eléison.}
\elcoldent{℟. Lord, have mercy upon us.}{℟. Kyrie, eléison.}
\elcoldent{℣. Lord, have mercy upon us.}{℣. Kyrie, eléison.}
\elcoldent{℟. Christ, have mercy upon us.}{℟. Christe, eléison.}
\elcoldent{℣. Christ, have mercy upon us.}{℣. Christe, eléison.}
\elcoldent{℟. Christ, have mercy upon us.}{℟. Christe, eléison.}
\elcoldent{℣. Lord, have mercy upon us.}{℣. Kyrie, eléison.}
\elcoldent{℟. Lord, have mercy upon us.}{℟. Kyrie, eléison.}
\elcoldent{℣. Lord, have mercy upon us.}{℣. Kyrie, eléison.}
\subsection{Gloria in Excelsis}
\elcol{
\lett{G}{lory} be to God on high, and on earth peace, towards men of good will. We praise thee. We bless thee. (He bows his head:) We worship thee. We glorify thee. We give thanks to thee for thy great glory. O Lord God, heavenly king, God the Father almighty. O Lord, the only-begotten Son, Jesus Christ. O Lord God, Lamb of God, Son of the Father. That takest away the sins of the world, have mercy upon us. Thou that takest away the sins of the world, receive our prayer. Thou that sittest at the right hand of the Father, have mercy upon us. For thou only art holy. Thou only art the Lord. Thou only, O Jesus Christ, with the Holy Ghost, art most high in the glory {\ding{64}} of God the Father. Amen.\\
}{
\lett{G}{lória} in excélsis Deo. Et in terra pax homínibus bon{\ae} voluntátis. Laudámus te. Benedícimus te. Adorámus te. Glorificámus te. Grátias ágimus tibi propter magnam glóriam tuam. Dómine Deus, Rex c{\ae}léstis, Deus Pater omnípotens. 
    Dómine Fili unigénite, Jesu Christe. Dómine Deus, Agnus Dei, Fílius Patris. Qui tollis peccáta mundi, miserére nobis. Qui tollis peccáta mundi, súscipe deprecatiónem nostram. Qui sedes ad déxteram Patris, miserére nobis. Quóniam tu solus Sanctus. Tu solus Dóminus. Tu solus Altíssimus, Jesu Christe. Cum Sancto Spíritu {\ding{64}} in glória Dei Patris. Amen.\\
}
\elcol{℣.The Lord be with you.}{℣. Dóminus vobíscum.}
\elcol{℟. And with thy spirit.}{℟. Et cum spiritu tuo.}
\elcol{℣. Let us pray.}{℣. Orémus.}
\subsection{The Collect}
\rub{The Priest prays the collect(s) for the day.}
\subsection{The Epistle}
\rub{The proper epistle for the day shall be read by the Subdeacon.\\At the end of the Epistle the People respond,}
\elcol{℟. Thanks be to God.}{℟. Deo grátias.}
\subsection{The Gospel}
\rub{The Gospel shall be read by the Deacon.\\At the end of the Gospel the People respond,}
\elcol{℟. Praise be to thee, O Christ.}{℟. Laus tibi, Christe.}
\rub{A sermon may be given.}
\subsection{The Nicene Creed}
\elcol{
\lett{I}{Believe} in one God the Father Almighty, Maker of heaven and earth, And of all things visible and invisible:\par
    And in one Lord Jesus Christ, the only-begotten Son of God; Begotten of his Father before all worlds, God of God, Light of Light, Very God of very God; Begotten, not made; Being of one substance with the Father; By whom all things were made: Who for us men and for our salvation came down from heaven, \inrub{Everyone genuflects.} And was incarnate by the Holy Ghost of the Virgin Mary, And was made man: \inrub{Everyone rises.} And was crucified also for us under Pontius Pilate; He suffered and was buried: And the third day he rose again according to the Scriptures: And ascended into heaven, And sitteth on the right hand of the Father: And he shall come again, with glory, to judge both the quick and the dead; Whose kingdom shall have no end.\par
    And I believe in the Holy Ghost, The Lord, and Giver of Life, Who proceedeth from the Father; Who with the Father and the Son together is worshiped and glorified; Who spake by the Prophets: And I believe one, holy, catholic, and apostolic Church: I acknowledge one Baptism for the remission of sins: And I look for the Resurrection of the dead: {\ding{64}} And the Life of the world to come. Amen.}
    {
    \lett{C}{redo} in unum Deum, Patrem omnipoténtem, factórem c{\ae}li et terr{\ae}, visibílium ómnium et invisibílium.\par
    Et in unum Dóminum Jesum Christum, Fílium Dei unigénitum. Et ex Patre natum ante ómnia sǽcula. Deum de Deo, lumen de lúmine, Deum verum de Deo vero. Génitum, non factum, consubstantiálem Patri: per quem ómnia facta sunt. Qui propter nos hómines et propter nostram salútem descéndit de c{\ae}lis. \inrub{Everyone genuflects.} Et incarnátus est de Spíritu Sancto ex María Vírgine: Et homo factus est. \inrub{Everyone rises.} Crucifíxus étiam pro nobis: sub Póntio Piláto passus, et sepúltus est. Et resurréxit tértia die, secúndum Scriptúras. Et ascéndit in c{\ae}lum: sedet ad déxteram Patris. Et íterum ventúrus est cum glória judicáre vivos et mórtuos: cujus regni non erit finis.\par
    Et in Spíritum Sanctum, Dóminum et vivificántem: qui ex Patre procédit. Qui cum Patre et Fílio simul adorátur et conglorificátur: qui locútus est per Prophétas. Et unam sanctam cathólicam et apostólicam Ecclésiam. Confíteor unum baptísma in remissiónem peccatórum. Et exspécto resurrectiónem mortuórum. {\ding{64}} Et vitam ventúri s{\ae}culi. Amen.
    }
\subsection{The Offertory}
\rub{The Priest here offers the bread and the wine and the secret prayers.}
\elcol{℣. Pray brethren, that this my sacrifice and yours may be acceptable unto God the Father Almighty.}{℣. Oráte, fratres: ut meum ac vestrum sacrifícium acceptábile fiat apud Deum Patrem omnipoténtem.}
\elcol{℟. May the Lord receive this sacrifice at thy hands, to the praise and glory of his name, both to our benefit, and that of all his holy Church.}{℟. Suscípiat Dóminus sacrifícium de mánibus tuis ad laudem et glóriam nominis sui, ad utilitátem quoque nostram, totiúsque Ecclési{\ae} su{\ae} sanct{\ae}.}
\subsection{The Sursum Corda}
\elcol{℣. The Lord be with you.}{℣. Dóminus vobíscum.}
\elcol{℟. And with thy spirit.}{℟. Et cum spiritu tuo.}
\elcol{℣. Lift up your hearts.}{℣. Sursum corda.}
\elcol{℟. We lift them up unto the Lord.}{℟. Habémus ad Dóminum.}
\elcol{℣. Let us give thanks unto our Lord God.}{℣. Grátias agámus Dómino, Deo nostro.}
\elcol{℟. It is meet and right so to do.}{℟. Dignum et justum est.}
\rub{The proper preface is then said (p. \pageref{prefaces}).}
\subsection{The Sanctus}
\elcol{
\lett{H}{oly, Holy, Holy,} Lord God of Sabaoth! Heaven and earth are full of thy glory! Hosanna in the highest! Blessed is he that cometh in the Name of the Lord! Hosanna in the highest!
}
{
\lett{S}{anctus, Sanctus, Sanctus,} Dóminus, Deus Sábaoth. Pleni sunt c{\ae}li et terra glória tua. Hosánna in excélsis. Benedíctus, qui venit in nómine Dómini. Hosánna in excélsis.
}
\subsection{The Canon of the Mass}
\elcol{\lett{T}{herefore}, most merciful Father, we humbly pray and beseech thee through Jesus Christ thy Son our Lord, that thou accept and bless these {\ding{64}} gifts, these {\ding{64}} offerings, these {\ding{64}} holy spotless sacrifices, which we offer thee in the first place for thy holy Catholic Church, that thou wouldst vouchsafe to keep her in peace, to guard, unite, and govern her throughout the whole world: together with thy servant \textit{N.} our Metropolitan, and \textit{N.}, our Bishop, and all the Orthodox who uphold the Catholic and Apostolic Faith.
}{\lett{T}{e} ígitur, clementíssime Pater, per Jesum Christum, Fílium tuum, Dóminum nostrum, súpplices rogámus, ac pétimus, uti accépta hábeas et benedícas, h{\ae}c {\ding{64}} dona, h{\ae}c {\ding{64}} múnera, h{\ae}c {\ding{64}} sancta sacrifícia illibáta, in primis, qu{\ae} tibi offérimus pro Ecclésia tua sancta cathólica: quam pacificáre, custodíre, adunáre et régere dignéris toto orbe terrárum: una cum fámulo tuo Metropolitáno nostro \textit{N.} et Antístite nostro \textit{N.} et ómnibus orthodóxis, atque cathólic{\ae} et apostólic{\ae} fídei cultóribus.
}
\elcol{\lett{R}{emember}, O Lord, thy servants and handmaids \textit{N.} and \textit{N.} and all here present whose faith and devotion are known and manifest unto thee: for whom we offer, or who themselves offer to thee, this sacrifice of praise, for themselves and those belonging to them: for the redemption of their souls, for the hope of their health and welfare, and who pay their vows to thee, the eternal, living, and true God.
}{\lett{M}{eménto}, Dómine, famulórum famularúmque tuarum \textit{N.} et \textit{N.} et ómnium circumstántium, quorum tibi fides cógnita est et nota devótio, pro quibus tibi offérimus: vel qui tibi ófferunt hoc sacrifícium laudis, pro se suísque ómnibus: pro redemptióne animárum suárum, pro spe salútis et incolumitátis su{\ae}: tibíque reddunt vota sua {\ae}térno Deo, vivo et vero.
}
\rub{Here specific intentions may be made.}
\elcol{
\lett{I}{n} communion with, and venerating first, the memory of the glorious and ever-Virgin Mary, Mother of our Lord and God Jesus Christ. And also of thy blessed Apostles and Martyrs Peter and Paul, Andrew, James, John, Thomas, James, Philip, Bartholomew, Matthew, Simon and Jude: Linus, Cletus, Clement, Xystus, Cornelius, Cyprian, Laurence, Chrysogonus, John and Paul, Cosmas and Damian and of all thy saints: through whose merits and prayers grant that in all things we may be guarded by the help of thy protection. Through the same Christ our Lord. Amen.
}{
\lett{C}{ommunicántes}, et memóriam venerántes, in primis gloriós{\ae} semper Vírginis Marí{\ae}, Genetrícis Dei et Dómini nostri Jesu Christi: sed et beatórum Apostolórum ac Mártyrum tuórum, Petri et Pauli, André{\ae}, Jacóbi, Joánnis, Thom{\ae}, Jacóbi, Philíppi, Bartholomǽi, Matthǽi, Simónis et Thaddǽi: Lini, Cleti, Cleméntis, Xysti, Cornélii, Cypriáni, Lauréntii, Chrysógoni, Joánnis et Pauli, Cosm{\ae} et Damiáni: et ómnium Sanctórum tuórum; quorum méritis precibúsque concédas, ut in ómnibus protectiónis tu{\ae} muniámur auxílio. Per eúndem Christum, Dóminum nostrum. Amen.
}
\subsection{Hanc Igitur}
\elcol{\lett{W}{e} therefore pray thee, O Lord, mercifully to accept this offering of our service and that of all thy family: order our days in thy peace, and bid us to be delivered from eternal damnation, and to be numbered in the flock of thine elect. Through Christ our Lord. Amen.}{\lett{H}{anc} ígitur oblatiónem servitútis nostr{\ae}, sed et cunct{\ae} famíli{\ae} tu{\ae}, quǽsumus, Dómine, ut placátus accípias: diésque nostros in tua pace dispónas, atque ab {\ae}térna damnatióne nos éripi, et in electórum tuórum júbeas grege numerári. Per Christum, Dóminum nostrum. Amen.}
\elcol{\lett{W}{hich} offering, we beseech thee, O God, to ble {\ding{64}}  ss, ap {\ding{64}} prove, ra {\ding{64}} tify, make reasonable and acceptable in every way, that it may become for us the Bo {\ding{64}} dy and Bl {\ding{64}} ood of thy most beloved Son, our Lord Jesus Christ.
}{\lett{Q}{uam} oblatiónem tu, Deus, in ómnibus, quǽsumus, bene {\ding{64}} díctam, adscríp {\ding{64}} tam, ra {\ding{64}} tam, rationábilem, acceptabilémque fácere dignéris: ut nobis Cor {\ding{64}} pus, et San {\ding{64}} guis fiat dilectíssimi Fílii tui, Dómini nostri Jesu Christi. 
}
\subsection{Consecration}
\elcol{\lett{W}{ho}, the day before he suffered, took bread into his holy and venerable hands and, with his eyes lifted up to heaven unto thee, God his almighty Father, giving thanks unto thee, he bles {\ding{64}} sed, brake and gave it to his disciples, saying: Take and eat, ye all, of this,
}{\lett{Q}{ui} prídie quam paterétur, accépit panem in sanctas ac venerábiles manus suas, elevátis óculis in c{\ae}lum ad te Deum, Patrem suum omnipoténtem, tibi grátias agens, bene {\ding{64}} díxit, fregit, dedítque discípulis suis, dicens: Accípite, et manducáte ex hoc omnes.
}
\elcol{
\begin{center}
	{\large{For this is my Body.}}
\end{center}
}{
\begin{center}
	{\large{Hoc est enim Corpus Meum.}}
\end{center}
}
\elcoldent{In like manner after he had supped, taking also this excellent chalice into his holy and venerable hands, again giving thanks unto thee, he ble {\ding{64}} ssed it, and gave it to his disciples saying: Take and drink, ye all, of this,
}{Símili modo postquam cenátum est, accípiens et hunc pr{\ae}clárum Cálicem in sanctas ac venerábiles manus suas: item tibi grátias agens, bene {\ding{64}} díxit, dedítque discípulis suis, dicens: Accípite, et bíbite ex eo omnes.
}
\elcol{
\begin{center}
{\large{For this is the Cup of my Blood of the New and Eternal Testament:}}
\end{center}
}{
\begin{center}
{\large{Hic est enim Calix Sanguinis Mei, Novi et {\AE}terni Testamenti:}}
\end{center}
}
\elcol{
{
\large
\centerline{The Mystery of Faith:}
}
}
{
{
\large
\centerline{Mysterium Fidei:}
}
}
\elcol{
\begin{center}
	{\large{which shall be shed for you and for many unto the remission of sins.}}
\end{center}
}{
\begin{center}
	{\large{qui pro vobis et pro multis effundetur in remissionem peccatorum.}}
\end{center}
}
\elcoldent{As oft as ye shall do these things, ye shall do them in remembrance of me.}{H{\ae}c quotiescúmque fecéritis, in mei memóriam faciétis.}
\elcol{\lett{W}{herefore}, O Lord, we thy servants, as also thy holy people, calling to mind the blessed Passion of the same Christ, thy Son our Lord, his Resurrection from the dead and glorious Ascension into heaven, offer unto thy most excellent majesty of thy gifts bestowed upon us a pure {\ding{64}} host, a holy {\ding{64}} host, a spotless {\ding{64}} host, the holy {\ding{64}} bread of eternal life, and the chalice {\ding{64}} of everlasting salvation.
}{\lett{U}{nde} et mémores, Dómine, nos servi tui, sed et plebs tua sancta, ejúsdem Christi Fílii tui, Dómini nostri, tam beát{\ae} passiónis, nec non et ab ínferis resurrectiónis, sed et in c{\ae}los gloriós{\ae} ascensiónis: offérimus pr{\ae}clár{\ae} majestáti tu{\ae} de tuis donis ac datis, hóstiam {\ding{64}} puram, hóstiam {\ding{64}} sanctam, hóstiam {\ding{64}} immaculátam, Panem {\ding{64}} sanctum vit{\ae} {\ae}térn{\ae}, et Cálicem {\ding{64}} salútis perpétu{\ae}.}
\elcol{\lett{U}{pon} which vouchsafe to look with a favourable and serene countenance, and to accept them as thou wert graciously pleased to accept the gifts of thy just servant Abel, and the sacrifice of our patriarch Abraham, and that which thy high priest Melchisedech offered unto thee, a holy sacrifice, a spotless victim.
}{\lett{S}{upra} qu{\ae} propítio ac seréno vultu respícere dignéris: et accépta habére, sicúti accépta habére dignátus es múnera púeri tui justi Abel, et sacrifícium Patriárch{\ae} nostri Abrah{\ae}: et quod tibi óbtulit summus sacérdos tuus Melchísedech, sanctum sacrifícium, immaculátam hóstiam.}
\subsection{Epikleses}%
\elcol{\lett{A}{nd} we beseech thee, O Lord, to send down thy Holy Spirit upon us and upon these offerings: that he would make this Bread the precious Bo {\ding{64}} dy of thy Christ, and that which is in this Cup the precious Bl {\ding{64}} ood of thy Son, our Lord Jesus Christ, changing them by thy Holy Spirit.
}{\selectlanguage{greek}%
\lett{Π}{αρακαλοῦμέν} σε καὶ δεόμεθα καὶ ἱκετεύομεν· κατάπεμψον τὸ Πνεῦμά σου τὸ Ἅγιον ἐφ’ ἡμᾶς, καὶ ἐπὶ τὰ προκείμενα δῶρα ταῦτα. Καὶ ποίησον τὸν μὲν Ἄρτον τοῦτον τίμιον σῶμα τοῦ Χριστοῦ σου. Τὸ δὲ ἐν τῷ Ποτηρίῳ τούτῳ, τίμιον αἷμα τοῦ Χριστοῦ σου. Μεταβαλὼν τῷ Πνεύματί σου τῷ Ἁγίῳ.
}
\elcol{\lett{W}{e} humbly beseech thee, almighty God, to command that these things be borne by the hands of thy holy angel to thine altar on high, into the presence of thy divine majesty, that so many of us as shall partake at this altar of the most sacred Body {\ding{64}} and Blood {\ding{64}} of thy Son, may be filled with all heavenly benediction. Through the same Christ our Lord. Amen.
}{\lett{S}{úpplices} te rogámus, omnípotens Deus: jube h{\ae}c perférri per manus sancti Angeli tui in sublíme altáre tuum, in conspéctu divín{\ae} majestátis tu{\ae}: ut, quotquot ex hac altáris participatióne sacrosánctum Fílii tui Cor {\ding{64}} pus, et Sán {\ding{64}} guinem sumpsérimus, omni benedictióne c{\ae}lésti et grátia repleámur. Per eúndem Christum, Dóminum nostrum. Amen.
}
\subsection{Commemorations}
\elcol{\lett{B}{e} mindful also, O Lord, of thy servants who are gone before us with the sign of faith, and who rest in the sleep of peace. \inrub{Here the departed are commemorated.} To them, O Lord, and to all who rest in Christ, grant we pray thee a place of refreshment, light and peace, through the same Christ our Lord. Amen.
}{\lett{M}{eménto} étiam, Dómine, famulórum famularúmque tuárum \textit{N.} et \textit{N.}, qui nos pr{\ae}cessérunt cum signo fídei, et dórmiunt in somno pacis. Ipsis, Dómine, et ómnibus in Christo quiescéntibus locum refrigérii, lucis, et pacis, ut indúlgeas, deprecámur. Per eúndem Christum, Dóminum nostrum. Amen.
}
\elcol{\lett{T}{o} us sinners also, thy servants, confiding in the multitude of thy mercies, grant some lot and partnership with thy holy apostles and martyrs John, Stephen, Matthias, Barnabas, Ignatius, Alexander, Marcellinus, Peter, Felicitas, Perpetua, Agatha, Lucia, Agnes, Cecilia, Anastasia and with all thy saints, into whose company we pray thee of thy mercy to admit us, not weighing our merits, but pardoning our offences. Through Christ our Lord, by whom O Lord, thou dost ever create, sanctify, {\ding{64}} quicken, {\ding{64}} bless {\ding{64}} and bestow upon us all these good things.
}{\lett{N}{obis} quoque peccatóribus fámulis tuis, de multitúdine miseratiónum tuárum sperántibus, partem áliquam et societátem donáre dignéris, cum tuis sanctis Apóstolis et Martýribus: cum Joánne, Stéphano, Matthía, Bárnaba, Ignátio, Alexándro, Marcellíno, Petro, Felicitáte, Perpétua, Agatha, Lúcia, Agnéte, C{\ae}cília, Anastásia, et ómnibus Sanctis tuis: intra quorum nos consórtium, non {\ae}stimátor mériti, sed véni{\ae}, quǽsumus, largítor admítte. Per Christum, Dóminum nostrum. Per quem h{\ae}c ómnia, Dómine, semper bona creas, sanctí {\ding{64}} ficas, viví {\ding{64}} ficas, bene {\ding{64}} dícis et pr{\ae}stas nobis.
}
\elcol{\lett{F}{or} by {\ding{64}} him, and with {\ding{64}} him, and in {\ding{64}} him is to thee, God the Father {\ding{64}} Almighty, in the unity of the Holy {\ding{64}} Spirit, all honour and glory,\\
throughout all ages of ages.\\
℟. Amen.
}{\lett{P}{er} ip {\ding{64}} sum, et cum ip {\ding{64}} so, et in ip {\ding{64}} so, est tibi Deo Patri {\ding{64}} omnipoténti, in unitáte Spíritus {\ding{64}} Sancti, omnis honor, et glória.\\
Per ómnia sǽcula s{\ae}culórum.\\
\textit{\scshape ℟.} Amen.
}
\subsection{Lord's Prayer}
\elcoldent{Let us pray.\\
Instructed by saving precepts and following thy divine institution, we presume to say:}{Orémus.\\
Pr{\ae}céptis salutáribus móniti, et divína institutióne formáti audémus dícere:}
\elcol{\lett{O}{ur} Father, who art in heaven, hallowed be thy Name, thy kingdom come, thy will be done, on earth as it is in heaven. Give us this day our daily bread. And forgive us our trespasses, as we forgive those who trespass against us. And lead us not into temptation,
}{\lett{P}{ater} noster, qui es in c{\ae}lis. Sanctificétur nomen tuum. Advéniat regnum tuum. Fiat volúntas tua, sicut in c{\ae}lo et in terra. Panem nostrum quotidiánum da nobis hódie. Et dimítte nobis débita nostra, sicut et nos dimíttimus debitóribus nostris. Et ne nos indúcas in tentatiónem:}
\elcol{℟. but deliver us from evil.}{℟. Sed líbera nos a malo.}
\elcol{℣. Amen.}{℣. Amen.}
\elcoldent{Deliver us, we beseech thee, O Lord, from all evils, past, present, and to come; and at the intercession of the blessed and glorious Mary, Ever-Virgin Mother of God, of thy blessed Apostles Peter and Paul, Andrew, and all thy Saints, graciously give peace in our time, that aided by the help of thy loving kindness, we may both be ever free from sin and secure from all disquietude. Through the same our Lord Jesus Christ, who liveth and reigneth with thee in the unity of the Holy Spirit, ever one God.
}{Líbera nos, quǽsumus, Dómine, ab ómnibus malis, pr{\ae}téritis, pr{\ae}séntibus et futúris: et intercedénte beáta et gloriósa semper Vírgine Dei Genetríce María, cum beátis Apóstolis tuis Petro et Paulo, atque Andréa, et ómnibus Sanctis, da propítius pacem in diébus nostris: ut, ope misericórdi{\ae} tu{\ae} adjúti, et a peccáto simus semper líberi et ab omni perturbatióne secúri. Per eúndem Dóminum nostrum Jesum Christum, Fílium tuum.  Qui tecum vivit et regnat in unitáte Spíritus Sancti Deus.}
\elcol{℣. Throughout all ages, world without end.}{℣. Per ómnia sǽcula s{\ae}culórum.}
\elcol{℟. Amen.}{℟. Amen.}
\elcol{℣. The peace of the Lord be always with you.}{℣. Pax Dómini sit semper vobíscum.}
\elcol{℟. And with thy spirit.}{℟. Et cum spíritu tuo.}
\subsection{Agnus Dei}
\elcol{\lett{O}{Lamb} of God, that takest away the sins of the world, have mercy upon us.}{\lett{A}{gnus} Dei, qui tollis peccáta mundi: miserére nobis.}
\elcol{O Lamb of God, that takest away the sins of the world, have mercy upon us.}{Agnus Dei, qui tollis peccáta mundi: miserére nobis.}
\elcol{O Lamb of God, that takest away the sins of the world, grant us thy peace.}{Agnus Dei, qui tollis peccáta mundi: dona nobis pacem.}
\rub{In Masses for the Dead, in place of ``have mercy upon us'' is said ``grant them rest'', and in place of ``grant us thy peace'', is said: ``grant them rest eternal.''\\The Priest first receives the Holy Communion himself while saying secret prayers. The Acolyte says the Confiteor,}
\subsection{The Second Confiteor}
\elcol{℟. I confess to God Almighty, to Blessed Mary Ever-Virgin, to blessed Michael the Archangel, to blessed John Baptist, to the holy Apostles Peter and Paul, to all the Saints, and to thee, father, that I have sinned exceedingly in thought, word and deed, by my fault, by my own fault, by my own most grievous fault. Wherefore I beg blessed Mary Ever-Virgin, blessed Michael the Archangel, blessed John the Baptist, the holy Apostles Peter and Paul, all the Saints, and thee, father, to pray for me to the Lord our God.}{℟. Confíteor Deo omnipoténti, beát{\ae} Marí{\ae} semper Vírgini, beáto Michaéli Archángelo, beáto Joánni Baptíst{\ae}, sanctis Apóstolis Petro et Paulo, ómnibus Sanctis, et tibi, pater: quia peccávi nimis cogitatióne, verbo et ópere: mea culpa, mea culpa, mea máxima culpa. Ideo precor beátam Maríam semper Vírginem, beátum Michaélem Archángelum, beátum Joánnem Baptístam, sanctos Apóstolos Petrum et Paulum, omnes Sanctos, et te, pater, oráre pro me ad Dóminum, Deum nostrum.}
\elcol{℣. God Almighty have mercy upon you, forgive you your sins, and bring you to everlasting life.}{℣. Misereátur vestri omnípotens Deus, et, dimíssis peccátis vestris, perdúcat vos ad vitam {\ae}térnam.}
\elcol{℟. Amen.}{℟. Amen.}
\elcol{℣. The Almighty and merciful Lord grant you pardon, {\ding{64}} absolution, and remission of your sins.}{℣. Indulgéntiam, {\ding{64}} absolutiónem et remissiónem peccatórum vestrórum tríbuat vobis omnípotens et miséricors Dóminus.}
\elcol{℟. Amen.}{℟. Amen.}
\rub{He then turns to the People, holding the Blessed Sacrament, saying,}
\elcol{℣. Behold the Lamb of God. Behold Him who taketh away the sins of the world.}{℣. Ecce Agnus Dei, ecce, qui tollit peccáta mundi.}
\elcol{℟. Lord, I am not worthy that thou shouldst come under my roof, but speak the word only, and my soul shall be healed.}{℟. Dómine, non sum dignus, ut intres sub tectum meum, sed tantum dic verbo, et sanábitur ánima mea.}
\elcol{℟. Lord, I am not worthy that thou shouldst come under my roof, but speak the word only, and my soul shall be healed.}{℟. Dómine, non sum dignus, ut intres sub tectum meum, sed tantum dic verbo, et sanábitur ánima mea.}
\elcol{℟. Lord, I am not worthy that thou shouldst come under my roof, but speak the word only, and my soul shall be healed.}{℟. Dómine, non sum dignus, ut intres sub tectum meum, sed tantum dic verbo, et sanábitur ánima mea.}
\rub{The Priest shall then say the Communion Antiphon. A hymn may be sung.}
\subsection{The Dismissal}
\rub{After the People have communicated and the Priest cleanses the vessels, he says,}
\elcol{℣. The Lord be with you.}{℣. Dóminus vobíscum.}
\elcol{℟. And with thy spirit.}{℟. Et cum spíritu tuo.}
\elcol{℣. Let us pray.\\}{℣. Orémus.\\}

\rub{The Priest then prays the Postcommunion Prayers.}
\elcol{℣. The Lord be with you.}{℣. Dóminus vobíscum.}
\elcol{℟. And with thy spirit.\\}{℟. Et cum spíritu tuo.\\}

	\rub{On days when the \textit{Gloria in exclesis} is said,}
	\elcol{℣. Depart in peace.}{℣. Ite, Missa est.}
	\elcol{℟. Thanks be to God.}{℟. Deo grátias.}
	\rub{On days when the \textit{Gloria in exclesis} is not said,}
	\elcol{℣. Let us bless the Lord.}{℣. Benedicámus Dómino.}
	\elcol{℟. Thanks be to God.}{℟. Deo grátias.}
	\rub{In Requiem Masses,}
	\elcol{℣. May they rest in peace.}{℣. Requiéscant in pace.}
	\elcol{℟. Amen.\\}{℟. Amen.\\}
	
	\rub{Then, the people kneeling, the Priest (the Bishop if he be present) shall let them depart with this Blessing, unless it be a Requiem Mass.}
\elcol{\lett{T}{he} blessing of God Almighty, the Father, {\ding{64}} the Son, and the Holy Spirit descend upon you, and remain with you always.}{\lett{B}{enedícat} vos omnípotens Deus, Pater, {\ding{64}} et Fílius, et Spíritus Sanctus.}
	\elcol{℟. Amen.}{℟. Amen.}
\subsection{The Last Gospel}
\elcol{℣. The Lord be with you.}{℣. Dóminus vobíscum.}
\elcol{℟. And with thy spirit.}{℟. Et cum spíritu tuo.}
\elcol{℣. The beginning of the Holy Gospel according to John.}{℣. Inítium sancti Evangélii secúndum Joánnem}
\elcol{℟. Glory be to thee, O Lord.}{℟. Glória tibi, Dómine.}
\elcol{\lett{I}{n} the beginning was the Word, and the Word was with God, and the Word was God. The same was in the beginning with God. All things were made by him; and without him was not any thing made that was made. In him was life; and the life was the light of men. And the light shineth in darkness; and the darkness comprehended it not. There was a man sent from God, whose name was John. The same came for a witness, to bear witness of the Light, that all men through him might believe. He was not that Light, but was sent to bear witness of that Light. That was the true Light, which lighteth every man that cometh into the world. He was in the world, and the world was made by him, and the world knew him not. He came unto his own, and his own received him not. But as many as received him, to them gave he power to become the sons of God, even to them that believe on his name: Which were born, not of blood, nor of the will of the flesh, nor of the will of man, but of God. \inrub{All genuflect.} And the Word was made flesh, and dwelt among us, \inrub{All rise.} (and we beheld his glory, the glory as of the only begotten of the Father,) full of grace and truth.}{\lett{I}{n} princípio erat Verbum, et Verbum erat apud Deum, et Deus erat Verbum. Hoc erat in princípio apud Deum. Omnia per ipsum facta sunt: et sine ipso factum est nihil, quod factum est: in ipso vita erat, et vita erat lux hóminum: et lux in ténebris lucet, et ténebr{\ae} eam non comprehendérunt.
Fuit homo missus a Deo, cui nomen erat Joánnes. Hic venit in testimónium, ut testimónium perhibéret de lúmine, ut omnes créderent per illum. Non erat ille lux, sed ut testimónium perhibéret de lúmine.
Erat lux vera, qu{\ae} illúminat omnem hóminem veniéntem in hunc mundum. In mundo erat, et mundus per ipsum factus est, et mundus eum non cognóvit. In própria venit, et sui eum non recepérunt. Quotquot autem recepérunt eum, dedit eis potestátem fílios Dei fíeri, his, qui credunt in nómine ejus: qui non ex sanguínibus, neque ex voluntáte carnis, neque ex voluntáte viri, sed ex Deo nati sunt. \inrub{Genuflectit dicens:} Et Verbum caro factum est, \inrub{Et surgens prosequitur:} et habitávit in nobis: et vídimus glóriam ejus, glóriam quasi Unigéniti a Patre, plenum gráti{\ae} et veritátis. 
}
\elcol{℟. Thanks be to God.}{℟. Deo grátias.}