\section{Marian Anthems}\label{mary}
\subsection{Alma Redemptoris Mater}
\rub{From Advent through Purification}
\englat{
\lett{G}{racious} Mother of our Redeemer, for ever abiding Heaven’s gateway, and star of ocean, O succour the people, who, though falling, strive to rise again. Thou Maiden who barest thy holy Creator, to the wonder of all nature; ever Virgin, after, as before thou receivedst that ``Ave'' from the mouth of Gabriel; have compassion on us sinners.
\begin{center}
	\textit{In Advent}
\end{center}
\par\noindent
	\textit{\scshape ℣.} The Angel of the Lord announced unto Mary.\\
	\textit{\scshape ℟.} And she conceived by the Holy Ghost.\\
	\textit{\scshape ℣.} Let us pray.\\
We beseech thee, O Lord, pour thy grace into our hearts: that, as we have known the incarnation of thy Son Jesus Christ by the message of an angel, so by his cross Cross and Passion we may be brought unto the glory of his Resurrection. Through the same Christ our Lord. Amen.
\begin{center}
	\textit{From Christmas}
\end{center}
\par\noindent
	\textit{\scshape ℣.} After Childbearing, O Virgin, thou didst remain inviolate.\\
	\textit{\scshape ℟.} Intercede for us, O Mother of God.\\
	\textit{\scshape ℣.} Let us pray.\\
O God, who by the fruitful virginity of Blessed Mary hast bestowed upon mankind the reward of eternal salvation: Grant, we beseech thee, that we may know the help of her intercession through whom we have been accounted worthy to receive the Author of our life, Jesus Christ thy Son our Lord. Amen.
}{
\lett{A}{lma} Redemptóris Mater, qu{\ae} pérvia c{\ae}li porta manes, 
et stella maris, succúrre cadénti,
Súrgere qui curat, pópulo: tu qu{\ae} genuísti,
Natúra miránte, tuum sanctum Genitórem,
Virgo prius ac postérius, Gabriélis ab ore
Sumens illud Ave, peccatórum miserére.
\begin{center}
	\textit{In Adventu}
\end{center}
\par\noindent
	\textit{\scshape ℣.} Ángelus Dómini nuntiávit Marí{\ae}.\\
	\textit{\scshape ℟.} Et concépit de Spíritu Sancto.\\
Orémus.\\
Grátiam tuam, quǽsumus, Dómine, méntibus nostris infúnde: ut, qui, Ángelo nuntiánte, Christi Fílii tui incarnatiónem cognóvimus; per passiónem ejus et crucem, ad resurrectiónis glóriam perducámur. Per eúmdem Christum Dóminum nóstrum. Amen.
\begin{center}
	\textit{De Nativitate}
\end{center}
\par\noindent
	\textit{\scshape ℣.} Post partum, Virgo, invioláta permansísti.\\
	\textit{\scshape ℟.} Dei Génitrix, intercéde pro nobis.\\
Orémus.\\
Deus, qui salútis {\ae}térn{\ae}, beát{\ae} Marí{\ae} virginitáte fecúnda, humáno géneri prǽmia pr{\ae}stitísti: tríbue, quǽsumus; ut ipsam pro nobis intercédere sentiámus, per quam merúimus auctórem vit{\ae} suscípere, Dóminum nóstrum Jesum Christum Fílium tuum. Amen.
}
\subsection{Ave, Regina C{\ae}lorum}
\rub{From Purification through Holy Wednesday}
\englat{
\lett{Q}{ueen} of the heavens, we hail thee,\\
Hail thee, Lady of all the Angels;\\
Thou the dawn, the door of morning\\
Whence the world’s true light is risen:\par
Joy to thee, O Virgin glorious,\\
Beautiful beyond all other;\\
Hail and farewell, O most gracious,\\
Intercede for us alway to Jesus.\par\noindent
	\textit{\scshape ℣.} Vouchsafe that I may praise thee, O holy Virgin.\\
	\textit{\scshape ℟.} Give me strength against thine enemies.\par
Let us pray.\\
Grant us, O merciful God, protection in our weakness: that we who celebrate the memory of the holy Mother of God may, through the aid of her intercession, rise again from our sins. Through the same Christ our Lord. Amen.
}{
\lett{A}{ve}, Regína c{\ae}lórum,\\
Ave, Dómina Angelórum:\\
Salve radix, salve porta,\\
Ex qua mundo lux est orta:\par
Gaude, Virgo gloriósa,\\
Super omnes speciósa,\\
Vale, o valde decóra,\\
Et pro nobis Christum exóra.\par\noindent
	\textit{\scshape ℣.} Dignáre me laudáre te, Virgo sacráta.\\
	\textit{\scshape ℟.} Da mihi virtútem contra hostes tuos.\par
Orémus.\\
Concéde, miséricors Deus, fragilitáti nostr{\ae} pr{\ae}sídium; ut, qui sanct{\ae} Dei Genetrícis memóriam ágimus; intercessiónis ejus auxílio, a nostris iniquitátibus resurgámus. Per eúmdem Christum Dóminum nóstrum. Amen.
}
\subsection{Regina C{\ae}li}
\rub{From Easter Sunday until Trinity Sunday}
\englat{
\lett{O}{Queen} of heaven, be joyful, alleluia;\\
Because he whom so meetly thou barest, alleluia,\\
Hath arisen, as he promised, alleluia:\\
Pray for us to the Father, alleluia.\par\noindent
	\textit{\scshape ℣.} Rejoice and be glad, O Virgin Mary, alleluia.\\
	\textit{\scshape ℟.} For the Lord is risen indeed, alleluia.\par
Let us pray.\\
O God, who, by the resurrection of thy Son Jesus Christ didst vouchsafe to give gladness unto the world: Grant, we beseech thee, that we, being holpen by the Virgin Mary, his Mother, may attain unto the joys of everlasting life. Through the same Christ our Lord. Amen.
}	
{
\lett{R}{egína} c{\ae}li, l{\ae}táre, allelúja;\\
Quia quem meruísti portáre, allelúja,\\
Resurréxit, sicut dixit, allelúja:\\
Ora pro nobis Deum, allelúja.\par\noindent
	\textit{\scshape ℣.} Gaude et l{\ae}táre, Virgo María, allelúja.\\
	\textit{\scshape ℟.} Quia surréxit Dóminus vere, allelúja.\par
Orémus.\\
Deus, qui per resurrectiónem Fílii tui, Dómini nostri Jesu Christi, mundum l{\ae}tificáre dignátus es: pr{\ae}sta, quǽsumus; ut, per eíus Genetrícem Vírginem Maríam, perpétu{\ae} capiámus gáudia vit{\ae}. Per eúmdem Christum Dóminum nóstrum. Amen.
}
\subsection{Salve Regina}
\rub{From Trinity Sunday until Advent Sunday}
\englat{
\lett{M}{ary}, we hail thee, Mother and Queen compassionate; Mary, our comfort, life and hope, we hail thee. To thee we exiles, children of Eve, lift our crying. To thee we are sighing, as mournful and weeping, we pass through this vale of sorrow. Turn thou therefore, O our intercessor, those thine eyes of pity and loving-kindness upon us sinners. Hereafter, when our earthly exile shall be ended, show us Jesus, the blessed fruit of thy womb. O gentle, O tender, O gracious Virgin Mary.\par\noindent
	\textit{\scshape ℣.} Pray for us, O holy Mother of God.\\
	\textit{\scshape ℟.} That we may be made worthy of the promises of Christ.\par
Let us pray.\\
Almighty and everlasting God, who by the cooperation of the Holy Ghost, didst prepare the body and soul of the glorious Virgin Mother Mary to become a habitation meet for thy Son: Grant that as we rejoice in her commemoration, we may be delivered by her loving intercession from our present evils and from eternal death. Through the same Christ our Lord. Amen.
}
{
\lett{S}{alve}, Regína, mater misericórdi{\ae};
vita, dulcédo et spes nóstra, salve.\\
Ad te clamámus éxsules fílii Hev{\ae}.\\
Ad te suspirámus geméntes et flentes\\
In hac lacrimárum valle.\\
Eja ergo, advocáta nostra,\\
illos tuos misericórdes óculos ad nos convérte.\\
Et Jesum, benedíctum fructum ventris tui,\\
nobis post hoc exsílium osténde.\\
O clemens, o pia, o dulcis Virgo María.\par\noindent
	\textit{\scshape ℣.} Ora pro nobis, sancta Dei Génitrix.\\
	\textit{\scshape ℟.} Ut digni efficiámur promissiónibus Christi.\par
Orémus.\\
Omnípotens sempitérne Deus, qui gloriós{\ae} Vírginis Matris Marí{\ae} corpus et ánimam, ut dignum Fílii tui habitáculum éffici mererétur, Spíritu Sancto cooperánte, pr{\ae}parásti: da, ut, cujus commemoratióne l{\ae}támur, ejus pia intercessióne, ab instántibus malis et a morte perpétua liberémur. Per eúmdem Christum Dóminum nóstrum. Amen.
}