\section{The Order of Mattins}
\fancyhead[RE,LO]{Morning Prayer}
\elcol{℣. O Lord, {\ding{61}} open thou our lips.}{℣. Dómine, {\ding{61}} lábia nostra apéries.}
\elcol{℟. And our mouth shall show forth thy praise.}{℟. Et os nostrum annuntiábit laudem tuam.}
\elcol{℣. O God, {\ding{64}} make speed to save us.}{℣. Deus, {\ding{64}} in adiutórium nostrum inténde.}
\elcol{℟. O Lord, make haste to help us.}{℟. Dómine, ad adiuvándum nos festína.}
\elcol{℣. Glory be to the Father, and to the Son, and to the Holy Ghost.}{℣. Glória Patri, et Fílio, * et Spirítui Sancto:}
\elcol{℟. As it was in the beginning, is now, and ever shall be, world without end. Amen.}{℟. Sicut erat in princípio, et nunc, et semper, * et in sǽcula s{\ae}culórum. Amen.}
\elcol{℣. Praise ye the Lord.}{℣. Laudáte Dominum.}
\elcol{℟. The Lord's Name be praised.}{℟. Sit Nomen Domini Benedictum.}
\subsection{The Invitatory}
\rub{Then shall be said or sung the following Canticle; except on those days for which other Canticles are appointed;\\
On the appointed days, the following antiphons may be sung or said at the indicated places throughout the \textit{Venite}.}
\par
\inv{On the Sundays in Advent:} Our King and Saviour draweth nigh; * O come, let us adore him.
\par
\inv{From Christmas Day until the Epiphany:} Alleluia. Unto us a child is born; * O come, let us adore him. Alleluia.
\par
\inv{On the Epiphany and seven days after, and on the Feast of the Transfiguration:} The Lord hath manifested forth his glory; * O come, let us adore him.
\par
\inv{From Low Sunday until Ascension Day:} Alleluia. The Lord is risen indeed; * O come let us adore him. Alleluia.
\par
\inv{From Ascension Day until Whitsunday:} Alleluia. Christ the Lord ascended into heaven; * O come, let us adore him. Alleluia.
\par
\inv{On Whitsunday and six days after:} Alleluia. The Spirit of the Lord filleth the world; * O come, let us adore him. Alleluia.
\par
\inv{On Sundays in Trinitytide and optionally on ferial days:} Father, Son, and Holy Ghost, One God; * O come, let us adore him.
\par
\inv{On the Purification and the Annunciation:} The Word was made flesh; * O come, let us adore him.
\par
\inv{On other Festivals for which a proper Epistle and Gospel are ordered:} The Lord is glorious in his saints; * O come, let us adore him.
\subsection{Venite, exultemus Domino.}
\elcol{\lett{O}{Come} let us sing unto the \textsc{Lord}; * let us heartily rejoice in the strength of our salvation.
\par
Let us come before his presence with thanksgiving; * and show ourselves glad in him with psalms.
\par
For the \textsc{Lord} is a great God; * and a great King above all gods.
\par
In his hand are all the corners of the earth; * and the strength of the hills is his also.
\par
The sea is his, and he made it; * and his hands prepared the dry land.
\par
O come, let us worship and fall down, * and kneel before the \textsc{Lord} our Maker.
\par
For he is the Lord our God; * and we are the people of his pasture, and the sheep of his hand.
}{\lett{V}{eníte}, exsultémus Dómino, jubilémus Deo, salutári nostro:\par
pr{\ae}occupémus fáciem ejus in confessióne, et in psalmis jubilémus ei.\par
Quóniam Deus magnus Dóminus, et Rex magnus super omnes deos, quóniam non repéllet Dóminus plebem suam:\par
quia in manu ejus sunt omnes fines terr{\ae}, et altitúdines móntium ipse cónspicit.\par
Quóniam ipsíus est mare, et ipse fecit illud, et áridam fundavérunt manus ejus\par
veníte, adorémus, et procidámus ante Deum: plorémus coram Dómino, qui fecit nos,\par
quia ipse est Dóminus, Deus noster; nos autem pópulus ejus, et oves páscu{\ae} ejus.
}
\rub{First Ending}
\elcol{O worship the \textsc{Lord} in the beauty of holiness; * let the whole earth stand in awe of him.\par
For he cometh, for he cometh to judge the earth; * and with righteousness to judge the world, and the peoples with his truth.\par
Glory be to the Father, and to the Son, and to the Holy Ghost.\par
As it was in the beginning, is now, and ever shall be, world without end. Amen.
}{Tóllite hóstias, et introíte in átria ejus: * adoráte Dóminum in átrio sancto ejus.\par
Judicábit orbem terr{\ae} in {\ae}quitáte, * et pópulos in veritáte sua.\par
Glória Patri, et Fílio, * et Spirítui Sancto:\par
Sicut erat in princípio, et nunc, et semper, * et in sǽcula s{\ae}culórum. Amen.
}\par
\rub{Second Ending}
\elcol{To-day if ye will hear his voice, harden not your hearts * as in the provocation, and as in the day of temptation in the wilderness;\par
When your fathers tempted me, * proved me, and saw my works.\par
Forty years long was I grieved with this generation, and said, * It is a people that do err in their hearts, for they have not known my ways:\par
Unto whom I sware in my wrath, * that they should not enter into my rest.
\par
Glory be to the Father, and to the Son, and to the Holy Ghost.\par
As it was in the beginning, is now, and ever shall be, world without end. Amen.
}{Hódie, si vocem ejus audiéritis, nolíte obduráre corda vestra, sicut in exacerbatióne secúndum diem tentatiónis in desérto: ubi tentavérunt me patres vestri, probavérunt et vidérunt ópera mea.
\par
Quadragínta annis próximus fui generatióni huic, et dixi; Semper hi errant corde, ipsi vero non cognovérunt vias meas: quibus jurávi in ira mea; Si introíbunt in réquiem meam.
\par
Glória Patri, et Fílio, * et Spirítui Sancto:\par
Sicut erat in princípio, et nunc, et semper, * et in sǽcula s{\ae}culórum. Amen.
}
\rub{After the Invitatory, a hymn may be sung.\\
Then shall be sung or said a portion of the Psalms according to the use of this Book. After each Psalm (or Psalm portion) and the \textit{Benedictus} shall be prayed the \textit{Gloria Patri.}\\
After which shall be read the First Lesson, according to the Daily Office Lectionary.}
\subsection{The First Canticle}
\subsubsection{Te Deum}
\rub{The Te Deum is prayed on all Sundays (outside Advent \& Lent) and Feast Days.}
\elcol{	
\lett{W}{e} praise thee, O God; we acknowledge thee to be the Lord.\par
All the earth doth worship thee, the Father everlasting.\par
    To thee all Angels cry aloud; the Heavens, and all the Powers therein;\par
    To thee Cherubim and Seraphim continually do cry,\par
    Holy, Holy, Holy, Lord God of Sabaoth;\par
    Heaven and earth are full of the Majesty of thy glory.\par
    The glorious company of the Apostles praise thee.\par
    The goodly fellowship of the Prophets praise thee.\par
    The noble army of Martyrs praise thee.\par
    The holy Church throughout all the world doth acknowledge thee;\par
    The Father of an infinite Majesty;\par
    Thine adorable, true and only Son;\par
    Also the Holy Ghost the Comforter.
\lett{T}{hou} art the King of Glory, O Christ.\par
Thou art the everlasting Son of the Father.\par
    When thou tookest upon thee to deliver man, thou didst humble thyself to be born of a Virgin.\par
    When thou hadst overcome the sharpness of death, thou didst open the Kingdom of Heaven to all believers.\par
    Thou sittest at the right hand of God, in the glory of the Father.\par
    We believe that thou shalt come to be our Judge.\par
    We therefore pray thee, help thy servants, whom thou hast redeemed with thy precious blood.\par
    Make them to be numbered with thy Saints, in glory everlasting.
\lett{O}{Lord}, save thy people, and bless thine heritage.\par
Govern them and lift them up for ever.\par
    Day by day we magnify thee;\par
    And we worship thy Name ever, world without end.\par
    Vouchsafe, O Lord, to keep us this day without sin.\par
    O Lord, have mercy upon us, have mercy upon us.\par
    O Lord, let thy mercy be upon us, as our trust is in thee.\par
    O Lord, in thee have I trusted; let me never be confounded.}
{
\lett{T}{e} Deum laudámus: * te Dóminum confitémur.\par
Te {\ae}térnum Patrem * omnis terra venerátur.\par
Tibi omnes Ángeli, * tibi C{\ae}li, et univérs{\ae} Potestáte\par
Tibi Chérubin et Séraphin * incessábili voce proclámant:\par
Sanctus, Sanctus, Sanctus * Dóminus Deus Sábaoth.\par
Pleni sunt c{\ae}li et terra * majestátis glóri{\ae} tu{\ae}.\par
Te gloriósus * Apostolórum chorus,\par
Te Prophetárum * laudábilis númerus,\par
Te Mártyrum candidátus * laudat exércitus.\par
Te per orbem terrárum * sancta confitétur Ecclésia,\par
Patrem * imméns{\ae} majestátis;\par
Venerándum tuum verum * et únicum Fílium;\par
Sanctum quoque * Paráclitum Spíritum.
\lett{T}{u} Rex glóri{\ae}, * Christe.\par
Tu Patris * sempitérnus es Fílius.\par
Tu, ad liberándum susceptúrus hóminem: * non horruísti Vírginis úterum.\par
Tu, devícto mortis acúleo, * aperuísti credéntibus regna c{\ae}lórum.\par
Tu ad déxteram Dei sedes, * in glória Patris.\par
Judex créderis * esse ventúrus. \par
Te ergo quǽsumus, tuis fámulis súbveni, * quos pr{\ae}cióso sánguine redemísti.\par
{\AE}térna fac cum Sanctis tuis * in glória munerári.
\lett{S}{alvum} fac pópulum tuum, Dómine, * et bénedic hereditáti tu{\ae}.\par
Et rege eos, * et extólle illos usque in {\ae}térnum.\par
Per síngulos dies * benedícimus te.\par
Et laudámus nomen tuum in sǽculum, * et in sǽculum sǽculi.\par
Dignáre, Dómine, die isto * sine peccáto nos custodíre.\par
Miserére nostri, Dómine, * miserére nostri.\par
Fiat misericórdia tua, Dómine, super nos, * quemádmodum sperávimus in te.\par
In te, Dómine, sperávi: * non confúndar in {\ae}térnum.
}
\subsubsection{Benedictus es, Domine}
\rub{The Benedictus es, Domine (or another daily canticle) is prayed on all ferial days.}
\elcol{\lett{B}{lessed} art thou, O Lord God of our fathers; * praised and exalted above all for ever.\par
    Blessed art thou for the Name of thy Majesty; * praised and exalted above all for ever.\par
    Blessed art thou in the temple of thy holiness; * Praised and exalted above all for ever.\par
    Blessed art thou that beholdest the depths, and dwellest between the Cherubim: * praised and exalted above all for ever.\par
    Blessed art thou on the glorious throne of thy Kingdom: * praised and exalted above all for ever.\par
    Blessed art thou in the firmament of heaven: * praised and exalted above all for ever.
}{\lett{B}{enedictus} es, Domine Deus patrum nostrorum: et laudabilis, et gloriosus, et superexaltatus in s{\ae}cula.\par
Et benedictum nomen glori{\ae} tu{\ae} sanctum: et laudabile, et superexaltatum in omnibus s{\ae}culis.\par
Benedictus es in templo sancto glori{\ae} tu{\ae}: et superlaudabilis, et supergloriosus in s{\ae}cula.\par
Benedictus es in throno regni tui: et superlaudabilis, et superexaltatus in s{\ae}cula.\par
Benedictus es, qui intueris abyssos, et sedes super cherubim: et laudabilis, et superexaltatus in s{\ae}cula.\par
Benedictus es in firmamento c{\ae}li: et laudabilis et gloriosus in s{\ae}cula.}
\rub{After the Second Lesson, a hymn may be said or sung followed by the \textit{Benedictus},}
\subsection{The Second Canticle}
\subsubsection{Benedictus}
\englat{	
\lett{B}{lessed} {\ding{64}} be the Lord God of Israel; * for he hath visited and redeemed his people;\\
    And hath raised up a mighty salvation for us, * in the house of his servant David;\\
    As he spake by the mouth of his holy Prophets, * which have been since the world began;\\
    That we should be saved from our enemies, * and from the hand of all that hate us.\par
    To perform the mercy promised to our forefathers, * and to remember his holy covenant;\\
    To perform the oath which he sware to our forefather Abraham, * that he would give us;\\
    That we being delivered out of the hand of our enemies * might serve him without fear;\\
    In holiness and righteousness before him, * all the days of our life.\\
    And thou, child, shalt be called the prophet of the Highest: * for thou shalt go. before the face of the Lord to prepare his ways;\\
    To give knowledge of salvation unto his people * for the remission of their sins,\\
    Through the tender mercy of our God; * whereby the day-spring from on high hath visited us;\\
    To give light to them that sit in darkness, and in the shadow of death, * and to guide our feet into the way of peace.}
{\lett{B}{enedictus} {\ding{64}} Dóminus, Deus Isra{\"e}l: * quia visitávit, et fecit redemptiónem plebi su{\ae}:\\
Et eréxit cornu salútis nobis: * in domo David, púeri sui.\\
Sicut locútus est per os sanctórum, * qui a sǽculo sunt, prophetárum ejus:\\
Salútem ex inimícis nostris, * et de manu ómnium, qui odérunt nos.\\
Ad faciéndam misericórdiam cum pátribus nostris: * et memorári testaménti sui sancti.\\
Jusjurándum, quod jurávit ad Ábraham patrem nostrum, * datúrum se nobis:\\
Ut sine timóre, de manu inimicórum nostrórum liberáti, * serviámus illi.\\
In sanctitáte, et justítia coram ipso, * ómnibus diébus nostris.\\
Et tu, puer, Prophéta Altíssimi vocáberis: * pr{\ae}íbis enim ante fáciem Dómini, paráre vias ejus:\\
Ad dandam sciéntiam salútis plebi ejus: * in remissiónem peccatórum eórum:\\
Per víscera misericórdi{\ae} Dei nostri: * in quibus visitávit nos, óriens ex alto:\\
Illumináre his, qui in ténebris, et in umbra mortis sedent: * ad dirigéndos pedes nostros in viam pacis.
}
\subsection{The Apostles' Creed}
\elcol{
\lett{I}{Believe} in God the Father Almighty, Maker of heaven and earth: And in Jesus Christ his only Son our Lord: Who was conceived by the Holy Ghost, Born of the Virgin Mary: Suffered under Pontius Pilate, Was crucified, dead, and buried: He descended into hell; The third day he rose again from the dead: He ascended into heaven, And sitteth on the right hand of God the Father Almighty: From thence he shall come to judge the quick and the dead.\par
    I believe in the Holy Ghost: The holy Catholic Church; The Communion of Saints: The Forgiveness of sins: The Resurrection of the body: {\ding{64}} And the Life everlasting. Amen.
}{
\lett{C}{redo} in Deum, Patrem omnipoténtem, Creatórem c{\ae}li et terr{\ae}. Et in Jesum Christum, Fílium ejus únicum, Dóminum nostrum: qui concéptus est de Spíritu Sancto, natus ex María Vírgine, passus sub Póntio Piláto, crucifíxus, mórtuus, et sepúltus: descéndit ad ínferos; tértia die resurréxit a mórtuis; ascéndit ad c{\ae}los; sedet ad déxteram Dei Patris omnipoténtis: inde ventúrus est judicáre vivos et mórtuos.\par
     Credo in Spíritum Sanctum, sanctam Ecclésiam cathólicam, Sanctórum communiónem, remissiónem peccatórum, carnis {\ding{64}} resurrectiónem, vitam {\ae}térnam. Amen.
 }
\rub{And after that, these Prayers following, the people devoutly kneeling; the Minister first pronouncing (the words in parentheses if he is not in major orders),}
\elcol{℣. The Lord be with you. (O Lord, hear my prayer.)}{℣. Dóminus vobíscum. (Dómine, exáudi oratiónem meam.)}
\elcol{℟. And with thy spirit. (And let my cry come unto thee.)}{℟. Et cum spiritu tuo. (Et clamor meus ad te véniat.)}
\elcol{℣. Let us pray.}{℣. Orémus.}
\elcol{℣. Lord, have mercy upon us.}{℣. Kýrie, eléison.}
\elcol{℟. Christ, have mercy upon us.}{℟. Christe, eléison.}
\elcol{℣. Lord, have mercy upon us.}{℣. Kýrie, eléison.}
\subsection{Our Father}
\rub{All kneeling shall say the Lord's Prayer.}
\englat{
\lett{O}{ur} Father, who art in heaven, Hallowed be thy Name. Thy kingdom come. Thy will be done on earth, As it is in heaven. Give us this day our daily bread. And forgive us our trespasses, As we forgive those who trespass against us. And lead us not into temptation; But deliver us from evil. Amen.}
{
\lett{P}{ater} noster, qui es in c{\ae}lis, sanctificétur nomen tuum: advéniat regnum tuum: fiat volúntas tua, sicut in c{\ae}lo et in terra. Panem nostrum quotidiánum da nobis hódie: et dimítte nobis débita nostra, sicut et nos dimíttimus debitóribus nostris: et ne nos indúcas in tentatiónem: sed líbera nos a malo. Amen.}
\subsection{The Preces}
\rub{The Minister, standing, shall say:}
\elcol{℣. O Lord, show thy mercy upon us.}{℣. Ostende nobis Domine misericordiam tuam.}
\elcol{℟. And grant us thy salvation.}{℟. Et salutare tuum da nobis.}
\elcol{℣. O God, make clean our hearts within us.}{℣. Cor mundum crea in nobis, O Deus.}
\elcol{℟. And take not thy Holy Spirit from us.}{℟. Et Spiritum Sanctum tuum ne auferas a nobis.}
\subsection{The Collects}
\rub{Then shall follow the Collect for the Day and then the following collects.}
\subsubsection{A Collect for Peace}
\lett{O}{God}, who art the author of peace and lover of concord, in knowledge of whom standeth our eternal life, whose service is perfect freedom; Defend us thy humble servants in all assaults of our enemies; that we, surely trusting in thy defence, may not fear the power of any adversaries, through the might of Jesus Christ our Lord. Amen.
\subsubsection{A Collect for Grace}
\lett{O}{Lord}, our heavenly Father, Almighty and everlasting God, who hast safely brought us to the beginning of this day; Defend us in the same with thy mighty power; and grant that this day we fall into no sin, neither run into any kind of danger; but that all our doings, being ordered by thy governance, may be righteous in thy sight; through Jesus Christ our Lord. Amen.
\subsection{Conclusion}
\elcol{℣. The Lord be with you. (O Lord, hear my prayer.)}{℣. Dóminus vobíscum. (Dómine, exáudi oratiónem meam.)}
\elcol{℟. And with thy spirit. (And let my cry come unto thee.)}{℟. Et cum spiritu tuo. (Et clamor meus ad te véniat.)}
\elcol{℣. Let us bless the Lord. (alleluia, alleluia.)}{℣. Benedicámus Dómino (allelúia, allelúia).}
\elcol{℟. Thanks be to God. (alleluia, alleluia.)}{℟. Deo grátias (allelúia, allelúia).}
\elcol{℣. May the souls {\ding{64}} of the faithful departed, through the mercy of God, rest in peace.}{℣. Fidélium ánim{\ae} {\ding{64}} per misericórdiam Dei requiéscant in pace.}
\elcol{℟. Amen.}{℟. Amen.}
\rub{The appropriate Marian Anthem (p. \pageref{mary}) may be said or sung here.}
\subsection{Third Collects}
\subsubsection{A Prayer for The President of the United States, and all in Civil Authority}
\lett{O}{Lord}, our heavenly Father, the high and mighty Ruler of the universe, who dost from thy throne behold all the dwellers upon earth; Most heartily we beseech thee, with thy favour to behold and bless thy servant THE PRESIDENT OF THE UNITED STATES, and all others in authority; and so replenish them with the grace of thy Holy Spirit, that they may always incline to thy will, and walk in thy way. Endue them plenteously with heavenly gifts; grant them in health and prosperity long to live; and finally, after this life, to attain everlasting joy and felicity; through Jesus Christ our Lord. Amen.
\subsubsection{A Prayer for the Clergy and People}
\lett{A}{lmighty} and everlasting God, from whom cometh every good and perfect gift; Send down upon our Bishops, and other Clergy, and upon the Congregations committed to their charge, the healthful Spirit of thy grace; and, that they may truly please thee, pour upon them the continual dew of thy blessing. Grant this, O Lord, for the honour of our Advocate arid Mediator, Jesus Christ. Amen.
\subsubsection{A Prayer for all Sorts \& Conditions of Men}
\lett{O}{God}, the Creator and Preserver of all mankind, we humbly beseech thee for all sorts and conditions of men; that thou wouldest be pleased to make thy ways known unto them, thy saving health unto all nations. More especially we pray for thy holy Church universal; that it may be so guided and governed by thy good Spirit, that all who profess and call themselves Christians may be led into the way of truth, and hold the faith in unity of spirit, in the bond of peace, and in righteousness of life. Finally, we commend to thy fatherly goodness all those who are any ways afflicted, or distressed, in mind, body, or estate; (especially those for whom our prayers are desired;) that it may please thee to comfort and relieve them, according to. their several necessities; giving them patience under their sufferings, and a happy issue out of all their afflictions. And this we beg for Jesus Christ's sake. Amen.
\subsubsection{A General Thanksgiving}
\rub{The General Thanksgiving may be said by the Congregation with the Minister.}
\lett{A}{lmighty} God, Father of all mercies, we, thine unworthy servants, do give thee most humble and hearty thanks for all thy goodness and loving-kindness to us and to all men; (to those who desire now to offer up their praises and thanksgivings for thy late mercies vouchsafed unto them.) We bless thee for our creation, preservation, and all the blessings of this life; but above all, for thine inestimable love in the redemption of the world by our Lord Jesus Christ; for the means of grace, and for the hope of glory. And, we beseech thee, give us that due sense of all thy mercies, that our hearts may he unfeignedly thankful; and that we show forth thy praise, not only with our lips, but in our lives, by giving up our selves to thy service, and by walking before thee in holiness and righteousness all our days; through Jesus Christ our Lord, to whom, with thee and the Holy Ghost, be all honour and glory, world without end. Amen.
\subsubsection{A Prayer of St. Chrysostom}
\lett{A}{lmighty} God, who hast given us grace at this time with one accord to make our common supplications unto thee; and dost promise that when two or three are gathered together in thy Name thou wilt grant their requests; Fulfil now, O Lord, the desires and petitions of thy servants, as may be most expedient for them; granting us in this world knowledge of thy truth, and in the world to come life everlasting. Amen.
\subsubsection{The Grace}
\lett{T}{he} grace of our Lord Jesus Christ, {\ding{64}} and the love of God, and the fellowship of the Holy Ghost, be with us all evermore. Amen.\par
\begin{center}
	\textsc{Here endeth the Order of Morning Prayer.}
\end{center}