\section{The Hour of None}
\fancyhead[RE,LO]{The Hour of None}
\subsection{Introduction}
\englat
{
\textit{\scshape ℣.} O God, {\ding{64}} make speed to save us.\\
\textit{\scshape ℟.} O Lord, make haste to help us.\\
\textit{\scshape ℣.} Glory be to the Father, and to the Son, and to the Holy Ghost.\\
\textit{\scshape ℟.} As it was in the beginning, is now, and ever shall be, world without end. Amen.\\
\textit{\scshape ℣.} Praise ye the Lord.\\
\textit{\scshape ℟.} The Lord's Name be praised.\\
}
{
\textit{\scshape ℣.} Deus, {\ding{64}} in adjutorium nostrum intende.\\
\textit{\scshape ℟.} Domine, ad adjuvandum nos festina.\\
\textit{\scshape ℣.} Glória Patri, et Fílio, * et Spirítui Sancto:\\
\textit{\scshape ℟.} Sicut erat in princípio, et nunc, et semper, * et in sǽcula s{\ae}culórum. Amen.\\
\textit{\scshape ℣.} Laudate Dominum.\\
\textit{\scshape ℟.} Sit Nomen Domini Benedictum.\\
}
\subsection{Rerum Deus tenax vigor}
\englat{
O God, the world's sustaining force,\\
Thyself unmoved, all motion's source.\\
Who from the morn till evening's ray\\
Through all its changes guidest the day:\\
\par
\noindent
O grant us light at eventide,\\
That life may unimpaired abide,\\
And that a holy death may be\\
The door of immortality.\\
\par
\noindent
O Father, that we ask be done\\
Through Jesus Christ, thine only Son,\\
Who, with the Holy Ghost and thee,\\
Shall live and reign eternally. Amen.
}{
Rerum, Deus, tenax vigor,\\
Immótus in te pérmanens,\\
Lucis diúrn{\ae} témpora\\
Succéssibus detérminans:\\
\par\noindent
Largíre lumen véspere,\\
Quo vita nusquam décidat,\\
Sed prǽmium mortis sacr{\ae}\\
Perénnis instet glória.\\
\par\noindent
Pr{\ae}sta, Pater piíssime,\\
Patríque compar Únice,\\
Cum Spíritu Paráclito\\
Regnans per omne sǽculum.\\
Amen.
}
\subsection{The Psalms}
\inv{Ant. throughout the Year:} Alleluia. *\\
\inv{Ant. in Advent:} Lo, now cometh. *\\
\inv{Ant. from Christmas until Candlemas.} Whom beheld ye. *\\
\inv{Ant. in Pre-Lent and Lent.} They that fear the Lord. *\\
\inv{Ant. in Eastertide.} Alleluia. *
\subsubsection{Psalm 119:XI \textit{Defecit anima mea.}}
\rub{Sunday, Monday, Wednesday, Friday}
\lett{M}{y} soul hath longed for thy salvation, * and I have a good hope because of thy word.
\par
    82 Mine eyes long sore for thy word; * saying, O when wilt thou comfort me?
\par
    83 For I am become like a bottle in the smoke; * yet do I not forget thy statutes.
\par
    84 How many are the days of thy servant? * when wilt thou be avenged of them that persecute me?
\par
    85 The proud have digged pits for me, * which are not after thy law.
\par
    86 All thy commandments are true: * they persecute me falsely; O be thou my help.
\par
    87 They had almost made an end of me upon earth; * but I forsook not thy commandments.
\par
    88 O quicken me after thy loving-kindness; * and so shall I keep the testimonies of thy mouth.
\subsubsection{Psalm 119:XII \textit{In {\ae}ternum, Domine.}}
\rub{Sunday, Monday, Wednesday, Friday}
\lett{O}{Lord}, thy word * endureth for ever in heaven.
\par
    90 Thy truth also remaineth from one generation to another; * thou hast laid the foundation of the earth, and it abideth.
\par
    91 They continue this day according to thine ordinance; * for all things serve thee.
\par
    92 If my delight had not been in thy law, * I should have perished in my trouble.
\par
    93 I will never forget thy commandments; * for with them thou hast quickened me.
\par
    94 I am thine: O save me, * for I have sought thy commandments.
\par
    95 The ungodly laid wait for me, to destroy me; * but I will consider thy testimonies.
\par
    96 I see that all things come to an end; * but thy commandment is exceeding broad.
\subsubsection{Psalm 119:XIII \textit{Quomodo dilexi!}}
\rub{Sunday, Monday, Wednesday, Friday}
\lett{L}{ord}, what love have I unto thy law! * all the day long is my study in it.
\par
    98 Thou, through thy commandments, hast made me wiser than mine enemies; * for they are ever with me.
\par
    99 I have more understanding than my teachers; * for thy testimonies are my study.
\par
    100 I am wiser than the aged; * because I keep thy commandments.
\par
    101 I have refrained my feet from every evil way, * that I may keep thy word.
\par
    102 I have not shrunk from thy judgments; * for thou teachest me.
\par
    103 O how sweet are thy words unto my throat; * yea, sweeter than honey unto my mouth!
\par
    104 Through thy commandments I get understanding: * therefore I hate all evil ways. 
\subsubsection{Psalm 121 \textit{Levavi oculos.}}
\rub{Tuesday, Thursday, Saturday}
\lett{I}{will} lift up mine eyes unto the hills; * from whence cometh my help?
\par
    2 My help cometh even from the \textsc{Lord}, * who hath made heaven and earth.
\par
    3 He will not suffer thy foot to be moved; * and he that keepeth thee will not sleep.
\par
    4 Behold, he that keepeth Israel * shall neither slumber nor sleep.
\par
    5 The \textsc{Lord} himself is thy keeper; * the \textsc{Lord} is thy defence upon thy right hand;
\par6 So that the sun shall not burn thee by day, * neither the moon by night.
\par
    7 The \textsc{Lord} shall preserve thee from all evil; * yea, it is even he that shall keep thy soul.
\par
    8 The \textsc{Lord} shall preserve thy going out, and thy coming in, * from this time forth for evermore.
\subsubsection{Psalm 122 \textit{L{\ae}tatus sum.}}
\rub{Tuesday, Thursday, Saturday}
\lett{I}{was} glad when they said unto me, * We will go into the house of the \textsc{Lord}.
\par
    2 Our feet shall stand in thy gates, * O Jerusalem.
\par
    3 Jerusalem is built as a city * that is at unity in itself.
\par
    4 For thither the tribes go up, even the tribes of the \textsc{Lord}, * to testify unto Israel, to give thanks unto the Name of the \textsc{Lord}.
\par
    5 For there is the seat of judgment, * even the seat of the house of David.
\par
    6 O pray for the peace of Jerusalem; * they shall prosper that love thee.
\par
    7 Peace be within thy walls, * and plenteousness within thy palaces.
\par
    8 For my brethren and companions' sakes, * I will wish thee prosperity.
\par
    9 Yea, because of the house of the \textsc{Lord} our God, * I will seek to do thee good.
\subsubsection{Psalm 123 \textit{Ad te levavi oculos meos.}}
\rub{Tuesday, Thursday, Saturday}
\lett{U}{nto} thee lift I up mine eyes, * O thou that dwellest in the heavens.
\par
    2 Behold, even as the eyes of servants look unto the hand of their masters, and as the eyes of a maiden unto the hand of her mistress, * even so our eyes wait upon the \textsc{Lord} our God, until he have mercy upon us.
\par
    3 Have mercy upon us, O \textsc{Lord}, have mercy upon us; * for we are utterly despised.
\par
    4 Our soul is filled with the scornful reproof of the wealthy, * and with the despitefulness of the proud.
\par\noindent
\inv{Ant. throughout the Year:} Alleluia. * O Give thanks unto the Lord, because his mercy endureth for ever. Alleluia.
\inv{Ant. in Advent:} Lo, now cometh * the fulness of the time, wherein God sent forth his Son into the world.
\inv{Ant. from Christmas until Candlemas.} Whom beheld ye, * O shepherds, tell us? Declare to us the tidings; on earth, who hath appeared? We beheld an infant, the new-born Lord and Saviour, amidst a choir of angels, alleluia, alleluia.
\inv{Ant. in Pre-Lent and Lent.} They that fear the Lord * confess that his mercy endureth for ever.
\inv{Ant. in Eastertide.} Alleluia, * alleluia, alleluia, alleluia.
\subsection{The Chapter}
\rub{Then the following, or another suitable Chapter, shall be read.}
One Lord, one Faith, one Baptism, one God and Father of all, who is above all, and through all, and in you all. \vr{Eph. 4:5}\\
\textit{\scshape ℟.} Thanks be to God.\\
\par\noindent
\textit{\scshape ℣.} I call with my whole heart; hear me, O Lord (alleluia, alleluia).\\
\textit{\scshape ℟.} I call with my whole heart; hear me, O Lord (alleluia, alleluia).\\
\textit{\scshape ℣.} I will keep thy statutes.\\
\textit{\scshape ℟.} Hear me, O Lord (alleluia, alleluia).\\
\textit{\scshape ℣.} Glory be to the Father, and to the Son, and to the Holy Ghost.\\
\textit{\scshape ℟.} I call with my whole heart; hear me, O Lord (alleluia, alleluia).\\
\par\noindent
\textit{\scshape ℣.} Cleanse me, O Lord, from my secret faults (alleluia).\\
\textit{\scshape ℟.} Keep thy servant also from presumptuous sins (alleluia).
\rub{All kneeling:}
\englat{
\textit{\scshape ℣.} Lord, have mercy upon us.\\
\textit{\scshape ℟.} Christ, have mercy upon us.\\
\textit{\scshape ℣.} Lord, have mercy upon us.}\\
\lett{O}{ur} Father, who art in heaven, Hallowed be thy Name. Thy kingdom come. Thy will be done on earth, As it is in heaven. Give us this day our daily bread. And forgive us our trespasses, As we forgive those who trespass against us. And lead us not into temptation; But deliver us from evil. Amen.
{
\textit{\scshape ℣.} Kýrie, eléison.\\
\textit{\scshape ℟.} Christe, eléison.\\
\textit{\scshape ℣.} Kýrie, eléison.\\
\lett{P}{ater} noster, qui es in c{\ae}lis, sanctificétur nomen tuum: advéniat regnum tuum: fiat volúntas tua, sicut in c{\ae}lo et in terra. Panem nostrum quotidiánum da nobis hódie: et dimítte nobis débita nostra, sicut et nos dimíttimus debitóribus nostris: et ne nos indúcas in tentatiónem: sed líbera nos a malo. Amen.}
\textit{\scshape ℣.} Cast me not away in the time of grace.\\
\textit{\scshape ℟.} Forsake me not when my strength faileth me.\\
\textit{\scshape ℣.} Hide not thy face from me.\\
\textit{\scshape ℟.} Lest I be like unto them that go down into the pit.\\
\par
\textit{\scshape ℣.} O Lord, hear our prayer.\\
\textit{\scshape ℟.} And let our cry come unto thee.\\
\textit{\scshape ℣.} The Lord be with you.\\
\textit{\scshape ℟.} And with thy spirit.\\
\textit{\scshape ℣.} Let us pray.
\rub{The Dominus Vobiscum is only said if the Minister is in Major Orders.}
\subsection{The Collect}
\lett{O}{Lord} Jesus Christ, who for our sakes didst tread the paths of death; make known to us the way of life; that as thou wast reckoned with the transgressors in thy death, and with the rich in thy burial, so we, who are dead in trespasses and sins, may be raised up by thee to the land of true riches. Who livest and reignest with the Father and the Holy Ghost, ever one God, world without end.\\
\textit{\scshape ℟.} Amen.\\
\textit{\scshape ℣.} The Lord be with you (O Lord, hear my prayer).\\
\textit{\scshape ℟.} And with thy spirit (And let my cry come unto thee).\\
\textit{\scshape ℣.} Let us bless the Lord.\\
\textit{\scshape ℟.} Thanks be to God.\\
\textit{\scshape ℣.} May the souls {\ding{64}} of the faithful departed, through the mercy of God, rest in peace.\\
\textit{\scshape ℟.} Amen.