\section{The Hour of Prime}
\fancyhead[RE,LO]{The Hour of Prime}
\subsection{Introduction}
\englat
{
\textit{\scshape ℣.} O God, {\ding{64}} make speed to save us.\\
\textit{\scshape ℟.} O Lord, make haste to help us.\\
\textit{\scshape ℣.} Glory be to the Father, and to the Son, and to the Holy Ghost.\\
\textit{\scshape ℟.} As it was in the beginning, is now, and ever shall be, world without end. Amen.\\
\textit{\scshape ℣.} Praise ye the Lord.\\
\textit{\scshape ℟.} The Lord's Name be praised.\\
}
{
\textit{\scshape ℣.} Deus, {\ding{64}} in adjutorium nostrum intende.\\
\textit{\scshape ℟.} Domine, ad adjuvandum nos festina.\\
\textit{\scshape ℣.} Glória Patri, et Fílio, * et Spirítui Sancto:\\
\textit{\scshape ℟.} Sicut erat in princípio, et nunc, et semper, * et in sǽcula s{\ae}culórum. Amen.\\
\textit{\scshape ℣.} Laudate Dominum.\\
\textit{\scshape ℟.} Sit Nomen Domini Benedictum.\\
}
\subsection{Jam Lucis Ordo Sidere}
\englat{
Now that the daylight fills the sky,\\
We lift our hearts to God on high,\\
That he, in all we do or say,\\
Would keep us free from harm today.\\
\par
Would guard our hearts and tongues from strife;\\
From anger's din would hide our life;\\
From all ill sights would turn our eyes;\\
Would close our ears from vanities;\\
\par
Would keep our inmost conscience pure;\\
Our souls from folly would secure;\\
Would bid us check the pride of sense\\
With due and holy abstinence.\\
\par
So we, when this new day is gone,\\
And night in turn is drawing on,\\
With conscience by the world unstained,\\
Shall praise his Name for victory gained.\\
\par
All laud to God the Father be;\\
All praise, eternal Son, to thee;\\
All glory, as is ever meet,\\
To God the Holy Paraclete. Amen.
}{
Jam lucis orto sídere,\\
Deum precémur súpplices,\\
Ut in diúrnis áctibus\\
Nos servet a nocéntibus.\\
\par
Linguam refrénans témperet,\\
Ne litis horror ínsonet:\\
Visum fovéndo cóntegat,\\
Ne vanitátes háuriat.\\
\par
Sint pura cordis íntima,\\
Absístat et vecórdia;\\
Carnis terat supérbiam\\
Potus cibíque párcitas.\\
\par
Ut, cum dies abscésserit,\\
Noctémque sors redúxerit,\\
Mundi per abstinéntiam\\
Ipsi canámus glóriam.\\
\par
Deo Patri sit glória,\\
Ejúsque soli Fílio,\\
Cum Spíritu Paráclito,\\
Nunc et per omne sǽculum.\\
Amen.
}
\subsection{The Psalms}
\inv{Ant. throughout the Year:} Alleluia. *\\
\inv{Ant. in Advent:} Lo, now cometh. *\\
\inv{Ant. from Christmas until Candlemas.} Whom beheld ye. *\\
\inv{Ant. in Pre-Lent and Lent.} They that fear the Lord. *\\
\inv{Ant. in Eastertide.} Alleluia. *
\subsubsection{Psalm 54. \textit{Deus, in Nomine.}}
\rub{Sunday}
\lett{S}{ave} me, O God, for thy Name's sake, * and avenge me in thy strength.\par
2 Hear my prayer, O God, * and hearken unto the words of my mouth.\par
3 For strangers are risen up against me; * and tyrants, which have not God before their eyes, seek after my soul.\par
4 Behold, God is my helper; * the Lord is with them that uphold my soul.\par
5 He shall reward evil unto mine enemies: * destroy thou them in thy truth.\par
6 An offering of a free heart will I give thee, and praise thy Name, O \textsc{Lord}; * because it is so comfortable.\par
7 For he hath delivered me out of all my trouble; * and mine eye hath seen his desire upon mine enemies.
\subsubsection{Psalm 118. \textit{Confitemini Domino.}}
\rub{Sunday}
\lett{O}{Give} thanks unto the \textsc{Lord}, for he is gracious; * because his mercy endureth for ever.
\par
    2 Let Israel now confess that he is gracious, * and that his mercy endureth for ever.
\par
    3 Let the house of Aaron now confess, * that his mercy endureth for ever.
\par
    4 Yea, let them now that fear the \textsc{Lord} confess, * that his mercy endureth for ever.
\par
    5 I called upon the \textsc{Lord} in trouble; * and the \textsc{Lord} heard me at large.
\par
    6 The \textsc{Lord} is on my side; * I will not fear what man doeth unto me.
\par
    7 The \textsc{Lord} taketh my part with them that help me; * therefore shall I see my desire upon mine enemies.
\par
    8 It is better to trust in the \textsc{Lord}, * than to put any confidence in man.
\par
    9 It is better to trust in the \textsc{Lord}, * than to put any confidence in princes.
\par
    10 All nations compassed me round about; * but in the Name of the \textsc{Lord} will I destroy them.
\par
    11 They kept me in on every side, they kept me in, I say, on every side; * but in the Name of the \textsc{Lord} will I destroy them.
\par
    12 They came about me like bees, and are extinct even as the fire among the thorns; * for in the Name of the \textsc{Lord} I will destroy them.
\par
    13 Thou hast thrust sore at me, that I might fall; * but the \textsc{Lord} was my help.
\par
    14 The \textsc{Lord} is my strength, and my song; * and is become my salvation.
\par
    15 The voice of joy and health is in the dwellings of the righteous; * the right hand of the \textsc{Lord} bringeth mighty things to pass.
\par
    16 The right hand of the \textsc{Lord} hath the preeminence; * the right hand of the \textsc{Lord} bringeth mighty things to pass.
\par
    17 I shall not die, but live, * and declare the works of the \textsc{Lord}.
\par
    18 The \textsc{Lord} hath chastened and corrected me; * but he hath not given me over unto death.
\par
    19 Open me the gates of righteousness, * that I may go into them, and give thanks unto the \textsc{Lord}.
\par
    20 This is the gate of the \textsc{Lord}, * the righteous shall enter into it.
\par
    21 I will thank thee; for thou hast heard me, * and art become my salvation.
\par
    22 The same stone which the builders refused, * is become the head-stone in the corner.
\par
    23 This is the \textsc{Lord's} doing, * and it is marvellous in our eyes.
\par
    24 This is the day which the \textsc{Lord} hath made; * we will rejoice and be glad in it.
\par
    25 Help me now, O \textsc{Lord}: * O \textsc{Lord}, send us now prosperity.
\par
    26 Blessed be he that cometh in the Name of the \textsc{Lord}: * we have wished you good luck, we that are of the house of the \textsc{Lord}.
\par
    27 God is the \textsc{Lord}, who hath showed us light: * bind the sacrifice with cords, yea, even unto the horns of the altar.
\par
    28 Thou art my God, and I will thank thee; * thou art my God, and I will praise thee.
\par
    29 O give thanks unto the \textsc{Lord}; for he is gracious, * and his mercy endureth for ever.
\subsubsection{Psalm 119:I-II \textit{Beati immaculati.}}
\rub{Sunday, Monday, Wednesday, Friday}
\lett{B}{lessed} are those that are undefiled in the way, and walk in the law of the \textsc{Lord}.\par
2 Blessed are they that keep his testimonies, * and seek him with their whole heart;\par
3 Even they who do no wickedness, * and walk in his ways.\par
4 Thou hast charged * that we shall diligently keep thy commandments.\par
5 O that my ways were made so direct, * that I might keep thy statutes!\par
6 So shall I not be confounded, * while I have respect unto all thy commandments.\par
7 I will thank thee with an unfeigned heart, * when I shall have learned the judgments of thy righteousness.\par   
8 I will keep thy statutes1; * O forsake me not utterly.
\par
9 Wherewithal shall a young man cleanse his way? * even by ruling himself after thy word.\par
10 With my whole heart have I sought thee; * O let me not go wrong out of thy commandments.\par
11 Thy word have I hid within my heart, * that I should not sin against thee.\par
12 Blessed art thou, O \textsc{Lord}; * O teach me thy statutes.\par
13 With my lips have I been telling * of all the judgments of thy mouth.\par
14 I have had as great delight in the way of thy testimonies, * as in all manner of riches.\par
15 I will talk of thy commandments, * and have respect unto thy ways.\par
16 My delight shall be in thy statutes, * and I will not forget thy word.
\subsubsection{Psalm 119:III-IV \textit{Retribue servo tuo.}}
\rub{Sunday, Monday, Wednesday, Friday}
\lett{O}{Do} well unto thy servant; * that I may live, and keep thy word.\par
18 Open thou mine eyes; * that I may see the wondrous things of thy law.\par
19 I am a stranger upon earth; * O hide not thy commandments from me.\par
20 My soul breaketh out for the very fervent desire * that it hath alway unto thy judgments.\par
21 Thou hast rebuked the proud; * and cursed are they that do err from thy commandments.\par
22 O turn from me shame and rebuke; * for I have kept thy testimonies.\par
23 Princes also did sit and speak against me; * but thy servant is occupied in thy statutes.\par
24 For thy testimonies are my delight, * and my counsellors.
\par
25 My soul cleaveth to the dust; * O quicken thou me, according to thy word.
\par
    26 I have acknowledged my ways, and thou heardest me: * O teach me thy statutes.
\par
    27 Make me to understand the way of thy commandments; * and so shall I talk of thy wondrous works.
\par
    28 My soul melteth away for very heaviness; * comfort thou me according unto thy word.
\par
    29 Take from me the way of lying, * and cause thou me to make much of thy law.
\par
    30 I have chosen the way of truth, * and thy judgments have I laid before me.
\par
    31 I have stuck unto thy testimonies; * O \textsc{Lord}, confound me not.
\par
    32 I will run the way of thy commandments, * when thou hast set my heart at liberty.
\subsubsection{Psalm XIV-XV. \textit{Lucerna pedibus meis.}}
\rub{Tuesday, Thursday, Saturday}
\lett{T}{hy} word is a lantern unto my feet, * and a light unto my paths.
\par
    106 I have sworn, and am stedfastly purposed, * to keep thy righteous judgments.
\par
    107 I am troubled above measure: * quicken me, O \textsc{Lord}, according to thy word.
\par
    108 Let the free-will offerings of my mouth please thee, O \textsc{Lord}; * and teach me thy judgments.
\par
    109 My soul is alway in my hand; * yet do I not forget thy law.
\par
    110 The ungodly have laid a snare for me; * but yet I swerved not from thy commandments.
\par
    111 Thy testimonies have I claimed as mine heritage for ever; * and why? they are the very joy of my heart.
\par
    112 I have applied my heart to fulfil thy statutes alway, * even unto the end.
\par
I hate them that imagine evil things; * but thy law do I love.
\par
    114 Thou art my defence and shield; * and my trust is in thy word.
\par
    115 Away from me, ye wicked; * I will keep the commandments of my God.
\par
    116 O stablish me according to thy word, that I may live; * and let me not be disappointed of my hope.
\par
    117 Hold thou me up, and I shall be safe; * yea, my delight shall be ever in thy statutes.
\par
    118 Thou hast trodden down all them that depart from thy statutes; * for they imagine but deceit.
\par
    119 Thou puttest away all the ungodly of the earth like dross; * therefore I love thy testimonies.
\par
    120 My flesh trembleth for fear of thee; * and I am afraid of thy judgments.
\subsubsection{Psalm 119:XVI-XVII. \textit{Feci judicium.}}
\rub{Tuesday, Thursday, Saturday}
\lett{I}{Deal} with the thing that is lawful and right; * O give me not over unto mine oppressors.
\par
    122 Make thou thy servant to delight in that which is good, * that the proud do me no wrong.
\par
    123 Mine eyes are wasted away with looking for thy health, * and for the word of thy righteousness.
\par
    124 O deal with thy servant according unto thy loving mercy, * and teach me thy statutes.
\par
    125 I am thy servant; O grant me understanding, * that I may know thy testimonies.
\par
    126 It is time for thee, \textsc{Lord}, to lay to thine hand; * for they have destroyed thy law.
\par
    127 For I love thy commandments * above gold and precious stones.
\par
    128 Therefore hold I straight all thy commandments; * and all false ways I utterly abhor.
\par
    129 Thy testimonies are wonderful; * therefore doth my soul keep them.
\par
    130 When thy word goeth forth, * it giveth light and understanding unto the simple.
\par
    131 I opened my mouth, and drew in my breath; * for my delight was in thy commandments.
\par
    132 O look thou upon me, and be merciful unto me, * as thou usest to do unto those that love thy Name.
\par
    133 Order my steps in thy word; * and so shall no wickedness have dominion over me.
\par
    134 O deliver me from the wrongful dealings of men; * and so shall I keep thy commandments.
\par
    135 Show the light of thy countenance upon thy servant, * and teach me thy statutes.
\par
    136 Mine eyes gush out with water, * because men keep not thy law.
\rub{On Sundays in Trinitytide and after Epiphany until Septuagesimma, the Athanasian Creed may be added here.}
\inv{Ant. throughout the Year:} Alleluia. * O Give thanks unto the Lord, because his mercy endureth for ever. Alleluia.\\
\inv{Ant. in Advent:} Lo, now cometh * the fulness of the time, wherein God sent forth his Son into the world.\\
\inv{Ant. from Christmas until Candlemas.} Whom beheld ye, * O shepherds, tell us? Declare to us the tidings; on earth, who hath appeared? We beheld an infant, the new-born Lord and Saviour, amidst a choir of angels, alleluia, alleluia.\\
\inv{Ant. in Pre-Lent and Lent.} They that fear the Lord * confess that his mercy endureth for ever.\\
\inv{Ant. in Eastertide.} Alleluia, * alleluia, alleluia, alleluia.
\subsection{The Chapter}
\subsubsection{Ferias throughout the Year.}
Love ye the truth and peace, saith the Lord God of hosts. \vr{Zech. 8:19}
\subsubsection{Sundays \& Feasts}
Now unto the King eternal, immortal, invisible, the only wise God, be honour and glory for ever and ever. \vr{I Tim. 1:17}
\subsubsection{Easter through Whitsuntide}
O Lord, be gracious unto us; we have waited for thee: be thou our arm every morning, our salvation also in the time of trouble. \vr{Is. 33:8}\\
\textit{\scshape ℟.} Thanks be to God.
\subsection{The Short Respond}
\englat{
\textit{\scshape ℣.} Chris Jesu, Son of the Living God, have mercy upon us (alleluia, alleluia).\\
\textit{\scshape ℟.} Chris Jesu, Son of the Living God, have mercy upon us (alleluia, alleluia).\\
\textit{\scshape ℣.} Thou that sittest at the right hand of the Father.\\
\textit{\scshape ℟.} Have mercy upon us \inrub{Replaced in Eastertide with: alleluia, alleluia}.\\
\textit{\scshape ℣.} Glory be to the Father, and to the Son, and to the Holy Ghost.\\
\textit{\scshape ℟.} Chris Jesu, Son of the Living God, have mercy upon us (alleluia, alleluia).\\
\par\noindent
\textit{\scshape ℣.} O Christ, arise, help us (alleluia).\\
\textit{\scshape ℟.} And deliver us for thy Name's sake (alleluia).
}{\textit{\scshape ℣.} Christe, Fili Dei vivi, Miserére nobis (allelúja, allelúja). \\
\textit{\scshape ℟.} Christe, Fili Dei vivi, Miserére nobis (allelúja, allelúja).\\
\textit{\scshape ℣.} Qui sedes ad déxteram Patris.\\
\textit{\scshape ℟.} Miserére nobis \inrub{Replaced in Eastertide with allelúja, allelúja}.\\
\textit{\scshape ℣.} Glória Patri, et Fílio, et Spirítui Sancto.\\
\textit{\scshape ℟.} Christe, Fili Dei vivi, * miserére nobis (allelúja, allelúja).\\
\par\noindent
\textit{\scshape ℣.} Exsúrge, Christe, ádjuva nos (allelúja).\\
\textit{\scshape ℟.} Et líbera nos propter nomen tuum (allelúja).
}
\rub{All kneeling,}
\englat{
\textit{\scshape ℣.} Lord, have mercy upon us.\\
\textit{\scshape ℟.} Christ, have mercy upon us.\\
\textit{\scshape ℣.} Lord, have mercy upon us.\\
\par\noindent
\lett{O}{ur} Father, who art in heaven, Hallowed be thy Name. Thy kingdom come. Thy will be done on earth, As it is in heaven. Give us this day our daily bread. And forgive us our trespasses, As we forgive those who trespass against us. And lead us not into temptation; But deliver us from evil. Amen.\\
\par\noindent
\textit{\scshape ℣.} O let my mouth be filled with thy praise.\\
\textit{\scshape ℟.} That I may sing of thy glory and honour all the day long.\\
\textit{\scshape ℣.} Turn thy face from my sins, O Lord.\\
\textit{\scshape ℟.} And put out all my misdeeds.\\
\textit{\scshape ℣.} Make me a clean heart, O God.\\
\textit{\scshape ℟.} And renew a right spirit within me.\\
\textit{\scshape ℣.} Cast me not away from thy presence.\\
\textit{\scshape ℟.} And take not thy Holy Spirit from me.\\
\textit{\scshape ℣.} O give me the comfort of thy help again.\\
\textit{\scshape ℟.} And stablish me with thy free Spirit.\\
\par\noindent
I confess to God Almighty, to blessed Mary ever-Virgin, to blessed Michael the Archangel, to blessed John Baptist, to the holy Apostles Peter and Paul, and to all the Saints that I have sinned exceedingly in thought, word, and deed, by my fault, by my own fault, by my own most grievous fault. Therefore I beg blessed Mary ever-Virgin, blessed Michael the Archangel, blessed John Baptist, the holy Apostles Peter and Paul, and all the Saints, to pray for me to the Lord our God.
}
{
\textit{\scshape ℣.} Kýrie, eléison.\\
\textit{\scshape ℟.} Christe, eléison.\\
\textit{\scshape ℣.} Kýrie, eléison.\\
\par\noindent
\lett{P}{ater} noster, qui es in c{\ae}lis, sanctificétur nomen tuum: advéniat regnum tuum: fiat volúntas tua, sicut in c{\ae}lo et in terra. Panem nostrum quotidiánum da nobis hódie: et dimítte nobis débita nostra, sicut et nos dimíttimus debitóribus nostris: et ne nos indúcas in tentatiónem: sed líbera nos a malo. Amen.\\
\par\noindent
\textit{\scshape ℣.} Repleátur os meum laude.\\
\textit{\scshape ℟.} Ut cántem glóriam tuam, tota die magnitúdinem tuam.\\
\textit{\scshape ℣.} Dómine, avérte fáciem tuam a peccátis meis.\\
\textit{\scshape ℟.} Et omnes iniquitátes meas dele.\\
\textit{\scshape ℣.} Cor mundum crea in me, Deus.\\
\textit{\scshape ℟.} Et spíritum rectum ínnova in viscéribus meis.\\
\textit{\scshape ℣.} Ne projícias me a fácie tua.\\
\textit{\scshape ℟.} Et spíritum sanctum tuum ne áuferas a me.\\
\textit{\scshape ℣.} Redde mihi lætítiam salutáris tui.\\
\textit{\scshape ℟.} Et spíritu principáli confírma me.\\
\par\noindent
Confíteor Deo omnipoténti, beátæ Maríæ semper Vírgini, beáto Michaéli Archángelo, beáto Joánni Baptístæ, sanctis Apóstolis Petro et Paulo, et ómnibus Sanctis, quia peccávi nimis, cogitatióne, verbo et ópere: mea culpa, mea culpa, mea máxima culpa. Ídeo precor beátam Maríam semper Vírginem, beátum Michaélem Archángelum, beátum Joánnem Baptístam, sanctos Apóstolos Petrum et Paulum, et omnes Sanctos, oráre pro me ad Dóminum Deum nostrum.
}
\textit{\scshape ℣.} God Almighty have mercy upon us, forgive us all our sins and deliver us from all evil, confirm and strengthen us in all goodness, and bring us to life everlasting. Amen.
\rub{If the Minister is a Priest, he declares,}
\textit{\scshape ℣.} May the Almighty and merciful Lord grant unto you {\ding{64}} pardon and remission of all your sins, time for amendment of life, and the grace and consolation of the Holy Spirit.\\
\textit{\scshape ℟.} Amen.
\englat{
\textit{\scshape ℣.} Wilt thou not turn again and quicken us.\\
\textit{\scshape ℟.} That thy people may rejoice in thee.\\
\textit{\scshape ℣.} O Lord, show thy mercy upon us.\\
\textit{\scshape ℟.} And grant us thy salvation.\\
\textit{\scshape ℣.} Vouchsafe, O Lord, to keep us this day without sin.\\
\textit{\scshape ℟.} O Lord, have mercy upon us, have mercy upon us.\\
\textit{\scshape ℣.} O Lord, hear our prayer.\\
\textit{\scshape ℟.} And let our cry come unto thee.\\
\textit{\scshape ℣.} The Lord be with you.\\
\textit{\scshape ℟.} And with thy spirit.\\
\textit{\scshape ℣.} Let us pray.
}{
\textit{\scshape ℣.} Deus, tu convérsus vivificábis nos.\\
\textit{\scshape ℟.} Et plebs tua l{\ae}tábitur in te.\\
\textit{\scshape ℣.} Osténde nobis, Dómine, misericórdiam tuam.\\
\textit{\scshape ℟.} Et salutáre tuum da nobis.\\
\textit{\scshape ℣.} Dignáre, Dómine, die isto sine peccáto nos custodíre.\\
\textit{\scshape ℟.} Miserére nostri, Dómine. Miserére nostri.\\
\textit{\scshape ℣.} Dómine, exáudi oratiónem meam.\\
\textit{\scshape ℟.} Et clamor meus ad te véniat.\\
\textit{\scshape ℣.} Dóminus vobíscum.\\
\textit{\scshape ℟.} Et cum spiritu tuo.\\
\textit{\scshape ℣.} Orémus.
}
\rub{The Dominus Vobiscum is only said if the Minister is in Major Orders.}
\lett{O}{Almighty} Lord and everlasting God, vouchsafe, we beseech thee, to direct, sanctify, and govern, both our hearts and bodies in the ways of thy laws and in the works of thy commandments: that through thy most mighty protection, both here and ever, we may be preserved in body and soul; through our Lord and Saviour Jesus Christ.\\
\textit{\scshape ℟.} Amen.\\
\englat{
\textit{\scshape ℣.} The Lord be with you. (O Lord, hear my prayer.)\\
\textit{\scshape ℟.} And with thy spirit. (And let my cry come unto thee.)\\
\textit{\scshape ℣.} Let us bless the Lord.\\
\textit{\scshape ℟.} Thanks be to God.
}
{
	\textit{\scshape ℣.} Dóminus vobíscum. (Dómine, exáudi oratiónem meam.)\\
	\textit{\scshape ℟.} Et cum spiritu tuo. (Et clamor meus ad te véniat.)\\
	\textit{\scshape ℣.} Benedicámus Dómino.\\
	\textit{\scshape ℟.} Deo grátias.
}
\par\noindent
\lett{T}{he} Lord {\ding{64}} bless us, preserve us from all evil, and bring us to everlasting life; and may the souls of the faithful departed, throguht he mercy of God, rest in peace.
\textit{\scshape ℟.} Deo grátias.