\section{The Hour of Sext}
\fancyhead[RE,LO]{The Hour of Sext}
\subsection{Introduction}
\englat
{%
\textit{\scshape ℣.} O God, {\ding{64}} make speed to save us.\\
\textit{\scshape ℟.} O Lord, make haste to help us.\\
\textit{\scshape ℣.} Glory be to the Father, and to the Son, and to the Holy Ghost.\\
\textit{\scshape ℟.} As it was in the beginning, is now, and ever shall be, world without end. Amen.\\
\textit{\scshape ℣.} Praise ye the Lord.\\
\textit{\scshape ℟.} The Lord's Name be praised.\\
}
{%
\textit{\scshape ℣.} Deus, {\ding{64}} in adjutorium nostrum intende.\\
\textit{\scshape ℟.} Domine, ad adjuvandum nos festina.\\
\textit{\scshape ℣.} Glória Patri, et Fílio, * et Spirítui Sancto:\\
\textit{\scshape ℟.} Sicut erat in princípio, et nunc, et semper, * et in sǽcula s{\ae}culórum. Amen.\\
\textit{\scshape ℣.} Laudate Dominum.\\
\textit{\scshape ℟.} Sit Nomen Domini Benedictum.\\
}
\subsection{Rector potens, verax Deus}
\englat{%
O God of truth, O Lord of might,\\
Who orderest time and change aright,\\
And sendest the early morning ray,\\
And lightest the glow of perfect day;\\
\par
\noindent
Extinguish thou each sinful fire,\\
And banish every ill desire:\\
And while thou keepest the body whole,\\
Shed forth thy peace upon the soul.\\
\par
\noindent
O Father, that we ask be done\\
Through Jesus Christ, thine only Son,\\
Who, with the Holy Ghost and thee,\\
Shall live and reign eternally. Amen.
}{%
Rector potens, verax Deus,\\
Qui témperas rerum vices,\\
Splendóre mane illúminas,\\
Et ígnibus merídiem:\\
\par\noindent
Exstíngue flammas lítium,\\
Aufer calórem nóxium,\\
Confer salútem córporum,\\
Verámque pacem córdium.\\
\par\noindent
Pr{\ae}sta, Pater piíssime,\\
Patríque compar Únice,\\
Cum Spíritu Paráclito\\
Regnans per omne sǽculum.\\
Amen.
}
\subsection{The Psalms}
\inv{Ant. throughout the Year:} Alleluia. *\\
\inv{Ant. in Advent:} Lo, now cometh. *\\
\inv{Ant. from Christmas until Candlemas.} Whom beheld ye. *\\
\inv{Ant. in Pre-Lent and Lent.} They that fear the Lord. *\\
\inv{Ant. in Eastertide.} Alleluia. *
\psa{Psalm 119:VIII \textit{Portio mea, Domine.}}
\rub{Sunday, Monday, Wednesday, Friday}
\lett{T}{hou} art my portion, O \textsc{Lord}; * I have promised to keep thy law.
\par
    58 I made my humble petition in thy presence with my whole heart; * O be merciful unto me, according to thy word.
\par
    59 I called mine own ways to remembrance, * and turned my feet unto thy testimonies.
\par
    60 I made haste, and prolonged not the time, * to keep thy commandments.
\par
    61 The snares of the ungodly have compassed me about; * but I have not forgotten thy law.
\par
    62 At midnight I will rise to give thanks unto thee, * because of thy righteous judgments.
\par
    63 I am a companion of all them that fear thee, * and keep thy commandments.
\par
    64 The earth, O \textsc{Lord}, is full of thy mercy: * O teach me thy statutes.
\psa{Psalm 119:IX \textit{Bonitatem fecisti.}}
\rub{Sunday, Monday, Wednesday, Friday}
\lett{O}{Lord}, thou hast dealt graciously with thy servant, * according unto thy word.
\par
    66 O teach me true understanding and knowledge; * for I have believed thy commandments.
\par
    67 Before I was troubled, I went wrong; * but now have I kept thy word.
\par
    68 Thou art good and gracious; * O teach me thy statutes.
\par
    69 The proud have imagined a lie against me; * but I will keep thy commandments with my whole heart.
\par
    70 Their heart is as fat as brawn; * but my delight hath been in thy law.
\par
    71 It is good for me that I have been in trouble; * that I may learn thy statutes.
\par
    72 The law of thy mouth is dearer unto me * than thousands of gold and silver.
\psa{Psalm 119:X \textit{Manus tuae fecerunt me.}}
\rub{Sunday, Monday, Wednesday, Friday}
\lett{T}{hy} hands have made me and fashioned me: * O give me understanding, that I may learn thy commandments.
\par
    74 They that fear thee will be glad when they see me; * because I have put my trust in thy word.
\par
    75 I know, O \textsc{Lord}, that thy judgments are right, * and that thou of very faithfulness hast caused me to be troubled.
\par
    76 O let thy merciful kindness be my comfort, * according to thy word unto thy servant.
\par
    77 O let thy loving mercies come unto me, that I may live; * for thy law is my delight.
\par
    78 Let the proud be confounded, for they go wickedly about to destroy me; * but I will be occupied in thy commandments.
\par
    79 Let such as fear thee, and have known thy testimonies, * be turned unto me.
\par
    80 O let my heart be sound in thy statutes, * that I be not ashamed.
\psa{Psalm 119:XXI \textit{Principes persecuti sunt.}}
\rub{Tuesday, Thursday, Saturday}
\lett{P}{rinces} have persecuted me without a cause; * but my heart standeth in awe of thy word.
\par
    162 I am as glad of thy word, * as one that findeth great spoils.
\par
    163 As for lies, I hate and abhor them; * but thy law do I love.
\par
    164 Seven times a day do I praise thee; * because of thy righteous judgments.
\par
    165 Great is the peace that they have who love thy law; * and they have none occasion of stumbling.
\par
    166 \textsc{Lord}, I have looked for thy saving health, * and done after thy commandments.
\par
    167 My soul hath kept thy testimonies, * and loved them exceedingly.
\par
    168 I have kept thy commandments and testimonies; * for all my ways are before thee.
\psa{Psalm 119:XXII \textit{Appropinquet deprecatio.}}
\rub{Tuesday, Thursday, Saturday}
\lett{L}{et} my complaint come before thee, O \textsc{Lord}; * give me understanding according to thy word.
\par
    170 Let my supplication come before thee; * deliver me according to thy word.
\par
    171 My lips shall speak of thy praise, * when thou hast taught me thy statutes.
\par
    172 Yea, my tongue shall sing of thy word; * for all thy commandments are righteous.
\par
    173 Let thine hand help me; * for I have chosen thy commandments.
\par
    174 I have longed for thy saving health, O \textsc{Lord}; * and in thy law is my delight.
\par
    175 O let my soul live, and it shall praise thee; * and thy judgments shall help me.
\par
    176 I have gone astray like a sheep that is lost; * O seek thy servant, for I do not forget thy commandments. 
\psa{Psalm 120 \textit{Ad Dominum.}}
\rub{Tuesday, Thursday, Saturday}
\lett{W}{hen} I was in trouble, I called upon the \textsc{Lord}, * and he heard me.
\par
    2 Deliver my soul, O \textsc{Lord}, from lying lips, * and from a deceitful tongue.
\par
    3 What reward shall be given or done unto thee, thou false tongue? * even mighty and sharp arrows, with hot burning coals.
\par
    4 Woe is me, that I am constrained to dwell with Meshech, * and to have my habitation among the tents of Kedar!
\par
    5 My soul hath long dwelt among them * that are enemies unto peace.
\par
    6 I labour for peace; but when I speak unto them thereof, * they make them ready to battle.\\
\par\noindent
\inv{Ant. throughout the Year:} Alleluia. * O Give thanks unto the Lord, because his mercy endureth for ever. Alleluia.\\
\inv{Ant. in Advent:} Lo, now cometh * the fulness of the time, wherein God sent forth his Son into the world.\\
\inv{Ant. from Christmas until Candlemas.} Whom beheld ye, * O shepherds, tell us? Declare to us the tidings; on earth, who hath appeared? We beheld an infant, the new-born Lord and Saviour, amidst a choir of angels, alleluia, alleluia.\\
\inv{Ant. in Pre-Lent and Lent.} They that fear the Lord * confess that his mercy endureth for ever.\\
\inv{Ant. in Eastertide.} Alleluia, * alleluia, alleluia, alleluia.
\subsection{The Chapter}
\rub{Then the following, or another suitable Chapter, shall be read.}
Bear ye one another's burdens, and so fulfil the law of Christ. \vr{Gal. 6:2}\\
\textit{\scshape ℟.} Thanks be to God.\\
\par\noindent
\textit{\scshape ℣.} O Lord, thy Word endureth for ever in heaven (alleluia, alleluia).\\
\textit{\scshape ℟.} O Lord, thy Word endureth for ever in heaven (alleluia, alleluia).\\
\textit{\scshape ℣.} Thy truth also remaineth from one generation to another.\\
\textit{\scshape ℟.} For ever in heaven. \inrub{Replaced in Eastertide with ``alleluia, alleluia.''}\\
\textit{\scshape ℣.} Glory be to the Father, and to the Son, and to the Holy Ghost.\\
\textit{\scshape ℟.} O Lord, thy Word endureth for ever in heaven (alleluia, alleluia).\\
\par\noindent
\textit{\scshape ℣.} The Lord is my Shepherd, therefore can I lack nothing (alleluia).\\
\textit{\scshape ℟.} He shall feed me in a green pasture (alleluia).
\rub{All kneeling:}
\englat{%
\textit{\scshape ℣.} Lord, have mercy upon us.\\
\textit{\scshape ℟.} Christ, have mercy upon us.\\
\textit{\scshape ℣.} Lord, have mercy upon us.\\
\lett{O}{ur} Father, who art in heaven, Hallowed be thy Name. Thy kingdom come. Thy will be done on earth, As it is in heaven. Give us this day our daily bread. And forgive us our trespasses, As we forgive those who trespass against us. And lead us not into temptation; But deliver us from evil. Amen.
}{%
\textit{\scshape ℣.} Kýrie, eléison.\\
\textit{\scshape ℟.} Christe, eléison.\\
\textit{\scshape ℣.} Kýrie, eléison.\\
\lett{P}{ater} noster, qui es in c{\ae}lis, sanctificétur nomen tuum: advéniat regnum tuum: fiat volúntas tua, sicut in c{\ae}lo et in terra. Panem nostrum quotidiánum da nobis hódie: et dimítte nobis débita nostra, sicut et nos dimíttimus debitóribus nostris: et ne nos indúcas in tentatiónem: sed líbera nos a malo. Amen.}
\vspace{1ex}
\textit{\scshape ℣.} Turn us again, O Lord God of Hosts.\\
\textit{\scshape ℟.} Show the light of thy countenance, and we shall be whole.\\
\textit{\scshape ℣.} O Christ, arise, help us.\\
\textit{\scshape ℟.} And deliver us for thy Name's sake.\\
\textit{\scshape ℣.} O Lord, hear our prayer.\\
\textit{\scshape ℟.} And let our cry come unto thee.\\
\textit{\scshape ℣.} The Lord be with you.\\
\textit{\scshape ℟.} And with thy spirit.\\
\textit{\scshape ℣.} Let us pray.
\rub{The Dominus Vobiscum is only said if the Minister is in Major Orders.}
\subsection{The Collect}
\lett{O}{Lord} Jesus Christ, who at the sixth hour of the day didst ascend the Cross for the redemption of the world, and didst shed thy precious Blood for the remission of our sins; grant, we humbly beseech thee, that by the merits of thy Passion and Wounds, we may, after our death, enter into the gates of Paradise with joy. Who liveth and reigneth with the Father and the Holy Ghost, ever one God, world without end.\\
\textit{\scshape ℟.} Amen.\\
\englat{
\textit{\scshape ℣.} The Lord be with you. (O Lord, hear my prayer.)\\
\textit{\scshape ℟.} And with thy spirit. (And let my cry come unto thee.)\\
\textit{\scshape ℣.} Let us bless the Lord.\\
\textit{\scshape ℟.} Thanks be to God.\\
\textit{\scshape ℣.} May the souls {\ding{64}} of the faithful departed, through the mercy of God, rest in peace.\\
\textit{\scshape ℟.} Amen.
}
{
	\textit{\scshape ℣.} Dóminus vobíscum. (Dómine, exáudi oratiónem meam.)\\
	\textit{\scshape ℟.} Et cum spiritu tuo. (Et clamor meus ad te véniat.)\\
	\textit{\scshape ℣.} Benedicámus Dómino.\\
	\textit{\scshape ℟.} Deo grátias.\\
	\textit{\scshape ℣.} Fidélium ánim{\ae} {\ding{64}} per misericórdiam Dei requiéscant in pace.\\
	\textit{\scshape ℟.} Amen.
}