\section{The Hour of Terce}
\fancyhead[RE,LO]{The Hour of Terce}
\subsection{Introduction}
\englat
{
\textit{\scshape ℣.} O God, {\ding{64}} make speed to save us.\\
\textit{\scshape ℟.} O Lord, make haste to help us.\\
\textit{\scshape ℣.} Glory be to the Father, and to the Son, and to the Holy Ghost.\\
\textit{\scshape ℟.} As it was in the beginning, is now, and ever shall be, world without end. Amen.\\
\textit{\scshape ℣.} Praise ye the Lord.\\
\textit{\scshape ℟.} The Lord's Name be praised.\\
}
{
\textit{\scshape ℣.} Deus, {\ding{64}} in adjutorium nostrum intende.\\
\textit{\scshape ℟.} Domine, ad adjuvandum nos festina.\\
\textit{\scshape ℣.} Glória Patri, et Fílio, * et Spirítui Sancto:\\
\textit{\scshape ℟.} Sicut erat in princípio, et nunc, et semper, * et in sǽcula s{\ae}culórum. Amen.\\
\textit{\scshape ℣.} Laudate Dominum.\\
\textit{\scshape ℟.} Sit Nomen Domini Benedictum.\\
}
\subsection{Nunc, Sancte, nobis, Spiritus}
\englat{
Come, Holy Ghost, who ever One\\
Art with the Father and the Son;\\
Come, Holy Ghost, our souls possess\\
With thy full flood of holiness.\\
\par
\noindent
By every power, by heart and tongue,\\
By act and deed, thy praise be sung:\\
Inflame with perfect love each sense,\\
That others' souls may kindle thence.\\
\par
\noindent
O Father, that we ask be done\\
Through Jesus Christ, thine only Son,\\
Who, with the Holy Ghost and thee,\\
Shall live and reign eternally. Amen.
}{
Nunc, Sancte, nobis, Spíritus,\\
Unum Patri cum Fílio,\\
Dignáre promptus íngeri\\
Nostro refúsus péctori.\\
\par\noindent
Os, lingua, mens, sensus, vigor\\
Confessiónem pérsonent.\\
Flamméscat igne cáritas,\\
Accéndat ardor próximos.\\
\par\noindent
Jesu, tibi sit glória,\\
Qui natus es de Vírgine,\\
Cum Patre et almo Spíritu,\\
In sempitérna sǽcula.\\
Amen.
}
\subsection{The Psalms}
\inv{Ant. throughout the Year:} Alleluia. *\\
\inv{Ant. in Advent:} Lo, now cometh. *\\
\inv{Ant. from Christmas until Candlemas.} Whom beheld ye. *\\
\inv{Ant. in Pre-Lent and Lent.} They that fear the Lord. *\\
\inv{Ant. in Eastertide.} Alleluia. *
\subsubsection{Psalm 119:V \textit{Legem pone.}}
\rub{Sunday, Monday, Wednesday, Friday}
\lett{T}{each} me, O \textsc{Lord}, the way of thy statutes, * and I shall keep it unto the end.
\par
    34 Give me understanding, and I shall keep thy law; * yea, I shall keep it with my whole heart.
\par
    35 Make me to go in the path of thy commandments; * for therein is my desire.
\par
    36 Incline my heart unto thy testimonies, * and not to covetousness.
\par
    37 O turn away mine eyes, lest they behold vanity; * and quicken thou me in thy way.
\par
    38 O stablish thy word in thy servant, * that I may fear thee.
\par
    39 Take away the rebuke that I am afraid of; * for thy judgments are good.
\par
    40 Behold, my delight is in thy commandments; * O quicken me in thy righteousness.
    \subsubsection{Psalm 119:VI \textit{Et veniat super me.}}
\rub{Sunday, Monday, Wednesday, Friday}
\lett{L}{et} thy loving mercy come also unto me, O \textsc{Lord}, * even thy salvation, according unto thy word.
\par
    42 So shall I make answer unto my blasphemers; * for my trust is in thy word.
\par
    43 O take not the word of thy truth utterly out of my mouth; * for my hope is in thy judgments.
\par
    44 So shall I alway keep thy law; * yea, for ever and ever.
\par
    45 And I will walk at liberty; * for I seek thy commandments.
\par
    46 I will speak of thy testimonies also, even before kings, * and will not be ashamed.
\par
    47 And my delight shall be in thy commandments, * which I have loved.
\par
    48 My hands also will I lift up unto thy commandments, which I have loved; * and my study shall be in thy statutes.
\subsubsection{Psalm VII \textit{Memor esto verbi tui.}}
\rub{Sunday, Monday, Wednesday, Friday}
\lett{O}{Think} upon thy servant, as concerning thy word, * wherein thou hast caused me to put my trust.
\par
    50 The same is my comfort in my trouble; * for thy word hath quickened me.
\par
    51 The proud have had me exceedingly in derision; * yet have I not shrinked from thy law.
\par
    52 For I remembered thine everlasting judgments, O \textsc{Lord}, * and received comfort.
\par
    53 I am horribly afraid, * for the ungodly that forsake thy law.
\par
    54 Thy statutes have been my songs, * in the house of my pilgrimage.
\par
    55 I have thought upon thy Name, O \textsc{Lord}, in the night season, * and have kept thy law.
\par
    56 This I had, * because I kept thy commandments.
\subsubsection{Psalm 119:XVIII \textit{Justus es, Domine.}}
\rub{Tuesday, Thursday, Saturday}
\lett{R}{ighteous} art thou, O \textsc{Lord}; * and true are thy judgments.
\par
    138 The testimonies that thou hast commanded * are exceeding righteous and true.
\par
    139 My zeal hath even consumed me; * because mine enemies have forgotten thy words.
\par
    140 Thy word is tried to the uttermost, * and thy servant loveth it.
\par
    141 I am small and of no reputation; * yet do I not forget thy commandments.
\par
    142 Thy righteousness is an everlasting righteousness, * and thy law is the truth.
\par
    143 Trouble and heaviness have taken hold upon me; * yet is my delight in thy commandments.
\par
    144 The righteousness of thy testimonies is everlasting: * O grant me understanding, and I shall live.
\subsubsection{Psalm 119:XIX \textit{Clamavi in toto corde meo.}}
\rub{Tuesday, Thursday, Saturday}
\lett{I}{Call} with my whole heart; * hear me, O \textsc{Lord}; I will keep thy statutes.
\par
    146 Yea, even unto thee do I call; * help me, and I shall keep thy testimonies.
\par
    147 Early in the morning do I cry unto thee; * for in thy word is my trust.
\par
    148 Mine eyes prevent the night watches; * that I might be occupied in thy word.
\par
    149 Hear my voice, O \textsc{Lord}, according unto thy lovingkindness; * quicken me, according to thy judgments.
\par
    150 They draw nigh that of malice persecute me, * and are far from thy law.
\par
    151 Be thou nigh at hand, O \textsc{Lord}; * for all thy commandments are true.
\par
    152 As concerning thy testimonies, I have known long since, * that thou hast grounded them for ever.
\subsubsection{Psalm 119:XX \textit{Vide humilitatem.}}
\rub{Tuesday, Thursday, Saturday}
\lett{O}{Consider} mine adversity, and deliver me, * for I do not forget thy law.
\par
    154 Avenge thou my cause, and deliver me; * quicken me according to thy word.
\par
    155 Health is far from the ungodly; * for they regard not thy statutes.
\par
    156 Great is thy mercy, O \textsc{Lord}; * quicken me, as thou art wont.
\par
    157 Many there are that trouble me, and persecute me; * yet do I not swerve from thy testimonies.
\par
    158 It grieveth me when I see the transgressors; * because they keep not thy law.
\par
    159 Consider, O \textsc{Lord}, how I love thy commandments; * O quicken me, according to thy loving-kindness.
\par
    160 Thy word is true from everlasting; * all the judgments of thy righteousness endure for evermore.
\par\noindent
\inv{Ant. throughout the Year:} Alleluia. * O Give thanks unto the Lord, because his mercy endureth for ever. Alleluia.
\inv{Ant. in Advent:} Lo, now cometh * the fulness of the time, wherein God sent forth his Son into the world.
\inv{Ant. from Christmas until Candlemas.} Whom beheld ye, * O shepherds, tell us? Declare to us the tidings; on earth, who hath appeared? We beheld an infant, the new-born Lord and Saviour, amidst a choir of angels, alleluia, alleluia.
\inv{Ant. in Pre-Lent and Lent.} They that fear the Lord * confess that his mercy endureth for ever.
\inv{Ant. in Eastertide.} Alleluia, * alleluia, alleluia, alleluia.
\subsection{The Chapter}
\rub{Then the following, or another suitable Chapter, shall be read.}
O the depth of the riches both of the wisdom and knowledge of God! How unsearchable are his judgements, and his ways past finding out. \vr{Rom. 11:33}\\
\textit{\scshape ℟.} Thanks be to God.\\
\par\noindent
\textit{\scshape ℣.} Incline my heart, O God, unto thy testimonies (alleluia, alleluia).\\
\textit{\scshape ℟.} Incline my heart, O God, unto thy testimonies (alleluia, alleluia).\\
\textit{\scshape ℣.} Turn away mine eyes, lest they behold vanity.\\
\textit{\scshape ℟.} Unto thy testimonies (alleluia, alleluia).\\
\textit{\scshape ℣.} Glory be to the Father, and to the Son, and to the Holy Ghost.\\
\textit{\scshape ℟.} Incline my heart, O God, unto thy testimonies (alleluia, alleluia).\\
\par\noindent
\textit{\scshape ℣.} I said, Lord, be merciful unto me (alleluia).\\
\textit{\scshape ℟.} For I have sinned against thee (alleluia).
\rub{All kneeling:}
\englat{
\textit{\scshape ℣.} Lord, have mercy upon us.\\
\textit{\scshape ℟.} Christ, have mercy upon us.\\
\textit{\scshape ℣.} Lord, have mercy upon us.}\\
\lett{O}{ur} Father, who art in heaven, Hallowed be thy Name. Thy kingdom come. Thy will be done on earth, As it is in heaven. Give us this day our daily bread. And forgive us our trespasses, As we forgive those who trespass against us. And lead us not into temptation; But deliver us from evil. Amen.
{
\textit{\scshape ℣.} Kýrie, eléison.\\
\textit{\scshape ℟.} Christe, eléison.\\
\textit{\scshape ℣.} Kýrie, eléison.\\
\lett{P}{ater} noster, qui es in c{\ae}lis, sanctificétur nomen tuum: advéniat regnum tuum: fiat volúntas tua, sicut in c{\ae}lo et in terra. Panem nostrum quotidiánum da nobis hódie: et dimítte nobis débita nostra, sicut et nos dimíttimus debitóribus nostris: et ne nos indúcas in tentatiónem: sed líbera nos a malo. Amen.}
\textit{\scshape ℣.} Send out thy light and thy truth, that they may lead me.\\
\textit{\scshape ℟.} And bring me unto thy holy hill, and to thy dwelling.\\
\textit{\scshape ℣.} O let us live.\\
\textit{\scshape ℟.} And we shall call upon thy name.\\
\textit{\scshape ℣.} O Lord, hear our prayer.\\
\textit{\scshape ℟.} And let our cry come unto thee.\\
\textit{\scshape ℣.} The Lord be with you.\\
\textit{\scshape ℟.} And with thy spirit.\\
\textit{\scshape ℣.} Let us pray.
\rub{The Dominus Vobiscum is only said if the Minister is in Major Orders.}
\subsection{The Collect}
\lett{H}{elp} us this day, O God, to serve thee devoutly, and the world busily. May we do our work wisely, give succour secretly, go to our meat appetitely, sit thereat discreetly, arise temperately, please our friends duly, go to our bread merrily, and sleep surely; for the joy of our Lord, Jesus Christ.\\
\textit{\scshape ℟.} Amen.\\
\englat{
\textit{\scshape ℣.} The Lord be with you. (O Lord, hear my prayer.)\\
\textit{\scshape ℟.} And with thy spirit. (And let my cry come unto thee.)\\
\textit{\scshape ℣.} Let us bless the Lord.\\
\textit{\scshape ℟.} Thanks be to God.\\
\textit{\scshape ℣.} May the souls {\ding{64}} of the faithful departed, through the mercy of God, rest in peace.\\
\textit{\scshape ℟.} Amen.
}
{
	\textit{\scshape ℣.} Dóminus vobíscum. (Dómine, exáudi oratiónem meam.)\\
	\textit{\scshape ℟.} Et cum spiritu tuo. (Et clamor meus ad te véniat.)\\
	\textit{\scshape ℣.} Benedicámus Dómino.\\
	\textit{\scshape ℟.} Deo grátias.\\
	\textit{\scshape ℣.} Fidélium ánim{\ae} {\ding{64}} per misericórdiam Dei requiéscant in pace.\\
	\textit{\scshape ℟.} Amen.
}