\section{The Liturgy of St. Tikhon}
\fancyhead[RE,LO]{The Liturgy of St. Tikhon}
\rub{It is a laudable custom to begin the liturgy with the Lord's Prayer.}
\subsection{The Collect for Purity}
\lett{A}{lmighty} God, unto whom all hearts be open, all desires known, and from whom no secrets are hid; Cleanse the thoughts of our hearts by the inspiration of thy Holy Spirit, that we may perfectly love thee, and worthily magnify thy holy name; through Christ our Lord. Amen.
\subsection{The Summary of the Law}
\rub{Instead of the Summary of the Law may be said the Decalogue (p. \ref{decalogue})}%
\noindent%
Hear what our Lord Jesus Christ saith.\\
\textsc{Thou} shalt love the Lord thy God with all thy heart, and with all thy soul, and with all thy mind. This is the first and great commandment. And the second is like unto it; Thou shalt love thy neighbour as thyself. On these two commandments hang all the Law and the Prophets.
\subsection{The Kyrie}
\elcoldent{℣. Lord, have mercy upon us.}{℣. Kyrie, eléison.}
\elcoldent{℟. Lord, have mercy upon us.}{℟. Kyrie, eléison.}
\elcoldent{℣. Lord, have mercy upon us.}{℣. Kyrie, eléison.}
\elcoldent{℟. Christ, have mercy upon us.}{℟. Christe, eléison.}
\elcoldent{℣. Christ, have mercy upon us.}{℣. Christe, eléison.}
\elcoldent{℟. Christ, have mercy upon us.}{℟. Christe, eléison.}
\elcoldent{℣. Lord, have mercy upon us.}{℣. Kyrie, eléison.}
\elcoldent{℟. Lord, have mercy upon us.}{℟. Kyrie, eléison.}
\elcoldent{℣. Lord, have mercy upon us.}{℣. Kyrie, eléison.}
\subsection{Gloria in Excelsis}
\elcol{\lett{G}{lory} be to God on high, and on earth peace, good will towards men. We praise thee, we bless thee, we worship thee, we glorify thee, we give thanks to thee for thy great glory, O Lord God, heavenly King, God the Father Almighty.\\
    O Lord, the only-begotten Son, Jesus Christ; O Lord God, Lamb of God, Son of the Father, that takest away the sins of the world, have mercy upon us. Thou that takest away the sins of the world, receive our prayer. Thou that sittest at the right hand of God the Father, have mercy upon us.\\
    For thou only art holy; thou only art the Lord; thou only, O Christ, with the Holy Ghost, {\ding{64}} art most high in the glory {\ding{64}} of God the Father. Amen.\\
    }
    {\lett{G}{lória} in excélsis Deo. Et in terra pax homínibus bon{\ae} voluntátis. Laudámus te. Benedícimus te. Adorámus te. Glorificámus te. Grátias ágimus tibi propter magnam glóriam tuam. Dómine Deus, Rex c{\ae}léstis, Deus Pater omnípotens. \\
    Dómine Fili unigénite, Jesu Christe. Dómine Deus, Agnus Dei, Fílius Patris. Qui tollis peccáta mundi, miserére nobis. Qui tollis peccáta mundi, súscipe deprecatiónem nostram. Qui sedes ad déxteram Patris, miserére nobis. Quóniam tu solus Sanctus. Tu solus Dóminus. Tu solus Altíssimus, Jesu Christe. Cum Sancto Spíritu {\ding{64}} in glória Dei Patris. Amen.\\
    }
\elcol{℣.The Lord be with you.}{℣. Dóminus vobíscum.}
\elcol{℟. And with thy spirit.}{℟. Et cum spiritu tuo.}
\elcol{℣. Let us pray.}{℣. Orémus.}
\subsection{The Collect}
\rub{The Priest prays the collect(s) for the day.}
\subsection{The Epistle}
\rub{The proper epistle for the day shall be read by the Subdeacon.\\At the end of the Epistle the People respond,}
\elcol{℟. Thanks be to God.}{℟. Deo grátias.}
\subsection{The Gospel}
\rub{The Gospel shall be read by the Deacon.\\At the end of the Gospel the People respond,}
\elcol{℟. Praise be to thee, O Christ.}{℟. Laus tibi, Christe.}
\subsection{The Nicene Creed}
\elcol{
\lett{I}{Believe} in one God the Father Almighty, Maker of heaven and earth, And of all things visible and invisible:\par
    And in one Lord Jesus Christ, the only-begotten Son of God; Begotten of his Father before all worlds, God of God, Light of Light, Very God of very God; Begotten, not made; Being of one substance with the Father; By whom all things were made: Who for us men and for our salvation came down from heaven, \inrub{Everyone genuflects.} And was incarnate by the Holy Ghost of the Virgin Mary, And was made man: \inrub{Everyone rises.} And was crucified also for us under Pontius Pilate; He suffered and was buried: And the third day he rose again according to the Scriptures: And ascended into heaven, And sitteth on the right hand of the Father: And he shall come again, with glory, to judge both the quick and the dead; Whose kingdom shall have no end.\par
    And I believe in the Holy Ghost, The Lord, and Giver of Life, Who proceedeth from the Father; Who with the Father and the Son together is worshiped and glorified; Who spake by the Prophets: And I believe one, holy, catholic, and apostolic Church: I acknowledge one Baptism for the remission of sins: And I look for the Resurrection of the dead: {\ding{64}} And the Life of the world to come. Amen.}
    {
    \lett{C}{redo} in unum Deum, Patrem omnipoténtem, factórem c{\ae}li et terr{\ae}, visibílium ómnium et invisibílium.\par
    Et in unum Dóminum Jesum Christum, Fílium Dei unigénitum. Et ex Patre natum ante ómnia sǽcula. Deum de Deo, lumen de lúmine, Deum verum de Deo vero. Génitum, non factum, consubstantiálem Patri: per quem ómnia facta sunt. Qui propter nos hómines et propter nostram salútem descéndit de c{\ae}lis. \inrub{Everyone genuflects.} Et incarnátus est de Spíritu Sancto ex María Vírgine: Et homo factus est. \inrub{Everyone rises.} Crucifíxus étiam pro nobis: sub Póntio Piláto passus, et sepúltus est. Et resurréxit tértia die, secúndum Scriptúras. Et ascéndit in c{\ae}lum: sedet ad déxteram Patris. Et íterum ventúrus est cum glória judicáre vivos et mórtuos: cujus regni non erit finis.\par
    Et in Spíritum Sanctum, Dóminum et vivificántem: qui ex Patre procédit. Qui cum Patre et Fílio simul adorátur et conglorificátur: qui locútus est per Prophétas. Et unam sanctam cathólicam et apostólicam Ecclésiam. Confíteor unum baptísma in remissiónem peccatórum. Et exspécto resurrectiónem mortuórum. {\ding{64}} Et vitam ventúri s{\ae}culi. Amen.
    }
\rub{After the Creed, a sermon may be given.}
\subsection{The Offertory}
\elcol{
    ℣. The Lord be with you.\par
	℟. And with thy spirit.\par
	℣. Let us pray.
	}{
    ℣. Dóminus vobíscum.\par
	℟. Et cum spiritu tuo.\par
	℣. Orémus.
}
\rub{The priest shall here offer the bread and the wine to God and pray the several secret prayers.}
\elcol{℣. Pray brethren, that this my sacrifice and yours may be acceptable unto God the Father Almighty.}{℣. Oráte, fratres: ut meum ac vestrum sacrifícium acceptábile fiat apud Deum Patrem omnipoténtem.}
\elcol{℟. May the Lord receive this sacrifice at thy hands, to the praise and glory of his name, both to our benefit, and that of all his holy Church.}{℟. Suscípiat Dóminus sacrifícium de mánibus tuis ad laudem et glóriam nominis sui, ad utilitátem quoque nostram, totiúsque Ecclési{\ae} su{\ae} sanct{\ae}.}
\textsc{Let} us pray for the whole state of Christ’s Church.\\
\lett{A}{lmighty} and ever living God, who by thy holy Apostle hast taught us to make prayers, and supplications, and to give thanks for all men: We humbly beseech thee most mercifully to accept (these) our Oblations, and to receive these our prayers, which we offer unto thy Divine Majesty: beseeching thee to inspire continually the Universal Church with the spirit of truth, unity, and concord: And grant that all those who do confess thy holy name may agree in the truth of thy holy Word, and live in unity and godly love. We beseech thee also, so to direct and dispose the hearts of all Christian Rulers, that they may truly and impartially administer justice, to the punishment of wickedness and vice, and to the maintenance of thy true religion and virtue. Give grace, O heavenly Father, to all Bishops and other Ministers, especially \textit{N.} our Patriarch, \textit{N.} our Metropolitan, and to the Holy Synod of \textit{N.}, that they may, both by their life and doctrine, set forth thy true and lively Word, and rightly and duly administer thy holy Sacraments. And to all thy people give thy heavenly grace; and especially to this congregation here present; that, with meek heart and due reverence, they may hear, and receive thy holy Word; truly serving thee in holiness and righteousness all the days of their life. And we most humbly beseech thee, of thy goodness, O Lord, to comfort and succor all those who, in this transitory life, are in trouble, sorrow, need, sickness, or any other adversity (the priest may commemorate specific names). And we also bless thy holy name for all thy servants departed this life in thy faith and fear; beseeching thee to grant them continual growth in thy love and service. And give us grace so to follow the good examples of blessed Mary and all thy Saints, that, through their intercessions, we (with them) may be partakers of thy heavenly kingdom. Grant this, O Father, for Jesus Christ's sake, our only Mediator and Advocate. Amen.
\subsection{The Sursum Corda}
\elcol{℣. The Lord be with you.}{℣. Dóminus vobíscum.}
\elcol{℟. And with thy spirit.}{℟. Et cum spiritu tuo.}
\elcol{℣. Lift up your hearts.}{℣. Sursum corda.}
\elcol{℟. We lift them up unto the Lord.}{℟. Habémus ad Dóminum.}
\elcol{℣. Let us give thanks unto our Lord God.}{℣. Grátias agámus Dómino, Deo nostro.}
\elcol{℟. It is meet and right so to do.}{℟. Dignum et justum est.}
\subsection{Proper Prefaces}\label{prefaces}
\rub{Then shall the Priest turn to the Altar, and say,}
\noindent
\textsc{It} is very meet, right, and our bounden duty, that we should at all times, and in all places, give thanks unto thee, O Lord, Holy Father, Almighty, Everlasting God.
\subsubsection{The Most Holy Trinity}
\rub{On Trinity Sunday, at Votive Masses of the Blessed Trinity, and on all Sundays throughout the year, except in Paschaltide and on Feasts having a proper Preface.}
\textsc{Who,} with thine only-begotten Son, and the Holy Ghost, art one God, one Lord, in Trinity of Persons and in Unity of Substance. For that which we believe of thy glory, O Father, the same we believe of the Son, and of the Holy Ghost, without any difference of inequality.
    Therefore with Angels, etc.
\subsubsection{Advent}
\rub{To be said in Masses of the season of Advent, except on Feasts which have a proper preface.}
\textsc{Through} Jesus Christ our Lord: Whom thou, merciful and faithful, didst promise as the Saviour to the lost race of man; a Saviour whose truth would instruct the ignorant, whose holiness would justify the wicked, and whose power would strengthen the infirm. Now that the time draweth night for the coming of he whom thou shalt send, and the day of our deliverance shineth forth, with faith in thy promises, we rejoice in holy exultation. Therefore with Angels, etc.
\subsubsection{Christmas}
\rub{From the Feast of the Nativity until the Epiphany, exclusive.}
\textsc{Because} thou didst give Jesus Christ, thine only Son, to be born as at this time for us; who, by the operation of the Holy Ghost, was made very man, of the substance of the Virgin Mary his mother; and that without spot of sin, to make us clean from all sin. Therefore with Angels, etc.
\subsubsection{Epiphany}
\rub{On the Feast of the Epiphany and during the Octave.}
\textsc{Through} Jesus Christ our Lord, who, in substance of our mortal flesh, manifested forth his glory; that he might bring us out of darkness into his own glorious light. 
    Therefore with Angels, etc.
\subsubsection{Feasts of Our Lord \& Our Lady}
\rub{On the Purification, Annunciation, Transfiguration, \& Corpus Christi (and during its Octave).}
\textsc{Because} in the Mystery of the Word made flesh, thou hast caused a new light to shine in our hearts, to give the knowledge of thy glory in the face of thy Son, Jesus Christ our Lord.
    Therefore with Angels, etc.
\subsubsection{Lent}
\rub{From Ash Wednesday until Passion Sunday exclusive, except on Feasts which have a proper Preface.}
\textsc{Because} that by our bodily fasting thou dost curb our sinfulness, dost raise our minds from things of earth, dost renew our strength and reward us with manifold blessings, through Jesus Christ our Lord. Therefore, etc.
\subsubsection{Holy Rood}
\rub{In Passiontide, except on Feasts having a proper Preface; on the Feasts of the Holy Rood and of the Precious Blood.}
\textsc{Because} by the Tree of the Cross thou hast wrought the salvation of the race of man; that whence death arose, thence also Life might rise again; and that he who by. a tree was once vanquisher, might also by a Tree be vanquished, through Jesus Christ our Lord. Therefore, etc.
\subsubsection{Easter}
\rub{Upon Easter-day, and seven days after.}
\textsc{But} chiefly are we bound to praise thee for the glorious Resurrection of thy Son Jesus Christ our Lord: for he is the very Paschal Lamb, which was offered for us, and hath taken away the sin of the world; who by his death hath destroyed death, and by his rising to life again hath restored to us everlasting life. 
    Therefore with Angels, etc.
\subsubsection{Ascension}
\rub{Upon Ascension-day, and seven days after.}
\textsc{Through} thy most dearly beloved Son Jesus Christ our Lord; who, after his most glorious Resurrection, manifestly appeared to all his Apostles, and in their sight ascended up into heaven, to prepare a place for us; that where he is, thither we might also ascend, and reign with him in glory. 
    Therefore with Angels, etc.
\subsubsection{Whitsuntide}
\rub{On Pentecost, its Octave, \& Votive Masses of the Holy Ghost}
\textsc{Through} Jesus Christ our Lord; according to whose most true promise, the Holy Ghost came down (as at this time) from heaven, lighting upon the Apostles, to teach them, and to lead them into all truth; giving them boldness with fervent zeal constantly to preach the Gospel unto all nations; whereby we have been brought out of darkness and error into the clear light and true knowledge of thee, and of thy Son Jesus Christ. 
    Therefore with Angels, etc.
\subsubsection{Blessed Sacrament}
\rub{For Votive Masses of the Blessed Sacrament \& Optionally on Corpus Christi}
\textsc{Through} Jesus Christ our Lord: Who, having abolished the empty shadows of fleshly victims, committed his own Body and Blood as a Sacrifice for us: that in every place may be offered to thy Name that clean Oblation, which alone hath been pleasing unto thee. Wherefore, in this mystery of thine inscrutable wisdom and boundless charity, that very thing which he accomplished upon the Cross ceaseth not its wonderful operation, he himself offering, he himself the Victim. Who inviteth us, made one Victim with him, to that sacred Banquet, in which he is received as our Food, the mind wiht grace is filled, and a pledge of future glory is bestowed upon us. Therefore, etc.
\subsubsection{Sacred Heart of Jesus}
\rub{On the Feast \& Octave of the Sacred Heart, except on feasts which have a proper Preface; also in Votive Masses of the Sacred Heart.}
\textsc{Who} didst will that thine only-begotten Son, as he hung upon the Cross, should be pierced by the soldier's spear: that his Heart thus opened, the shrine of heavenly bounty, might pour forth upon us streams of mercy and of grace; and, burning with unceasing love for us, might likewise be a place of rest for the godly, and an abiding refuge of salvation for the penitent. Therefore, etc.
\subsubsection{All Saints}
\rub{Upon All Saints’ Day, and seven days after.}
\textsc{Who,} in the multitude of thy saints, hast compassed us about with so great a cloud of witnesses that we, rejoicing in their fellowship, may run with patience the race that is set before us, and together with them may receive the crown of glory that fadeth not away.
    Therefore with Angels, etc.
\subsubsection{Requiem}
\rub{On All Souls' Day \& in Requiem Masses}
\textsc{Through} Jesus Christ our Lord, in whom hath been manifested forth unto us the hope of a glorious resurrection; giving unto them that bewail the certain condition of their mortality the comfort of thy promise of immortality to come; for thou, O Lord, takest not away but changest the life of thy faithful people, granting unto them at the dissolution of the tabernacle of their earthly pilgrimage a dwelling-place in the heavens. Therefore, etc.
\subsection{The Sanctus}
\lett{T}{herefore} with Angels and Archangels, and with all the company of heaven, we laud and magnify thy glorious Name; evermore praising thee, and saying:
\par
\noindent
\elcol{
\lett{H}{oly, Holy, Holy,} Lord God of hosts, Heaven and earth are full of thy glory: Glory be to thee, O Lord Most High. Blessed {\ding{64}} is he that cometh in the Name of the Lord. Hosanna in the Highest.
}
{
\lett{S}{anctus, Sanctus, Sanctus,} Dóminus, Deus Sábaoth. Pleni sunt c{\ae}li et terra glória tua. Hosánna in excélsis. Benedíctus, {\ding{64}} qui venit in nómine Dómini. Hosánna in excélsis.
}
\subsection{Consecration}
\lett{A}{ll} glory be to thee, Almighty God, our heavenly Father, for that thou, of thy tender mercy, didst give thine only Son Jesus Christ to suffer death upon the Cross for our redemption; who, by his own oblation of himself once offered, made a full, perfect, and sufficient sacrifice, oblation, and satisfaction, for the sins of the whole world; and did institute, and in his holy Gospel command us to continue, a perpetual memory of that his precious death and sacrifice, until his coming again:\\
\par
\noindent
For in the night in which he was betrayed, he took Bread; and when he had given thanks, he brake it, and gave it to his disciples, saying, Take, eat,
\begin{center}
\large{this is my Body, which is given for you; Do this in remembrance of me.}
\end{center}
Likewise, after supper, he took the Cup; and when he had given thanks, he gave it to them, saying, Drink ye all of this;
\begin{center}
\large{for this is my Blood of the New Testament, which is shed for you, and for many, for the remission of sins;}
\end{center}
Do this, as oft as ye shall drink it, in remembrance of me.\\
\par
\lett{W}{herefore} O Lord and heavenly Father, according to the institution of thy dearly beloved Son our Saviour Jesus Christ, we, thy humble servants, do celebrate and make here before thy Divine Majesty, with these thy holy gifts, which we now offer unto thee, the memorial thy Son hath commanded us to make; having in remembrance his blessed passion and precious death, his mighty resurrection and glorious ascension; rendering unto thee most hearty thanks for the innumerable benefits procured unto us by the same.
\subsection{The Epiklesis}
\lett{A}{nd} we most humbly beseech thee, O merciful Father, to hear us: and, of thy almighty goodness, vouchsafe to send down thy Holy Spirit upon these thy gifts and creatures of Bread and Wine, that they may be changed into the Body and Blood of thy most dearly beloved Son. Grant that we, receiving them according to thy Son our Saviour Jesus Christ's holy institution, in remembrance of his Death and Passion, may be partakers of his most blessed Body and Blood.
\par
\lett{A}{nd} we earnestly desire thy fatherly goodness, mercifully to accept this our Sacrifice of Praise and Thanksgiving; most humbly beseeching thee to grant that, by the merits and death of thy Son Jesus Christ, and through faith in his Blood, we, and all thy whole Church, may obtain remission of our sins, and all other benefits of his Passion.
\par
\lett{A}{nd} here we offer and present unto thee, O Lord, our selves, our souls and bodies, to be a reasonable, holy, and living sacrifice unto thee; humbly beseeching thee, that we, and all others who shall be partakers of this Holy Communion, may worthily receive the most precious Body and Blood of thy Son Jesus Christ, be filled with thy grace and heavenly benediction, and made one body with him, that he may dwell in us, and we in him. 
\par
\lett{R}{emember} Lord, also the souls of thy servants and handmaidens, which are gone before us with the mark of faith, and rest in the sleep of peace. \inrub{The Priest now prays for the souls of the dead.} We beseech thee, O Lord, that unto them, and unto all such as rest in Christ, thou wilt grant a place of refreshing, of light, and of peace. And vouchsafe to give unto us some portion and fellowship with thy holy Apostles and Martyrs; with John, Stephen, Matthias, Barnabas, Ignatius, Alexander, Marcellinus, Peter, Felicitas, Perpetua, Agatha, Lucia, Agnes, Cecilia, Anastasia, and with all thy Saints; within whose fellowship we beseech thee to admit us.
\par
\lett{A}{nd} although we are unworthy, through our manifold sins, to offer unto thee any sacrifice; yet we beseech thee to accept this our bounden duty and service; not weighing our merits, but pardoning our offences.
\par
\lett{T}{hrough} Jesus Christ our Lord; By whom, and with whom, in the unity of the Holy Ghost, all honour and glory be unto thee, O Father Almighty, world without end. ℟. Amen.
\subsection{Pater Noster}
\noindent
    Let us pray.\\
    And now, as our Saviour Christ hath taught us, we are bold to say,\par
\lett{O}{ur} Father, who art in heaven, Hallowed be thy Name. Thy kingdom come. Thy will be done, on earth, As it is in heaven. Give us this day our daily bread. And forgive us our trespasses, As we forgive those who trespass against us. And lead us not into temptation, But deliver us from evil. For thine is the kingdom, and the power, and the glory, for ever and ever. Amen.
\subsection{Agnus Dei}
\elcol{\lett{O}{Lamb} of God, that takest away the sins of the world, have mercy upon us.}{\lett{A}{gnus} Dei, qui tollis peccáta mundi: miserére nobis.}
\elcol{O Lamb of God, that takest away the sins of the world, have mercy upon us.}{Agnus Dei, qui tollis peccáta mundi: miserére nobis.}
\elcol{O Lamb of God, that takest away the sins of the world, grant us thy peace.}{Agnus Dei, qui tollis peccáta mundi: dona nobis pacem.}
\rub{In Requiem Masses, ``Have mercy upon us'' is replaced by ``grant them rest'' and ``grant us thy peace'' by ``grant them rest eternal.''}
\subsection{The Confession}
\rub{The Confession may be omitted or said immediately after the Offertory Prayer at the discretion of the Priest, who says,}
\lett{Y}{e} who do truly and earnestly repent you of your sins, and are in love and charity with your neighbours, and intend to lead a new life, following the commandments of God, and walking from henceforth in his holy ways; Draw near with faith, and take this holy Sacrament to your comfort; and make your humble confession to Almighty God, devoutly kneeling.
\rub{Then shall this General Confession be made, by the Priest and all those who are minded to receive the Holy Communion, humbly kneeling.}
\lett{A}{lmighty} God, Father of our Lord Jesus Christ, Maker of all things, Judge of all men; We acknowledge and bewail our manifold sins and wickedness, Which we, from time to time, most grievously have committed, By thought, word, and deed, Against thy Divine Majesty, Provoking most justly thy wrath and indignation against us. We do earnestly repent, And are heartily sorry for these our misdoings; The remembrance of them is grievous unto us; The burden of them is intolerable. Have mercy upon us, Have mercy upon us, most merciful Father; For thy Son our Lord Jesus Christ’s sake, Forgive us all that is past; And grant that we may ever hereafter Serve and please thee In newness of life, To the honour and glory of thy Name; Through Jesus Christ our Lord. Amen.
\rub{Then shall the Priest (the Bishop if he be present) stand up, and turning to the People, say,}
\lett{A}{lmighty} God, our heavenly Father, who of his great mercy hath promised forgiveness of sins to all those who with hearty repentance and true faith turn unto him; Have mercy upon you; pardon {\ding{64}} and deliver you from all your sins; confirm and strengthen you in all goodness; and bring you to everlasting life; through Jesus Christ our Lord. Amen.
\subsection{The Comfortable Words}
\rub{Then shall the Priest say,}
    Hear what comfortable words our Saviour Christ saith unto all who truly turn to him.\\
\par
\noindent
\textsc{Come} unto me, all ye that travail and are heavy laden, and I will refresh you.\vr{Matt. 11:28}\par
    So God loved the world, that he gave his only-begotten Son, to the end that all that believe in him should not perish, but have everlasting life.\vr{John 3:16}\\
    \par
    Hear also what Saint Paul saith.\\
    This is a true saying, and worthy of all men to be received, That Christ Jesus came into the world to save sinners.  \inrub{1 Tim. 1:15}\\
\par
    Hear also what Saint John saith.\\
    If any man sin, we have an Advocate with the Father, Jesus Christ the righteous; and he is the Propitiation for our sins.  \inrub{1 John 2:1-2}
\subsection{Prayer of Humble Access}
\lett{W}{e} do not presume to come to this thy Table, O merciful Lord, trusting in our own righteousness, but in thy manifold and great mercies. We are not worthy so much as to gather up the crumbs under thy Table. But thou art the same Lord, whose property is always to have mercy: Grant us therefore, gracious Lord, so to eat the flesh of thy dear Son Jesus Christ, and to drink his blood, that our sinful bodies may be made clean by his Body, and our souls washed through his most precious Blood, and that we may evermore dwell in him, and he in us. Amen.
\rub{The Priest first receives the Holy Communion himself while saying silent prayers. he then turns to the People, holding the Blessed Sacrament, saying,}
\elcol{℣. Behold the Lamb of God. Behold Him who taketh away the sins of the world.}{℣. Ecce Agnus Dei, ecce, qui tollit peccáta mundi.}
\elcol{℟. Lord, I am not worthy that thou shouldst come under my roof, but speak the word only, and my soul shall be healed.}{℟. Dómine, non sum dignus, ut intres sub tectum meum, sed tantum dic verbo, et sanábitur ánima mea.}
\elcol{℟. Lord, I am not worthy that thou shouldst come under my roof, but speak the word only, and my soul shall be healed.}{℟. Dómine, non sum dignus, ut intres sub tectum meum, sed tantum dic verbo, et sanábitur ánima mea.}
\elcol{℟. Lord, I am not worthy that thou shouldst come under my roof, but speak the word only, and my soul shall be healed.}{℟. Dómine, non sum dignus, ut intres sub tectum meum, sed tantum dic verbo, et sanábitur ánima mea.}
\subsection{Devotion before Communion}
\rub{Then shall the People sing or say these prayers before receiving Holy Communion,}
\lett{I}{Believe}, O Lord, and I confess that thou art truly the Christ, the Son of the Living God, who didst come into the world to save sinners, of whom I am first. And I believe that this is truly thine own immaculate Body, and that this is truly thine own precious Blood. Wherefore I pray thee: have mercy upon me and forgive my transgressions both voluntary and involuntary, of word and of deed, of knowledge and of ignorance. And make me worthy to partake without condemnation of thine immaculate Mysteries, unto the remission of my sins and unto life everlasting. Amen.
\par
Of Thy Mystical Supper, O Son of God, accept me today as a communicant; for I will not speak of Thy Mystery to thine enemies, neither will I give thee a kiss as did Judas; but like the thief will I confess Thee: Remember me, O Lord in thy Kingdom.
\par
Not unto judgement, nor unto condemnation, be my partaking of thy holy Mysteries, O Lord, but unto the healing of soul and body.
\rub{The Priest shall then say the Communion Antiphon. A hymn may then be sung.}
\subsection{The Thanksgiving Prayer}
\rub{When all have communicated, the Priest shall return to the Altar and cleanse the vessels and his hands, after which the Priest alone says,}
Let us pray.
\par
\lett{A}{lmighty} and everliving God, we most heartily thank thee, for that thou dost vouchsafe to feed us who have duly received these holy mysteries with the spiritual food of the most precious Body and Blood of thy Son our Saviour Jesus Christ; and dost assure us thereby of thy favour and goodness towards us; and that we are very members incorporate in the mystical body of thy Son, which is the blessed company of all faithful people; and are also heirs through hope of thy everlasting kingdom, by the merits of his most precious death and passion. And we humbly beseech thee, O heavenly Father, so to assist us with thy grace, that we may continue in that holy fellowship, and do all such good works as thou hast prepared for us to walk in; through Jesus Christ our Lord, to whom, with thee and the Holy Ghost, be all honour and glory, world without end. Amen.
\subsection{The Dismissal}
\elcol{℣. The Lord be with you.}{℣. Dóminus vobíscum.}
\elcol{℟. And with thy spirit.}{℟. Et cum spíritu tuo.}
\elcol{℣. Let us pray.\\}{℣. Orémus.\\}

\rub{The Priest then prays the Postcommunion Prayers.}
\elcol{℣. The Lord be with you.}{℣. Dóminus vobíscum.}
\elcol{℟. And with thy spirit.}{℟. Et cum spíritu tuo.}
	\rub{On days when the \textit{Gloria in exclesis} is said,}
	\elcol{℣. Depart in peace.}{℣. Ite, Missa est.}
	\elcol{℟. Thanks be to God.}{℟. Deo grátias.}
	\rub{On days when the \textit{Gloria in exclesis} is not said,}
	\elcol{℣. Let us bless the Lord.}{℣. Benedicámus Dómino.}
	\elcol{℟. Thanks be to God.}{℟. Deo grátias.}
	\rub{In Requiem Masses,}
	\elcol{℣. May they rest in peace.}{℣. Requiéscant in pace.}
	\elcol{℟. Amen.\\}{℟. Amen.\\}
	
	\rub{Then, the people kneeling, the Priest (the Bishop if he be present) shall let them depart with this Blessing, unless it be a Requiem Mass.}
\lett{T}{he} Peace of God, which passeth all understanding, keep your hearts and minds in the knowledge and love of God, and of his Son Jesus Christ our Lord: And the Blessing of God Almighty, the Father, {\ding{64}} the Son, and the Holy Ghost, be amongst you, and remain with you always. Amen.
\subsection{The Last Gospel}
\elcol{℣. The Lord be with you.}{℣. Dóminus vobíscum.}
\elcol{℟. And with thy spirit.}{℟. Et cum spíritu tuo.}
\elcol{℣. The beginning of the Holy Gospel according to John.}{℣. Inítium sancti Evangélii secúndum Joánnem}
\elcol{℟. Glory be to thee, O Lord.}{℟. Glória tibi, Dómine.}
\elcol{In the beginning was the Word, and the Word was with God, and the Word was God. The same was in the beginning with God. All things were made by him; and without him was not any thing made that was made. In him was life; and the life was the light of men. And the light shineth in darkness; and the darkness comprehended it not. There was a man sent from God, whose name was John. The same came for a witness, to bear witness of the Light, that all men through him might believe. He was not that Light, but was sent to bear witness of that Light. That was the true Light, which lighteth every man that cometh into the world. He was in the world, and the world was made by him, and the world knew him not. He came unto his own, and his own received him not. But as many as received him, to them gave he power to become the sons of God, even to them that believe on his name: Which were born, not of blood, nor of the will of the flesh, nor of the will of man, but of God. \inrub{All genuflect.} And the Word was made flesh, and dwelt among us, \inrub{All rise.} (and we beheld his glory, the glory as of the only begotten of the Father,) full of grace and truth.}{In princípio erat Verbum, et Verbum erat apud Deum, et Deus erat Verbum. Hoc erat in princípio apud Deum. Omnia per ipsum facta sunt: et sine ipso factum est nihil, quod factum est: in ipso vita erat, et vita erat lux hóminum: et lux in ténebris lucet, et ténebræ eam non comprehendérunt.
Fuit homo missus a Deo, cui nomen erat Joánnes. Hic venit in testimónium, ut testimónium perhibéret de lúmine, ut omnes créderent per illum. Non erat ille lux, sed ut testimónium perhibéret de lúmine.
Erat lux vera, quæ illúminat omnem hóminem veniéntem in hunc mundum. In mundo erat, et mundus per ipsum factus est, et mundus eum non cognóvit. In própria venit, et sui eum non recepérunt. Quotquot autem recepérunt eum, dedit eis potestátem fílios Dei fíeri, his, qui credunt in nómine ejus: qui non ex sanguínibus, neque ex voluntáte carnis, neque ex voluntáte viri, sed ex Deo nati sunt. \inrub{Genuflectit dicens:} Et Verbum caro factum est, \inrub{Et surgens prosequitur:} et habitávit in nobis: et vídimus glóriam ejus, glóriam quasi Unigéniti a Patre, plenum grátiæ et veritátis. 
}
\elcol{℟. Thanks be to God.}{℟. Deo grátias.}